### 2016-07-13 paper.tex
First commit

% ---------------------------------------------------------------------------
% Author guideline and sample document for EG publication using LaTeX2e input
% D.Fellner, v1.13, Jul 31, 2008

\documentclass{egpubl}
\usepackage{eg2015}

% --- for  Annual CONFERENCE
%\ConferenceSubmission   % uncomment for Conference submission
\ConferencePaper        % uncomment for (final) Conference Paper
% \STAR                   % uncomment for STAR contribution
% \Tutorial               % uncomment for Tutorial contribution
% \ShortPresentation      % uncomment for (final) Short Conference Presentation
% \Areas                  % uncomment for Areas contribution
% \MedicalPrize           % uncomment for Medical Prize contribution
% \Education              % uncomment for Education contribution
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to Computer Graphics Forum, special issue
% \SpecialIssuePaper         % uncomment for final version of Journal Paper, special issue
%
% --- for  EG Workshop Proceedings
% \WsSubmission    % uncomment for submission to EG Workshop
% \WsPaper         % uncomment for final version of EG Workshop contribution
%
 \electronicVersion % can be used both for the printed and electronic version

% !! *please* don't change anything above
% !! unless you REALLY know what you are doing
% ------------------------------------------------------------------------

% for including postscript figures
% mind: package option 'draft' will replace PS figure by a filname within a frame
%\ifpdf \usepackage[pdftex]{graphicx} \pdfcompresslevel=9
%\else \usepackage[dvips]{graphicx} \fi

\PrintedOrElectronic

% prepare for electronic version of your document
\usepackage{t1enc,dfadobe}

\usepackage{egweblnk}
\usepackage{cite}
\usepackage{balance}

% For backwards compatibility to old LaTeX type font selection.
% Uncomment if your document adheres to LaTeX2e recommendations.
% \let\rm=\rmfamily    \let\sf=\sffamily    \let\tt=\ttfamily
% \let\it=\itshape     \let\sl=\slshape     \let\sc=\scshape
% \let\bf=\bfseries

% end of prologue

%\input{EGauthorGuidelines-body.inc}
% ---------------------------------------------------------------------
% EG author guidelines plus sample file for EG publication using LaTeX2e input
% D.Fellner, v1.17, Sep 23, 2010

\title[PGVD]%
      {An Adaptive, Parallel Algorithm for Approximating the Generalized Voronoi Diagram}

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
% For Computer Graphics Forum: Please use the abbreviation of your first name.

%% author example
%\author[D. Fellner \& S. Behnke]
%       {D.\,W. Fellner\thanks{Chairman Eurographics Publications Board}$^{1,2}$
%        and S. Behnke$^{2}$
%        S. Spencer$^2$\thanks{Chairman Siggraph Publications Board} \\
%         $^1$TU Darmstadt \& Fraunhofer IGD, Germany \\
%         $^2$Institut f{\"u}r ComputerGraphik \& Wissensvisualisierung, TU Graz, Austria
%       }

\author[J. Edwards \& N. Vollmer \& N. Harrison]
%       {John Edwards\thanks{jedwards@sci.utah.edu}$^1$ and
%       {John Edwards\\
       {John Edwards and
         Nathan Vollmer and
         Nicholas Harrison \\
         \parbox[t]{10cm}{\centering Idaho State University}
%         $^2$Google, Inc. \\
%         \footnotesize{$^3$The University of Texas at Austin}}
       }
%\author[paper1040]
%       {paper1040}

% ------------------------------------------------------------------------

% if the Editors-in-Chief have given you the data, you may uncomment
% the following five lines and insert it here
%
% \volume{27}   % the volume in which the issue will be published;
% \issue{1}     % the issue number of the publication
% \pStartPage{1}      % set starting page

\renewcommand{\paragraph}[1]{\noindent \textbf{#1}}
\input{common}

%-------------------------------------------------------------------------
\begin{document}

%\teaser{
%  \subfloat[][]{
%    \label{fig:gears1}
%    \begin{tikzpicture}
%      \node[anchor=south west,inner sep=0] at (0,0) {
%        \begin{tabular}[b]{c}
%          \includegraphics[height=0.8in]{gears-far1.png} \\
%          \includegraphics[height=0.8in]{gears-close1.png}
%        \end{tabular}
%      };
%      \draw[black,thick] (1.4,3.2) rectangle (2.0,3.6);
%      \draw[black,dashed] (1.4,3.6) -- (0.22,2.15);
%      \draw[black,dashed] (2.0,3.6) -- (2.95,2.15);
%      \draw[black,thick] (0.22,0.1) rectangle (2.95,2.15);
%    \end{tikzpicture}
%  }
%  \subfloat[][]{
%    \label{fig:gears}
%    \begin{tikzpicture}
%      \node[anchor=south west,inner sep=0] at (0,0) {
%        \begin{tabular}[b]{c}
%          \includegraphics[height=0.8in]{gears-far4.png} \\
%          \includegraphics[height=0.8in]{gears-close4.png}
%        \end{tabular}
%      };
%      \draw[black,thick] (1.3,3.1) rectangle (1.9,3.5);
%      \draw[black,dashed] (1.3,3.5) -- (0.22,2.15);
%      \draw[black,dashed] (1.9,3.5) -- (2.92,2.15);
%      \draw[black,thick] (0.22,0.1) rectangle (2.92,2.15);
%    \end{tikzpicture}
%  }
%  \hspace{3mm}
%  \subfloat[][]{
%    \label{fig:knife}
%    \includegraphics[trim=4cm 0cm 4cm 2.5cm, clip=true, height=1.4in]
%                    {knife-above/slice-00000.png}
%  }
%  \subfloat[][]{
%    \label{fig:knife2}
%    \includegraphics[trim=2mm 0cm 2mm 2cm, clip=true, height=1.4in]
%                    {knife-above/slice-00110.png}
%  }
%  \caption{Two example applications of the \red{approximated} generalized Voronoi diagram (GVD) computed by our novel, adaptive algorithm. Previous GVD methods require a gridded space of $2^{24}$ (gears dataset) and $2^{36}$ (knives dataset) voxels to resolve the closely spaced objects.
%    \protect\subref{fig:gears1} Two gears with regions of very tight spacing.
%    \protect\subref{fig:gears} The GVD of the gears model.  The surface is colored red in areas of very close tolerance.
%    \protect\subref{fig:knife} Three butter knives in a wood block.  To animate removal of the knives without intersecting the block requires extreme care because of close mesh spacing.
%    \protect\subref{fig:knife2} Intersection-free motion is guaranteed by computing motion vectors based on the GVD and allowing motion only within a Voronoi cell.
%  }
%  \label{fig:teaser}
%}
%
\maketitle

\begin{abstract}
We present a parallel algorithm that approximates the Generalized Voronoi Diagram (GVD) on arbitrary collections of 2D or 3D geometric objects. Our algorithm is hierarchical, not uniformly gridded, enabling us to compute the GVD on datasets with extremely close object tolerances, including intersections. We demonstrate our method on a variety of data and show example applications of the GVD in 2D and 3D. \\

%   Leave one blank line after the abstract, 
%   then add the subject categories according to the ACM Classification Index 
%   (see http://www.acm.org/class/1998/).

\begin{classification} % according to http://www.acm.org/class/1998/
  \CCScat{I.3.5}{Computer Graphics}{Computational Geometry and Object Modeling}{Boundary representations}
  \CCScat{I.3.6}{Computer Graphics}{Methodology and Techniques}{Graphics data structures and data types}
  %\CCScat{Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
\end{classification}

\end{abstract}

%-------------------------------------------------------------------------
% Body
%-------------------------------------------------------------------------

%-------------------------------------------------------------------------------
% introduction
%-------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
Previous paper: \cite{edwards2015approximating}

\red{The generalized Voronoi diagram (GVD) is an important structure that divides space into a complex of generalized Voronoi cells (GVCs) around objects.  Similar to the ordinary Voronoi diagram, each GVC contains exactly one object, or site, and every point in the GVC is closer to its contained object than to any other object.  The generalized Voronoi diagram is the boundary of the cell complex, and thus every point on the GVD is equidistant from two or more closest objects.  Applications of the GVD range from motion path planning to GIS analysis to mosaicking.

Ordinary Voronoi diagrams have been studied extensively and efficient algorithms exist to compute them, but the GVD is difficult to compute analytically in general \cite{boissonnat2006curved,hoff1999fast} and so the majority of approaches compute an approximation.  Whereas most algorithms are efficient and robust on certain datasets, all algorithms to our knowledge require inordinate amounts of memory on datasets where objects are very closely spaced relative to the size of the domain.  The failures occur because the space is uniformly gridded.  In such approaches, voxel size must be small enough to resolve object spacings, and if two objects are very close to each other the number of voxels can become prohibitively large.

We present an algorithm to compute a GVD approximation on arbitrary datasets, including those with closely spaced objects.  The approach applies a distance transform over an octree representation of the objects.  Our octree, its associated data structure, and our distance transform are novel and optimized to GVD \red{approximation. For the remainder of the paper, ``GVD'' will refer to the approximated Generalized Voronoi Diagram.}

This paper demonstrates GVD computation on data beyond the computational abilities of previous algorithms, unlocking interesting and important applications.  Our approach allows GVD-based proximity queries and other applications using a larger class of meaningful datasets.
}

\paragraph{Main contributions} The primary technical contributions described in the paper are as follows.

\begin{enumerate}
\item Contribution 1
\item Contribution 2
\item Contribution 3
\end{enumerate}

We demonstrate various applications of the GVD...

Our GVD algorithm has a few main steps: 1) ...

%-------------------------------------------------------------------------------
% Related work
%-------------------------------------------------------------------------------
\section{Related work}
Related work falls into two categories: algorithms that compute the GVD and algorithms that compute distance fields, many of which are adaptive.

\paragraph{Generalized Voronoi diagrams}
A theoretical framework for generalized Voronoi diagrams can be found in Boissonnat \etal \shortcite{boissonnat2006curved}. Ordinary Voronoi diagrams are well studied and efficient algorithms exist that compute them exactly \cite{de2008computational}, but exact algorithms for the generalized Voronoi diagram are limited to a small set of special cases \cite{lee1982medial,karavelas2004robust}. In an early work, Lavender \etal \shortcite{lavender1992voronoi} define and compute GVDs over a set of solid models using an octree.  Etzion and Rappoport \shortcite{etzion2002computing} represent the GVD bisector symbolically for lazy evaluation, but are limited to sites that are polyhedra.  Boada \etal \shortcite{boada2002voronoi,boada2008approximations} use an adaptive approach to GVD computation, but their algorithm is restricted to GVDs with connected regions and is inefficient for polyhedral objects with many facets.  Two other works are adaptive \cite{teichmann1997polygonal,vleugels1998approximating} but are computationally expensive and are restricted to convex sites.

In recent years Voronoi diagram algorithms that take advantage of fast graphics hardware have become more common \cite{cao2010parallel,fischer2006fast,hsieh2005simple,rong2007variants,sud2006interactive,sud2006fast,hoff1999fast,wu2008gpu}.  These algorithms are efficient and generalize well to the GVD, but most are limited to a subset of site types.  More importantly, all of them use uniform grids and require an extraordinary number of voxels to resolve closely spaced objects (for example, Figs. \ref{fig:knife} and \ref{fig:path} would require $2^{36}$ and $2^{48}$ voxels, respectively).  To our knowledge, ours is the first fully adaptive algorithm that computes the generalized Voronoi diagram for arbitrary datasets.

\paragraph{Distance fields and octrees}
The GVD is a subset of the locus of distance field critical points, a property that we take advantage of. In that light, the GVD could be a post-processing step to any method that computes a distance field.  Distance transforms compute a distance field, but most are uniformly gridded \cite{jones20063d} and are thus no more suitable than GVD algorithms that use the GPU.

% JME: note that in the Bastos work, ``reconstruction'' is reconstruction
% of the surface from the octree at rendering time.
Two seminal works adaptively compute the Adaptive Distance Field (ADF) on octree vertices.  Strain~\shortcite{strain1999fast} fully resolves the octree everywhere on the object surface, and Frisken \etal~\shortcite{frisken2000adaptively} resolve the octree fully only in areas of small local feature size.  Both approaches are designed to retain features of a single object rather than resolving between multiple objects, as is required for GVD computation.  Qu \etal \shortcite{qu2004feature} implement an energy-minimizing distance field algorithm that preserves features at the expense of efficiency.  Many recent works on fast octree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}. Most octree approaches that support surfaces \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed} are designed for efficient rendering, and actual construction of the octree is implemented on the CPU.

Two works \cite{bastos2008gpu,park2010cuda} implement the ADF using GPU parallelism to compute the distance value at sample points, but building the octree itself is done sequentially.  Yin \etal~\shortcite{yin2011fast} compute the distance field entirely on the GPU using a bottom-up approach by initially subdividing into a complete octree, resulting in memory usage that is no better than using a uniform grid.  A method by Kim and Liu~\shortcite{kim2014exact} computes the octree and a BVH entirely on the GPU. However, octree construction is performed on barycenters of triangles, and so a leaf octree cell can have an arbitrary number of triangle intersections as long as it contains no more than one triangle's barycenter.  We have found no GPU octree construction method that can resolve between objects.

%-----------------------------------------------------------
% Build octree
%-----------------------------------------------------------
\section{Build octree}
\label{sec:build-octree}
\red{Our algorithm works in both 2D and 3D. Lacking a dimension-independent term, we use ``octree'' as a general term to refer to both quadtrees and octrees.}

In the algorithm below, we use the Karras algorithm as a subroutine. It takes vertices of the objects as data points. Given enough parallel units, it runs in $O(\log{N})$ time, where $N$ is the number of vertices.

We use $\container{l}{O}$ to denote the smallest octree cell in octree $O$ that fully contains the geometric object $l$. The complexity of the $\container{l}{O}$ function is $O(\log{N})$. Thus, step 2 of the algorithm (lines \ref{alg:octree_containing_begin}-\ref{alg:octree_containing_end}) runs in $O(\log{N})$ time.

Step 3 of the algorithm (lines \ref{alg:octree_intersections_begin}-\ref{alg:octree_intersections_end}) runs in $O(L)$ time, where $L$ is the number of lines in all objects. Even though the loop is doubly-nested, each line is stored in a unique internal node, so no more than $L$ lines will be visited in the loops. In practice, far fewer than $L$ lines will be checked for each leaf cell, because most datasets have lines that are completely contained in internal cells that are reasonably low in the tree.

Steps 1-3 are $O(\log{N}+L) = O(2L) = O(L)$. However, average case is $O(\log{N})$, considering that most lines are contained entirely in a cell at reasonably low depth.

In Step 4, Stack is preallocated to size $M\cdot2^D$ where $M$ is the maximum octree depth and $D$ is the dimension. A conflict cell is a cell that intersects at least two different objects, or two lines of different labels.

%\SetKwFor{ForPar}{for}{do in parallel}{endfpar}

\algorithmspace
\begin{algorithm}
  \DontPrintSemicolon
  \LinesNumbered
  \KwIn{Objects, Vertices}
  \BlankLine
  \tcp{1. Build initial octree IOctree}
  IOctree := karras(Vertices)\;

  \tcp{2. Find containing internal cells}
  \ForPar{line l in Objects}{
    \label{alg:octree_containing_begin}
    c := $\container{l}{IOctree}$\;
    c.numLines++\;
  }
  Allocate c.lines using parallel prefix sums (scan)\;
  \ForPar{line l in Objects}{
    c := $\container{l}{IOctree}$\;
    c.lines.append(l)\;
  } \label{alg:octree_containing_end}
  \tcp{3. Find line-leaf cell intersections}
  \ForPar{leaf cell c in IOctree}{
    \label{alg:octree_intersections_begin}
    \ForEach{cell a in direct\_ancestors(c)}{
      \ForEach{line l in a.lines}{
        \If{l intersects c}{
          c.numLines++\;
        }
      }
    }
  }
  Allocate c.lines using parallel prefix sums (scan)\;
  \ForPar{leaf cell c in IOctree}{
    \ForEach{cell a in direct\_ancestors(c)}{
      \ForEach{line l in a.lines}{
        \If{l intersects c}{
          c.lines.append(l)\;
        }
      }
    }
  } \label{alg:octree_intersections_end}

  \tcp{4. Octree refinement}
  \ForPar{leaf cell c in IOctree}{
    c' := c\;
    \While{c' is conflict cell}{
      $(c'_0,c'_1,\dots,c'_{2^D-1})$ := subdivide c'\;
      push $(c'_0,c'_1,\dots,c'_{2^D-1})$ onto Stack\;
      c' := Stack.pop
    }
  }
\caption{BUILD\_OCTREE}
\label{alg:octree}
\end{algorithm}
\algorithmspace

\begin{figure}
  \centering
  \subfloat[][]{
    \label{fig:octree-cartesian-1}
    \includegraphics[width=0.48\columnwidth]{octree-cartesian-1.pdf} }
  \subfloat[][]{
    \label{fig:octree-hierarchical-1}
    \includegraphics[width=0.48\columnwidth]{octree-hierarchical-1.pdf} } \\
  \subfloat[][]{
    \label{fig:octree-cartesian-2}
    \includegraphics[width=0.48\columnwidth]{octree-cartesian-2.pdf} }
  \subfloat[][]{
    \label{fig:octree-hierarchical-2}
    \includegraphics[width=0.48\columnwidth]{octree-hierarchical-2.pdf} }
  \caption{The $\container{l}{O}$ function finds the smallest octree cell that completely contains line $l$.
    \protect\subref{fig:octree-cartesian} Cartesian view of the octree with a red and a blue object.
    \protect\subref{fig:octree-hierarchical} Hierarchical view of the octree.
  }
  \label{fig:steps}
\end{figure}

In Fig. \ref{fig:steps}, R (Root) is the smallest containing cell for lines A, B, and C, cell 20 contains line D, and cell 2 contains lines E and F. After Step 3 of the algorithm, line A is stored in leaf cells 202, 203, 21, and 3. Conflict cells, which are the only cells that are subdivided, are 203 and 21.

%\begin{table}
%  \centering
%  \begin{tabular}{ll}
%    \toprule
%    line &  $\container{line}{O}$ \\
%    \midrule
%    A & R \\
%    B & R \\
%    C & R \\
%    D & 20 \\
%    E & 2 \\
%    F & 2 \\
%    \bottomrule
%  \end{tabular}
%\end{table}

%-----------------------------------------------------------
% Compute GVD surface
%-----------------------------------------------------------
\section{Compute GVD surface}
\label{sec:bisector}

%-------------------------------------------------------------------------------
% Results
%-------------------------------------------------------------------------------
\section{Results and applications}
Our implementation\footnote{Source code is available at \url{http://cedmav.org/research/project/33-gvds.html}.} of the algorithm supports \red{polygons and} triangulated objects, and our wavefront initialization step is implemented on the GPU using OpenCL. All tests were run on a MacBook Pro laptop with a dual-core 2.9 GHz processor, 8 GB memory, and Intel HD 4000 graphics card. Figure \ref{fig:bunny} shows our implementation of the GVD computation pipeline, and Figure \ref{fig:pipes} shows the computed GVD on a more challenging dataset.  We compare our method with other work and then show examples in three application settings: path planning, proximity queries, and exploded diagrams.

\subsection{Comparison to other methods}

% # octree vertices
%  52K ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
% 170K ./viewer3 -l 12 ~/data/knife/knife-holder.obj ~/data/knife/knife[2-4].obj
% 146K ./viewer2 -l 24 ../data2/filled_ut/*.dat
% 2.8M ./viewer3 --uniform-colors -a 1 -l 8 ~/data/neuron/improved/a*.obj
% 1.3M ./viewer3 -l 8 ~/data/rice-dwarf/decimated2/n1UF2a-*.obj

% memory (54 bytes per octree cell)
%  2.8Mb ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
%  9.2Mb ./viewer3 -l 12 ~/data/knife/knife-holder.obj ~/data/knife/knife[2-4].obj
% 7.9Mb ./viewer2 -l 24 ../data2/filled_ut/*.dat
% 151.2Mb ./viewer3 --uniform-colors -a 1 -l 8 ~/data/neuron/improved/a*.obj
% 70.2Mb ./viewer3 -l 8 ~/data/rice-dwarf/decimated2/n1UF2a-*.obj

\begin{table*}
  \centering
  \footnotesize{
  \begin{tabular}{l c c c c c c c}
    \toprule
    dataset & objects & object          & octree   & octree          & octree &
    GVD   & GVD             \\
            &         & $\Delta$s       & depth    & cells           & memory &
    (sec) & $\Delta$s       \\
            &         & ($\times 10^3$) &          & ($\times 10^3$) & (Mb)   &
          & ($\times 10^3$) \\
    \midrule
    % ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
    Fig. \ref{fig:gears} & 3 & 7 & 8 & 54 & 3 & 0.9 & 83\\
    % ./viewer3 -l 12 ~/data/knife/knife-holder.obj ~/data/knife/knife[2-4].obj
    Fig. \ref{fig:knife} & 4 & 15 & 12 & 146 & 9 & 3.9 & 232 \\
    % ./viewer2 -l 24 ../data2/filled_ut/*.dat
    Fig. \ref{fig:path} & 470 & 5 & 24 & 158 & 8 & 2.0 & 151 \\
    % ./viewer3 --uniform-colors -a 1 -l 8 ~/data/neuron/improved/a*.obj
    Fig. \ref{fig:axons} & 448 & 4015 & 8 & 2716 & 151 & 195 & 8100 \\
    % ./viewer3 -l 8 ~/data/rice-dwarf/decimated2/n1UF2a-*.obj
    Fig. \ref{fig:mol-explode} & 35 & 1500 & 8 & 496 & 70 & 19 & 2700 \\
    \bottomrule
  \end{tabular}}
  \caption{Table of octree/GVD computation statistics and timings on datasets that are unmanageable using other methods. \red{Columns are: \emph{objects} - the number of objects in the dataset; \emph{object $\Delta$s} - the number of line segments (2D) or triangles (3D) of all objects in the dataset; \emph{octree depth} - required octree depth in order to resolve objects; \emph{octree cells} - total number of leaf octree cells; \emph{octree memory} - amount of memory used by the octree; \emph{GVD (sec)} - seconds to perform all steps of GVD computation; \emph{GVD $\Delta$s} - number of line segments (2D) or triangles (3D) in the GVD.}}
  \label{tab:timings}
\end{table*}

%-------------------------------------------------------------------------------
% Conclusions
%-------------------------------------------------------------------------------
\section{Conclusions}

%-------------------------------------------------------------------------------
% Acknowledgements
%-------------------------------------------------------------------------------
%\section*{Acknowledgements}
%Thanks to Kristen Harris for use of the neuronal data and Jonathan Bronson for the heart data. The work of JE and VP was supported in part by NSF IIS-1314896, NSF ACI-0904631, DOE/NEUP 120341, DOE/UV-CDAT DESC0006872, DOE/Codesign P01180734, DOE/SciDAC DESC0007446, DOE/PIPER DESC0010498, and DOE/CCMSC DENA0002375. This work initiated at the University of Texas when JE, ED  and CB were supported in part by NIH contract R01-EB00487, NSF Grant OCI-1216701 and SNL contract 1439100.

%-------------------------------------------------------------------------
% Bibliography
%-------------------------------------------------------------------------

%\bibliographystyle{eg-alpha}
\bibliographystyle{eg-alpha-doi}
\balance
\bibliography{paper}

\end{document}

### 2016-07-28 paper.tex
Added derivation for resolution of conflict cells using closed solution

@@
 \usepackage{balance}
+\usepackage{color}
 
@@
 \title[PGVD]%
-      {An Adaptive, Parallel Algorithm for Approximating the Generalized Voronoi Diagram}
+      {Parallel Octree Construction on Collections of Objects}
 
@@
 \begin{abstract}
-We present a parallel algorithm that approximates the Generalized Voronoi Diagram (GVD) on arbitrary collections of 2D or 3D geometric objects.
-Our algorithm is hierarchical, not uniformly gridded, enabling us to compute the GVD on datasets with extremely close object tolerances, including intersections.
-We demonstrate our method on a variety of data and show example applications of the GVD in 2D and 3D.
-\\
+We present a parallel octree algorithm that resolves between geometric objects.
+The octree has the property that no octree cell intersects more than one labeled object.
+Previous parallel algorithms either spawn kernels hierarchically, separate points only, or make no hard guarantees of object separation.
+Our algorithm runs in \red{complexity?} in the average case and has excellent results in practice.
+We demonstrate with results on 2D and 3D datasets.
 
@@
 
-\red{The generalized Voronoi diagram (GVD) is an important structure that divides space into a complex of generalized Voronoi cells (GVCs) around objects.
+The generalized Voronoi diagram (GVD) is an important structure that divides space into a complex of generalized Voronoi cells (GVCs) around objects.
 Similar to the ordinary Voronoi diagram, each GVC contains exactly one object, or site, and every point in the GVC is closer to its contained object than to any other object.
@@
 The approach applies a distance transform over an octree representation of the objects.
-Our octree, its associated data structure, and our distance transform are novel and optimized to GVD \red{approximation.
-For the remainder of the paper, ``GVD'' will refer to the approximated Generalized Voronoi Diagram.}
+Our octree, its associated data structure, and our distance transform are novel and optimized to GVD approximation.
+For the remainder of the paper, ``GVD'' will refer to the approximated Generalized Voronoi Diagram.
 
@@
 Our approach allows GVD-based proximity queries and other applications using a larger class of meaningful datasets.
-}
 
-\paragraph{Main contributions} The primary technical contributions described in the paper are as follows.
+Our algorithm has three steps:
 
 \begin{enumerate}
-\item Contribution 1
-\item Contribution 2
-\item Contribution 3
+\item Construct an octree on object vertices using Karras' algorithm \cite{karras2012maximizing}
+\item Detect octree cells that intersect more than one object, which we call ``conflict cells'' (contribution)
+\item Subdivide conflict cells to resolve objects (contribution)
 \end{enumerate}
-
-We demonstrate various applications of the GVD...
-
-Our GVD algorithm has a few main steps: 1) ...
 
@@
 %-----------------------------------------------------------
-\section{Build octree}
-\label{sec:build-octree}
-\red{Our algorithm works in both 2D and 3D. Lacking a dimension-independent term, we use ``octree'' as a general term to refer to both quadtrees and octrees.}
+\section{Algorithm}
+\label{sec:algorithm}
+Our algorithm works in both 2D and 3D.
+Lacking a dimension-independent term, we use ``octree'' as a general term to refer to both quadtrees and octrees.
+We refer to object facets, which are lines in 2D and triangles in 3D.
 
-In the algorithm below, we use the Karras algorithm as a subroutine.
-It takes vertices of the objects as data points.
-Given enough parallel units, it runs in $O(\log{N})$ time, where $N$ is the number of vertices.
+\subsection{Resolving conflict cells}
 
-We use $\container{l}{O}$ to denote the smallest octree cell in octree $O$ that fully contains the geometric object $l$.
-The complexity of the $\container{l}{O}$ function is $O(\log{N})$.
-Thus, step 2 of the algorithm (lines \ref{alg:octree_containing_begin}-\ref{alg:octree_containing_end}) runs in $O(\log{N})$ time.
+%\begin{adjustbox}{valign=t}
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:conflict-resolution-x}
+    \includegraphics[width=0.48\columnwidth]{conflict-resolution-x.pdf} }
+  \subfloat[][]{
+    \label{fig:conflict-resolution-even}
+    \includegraphics[width=0.48\columnwidth]{conflict-resolution-even.pdf} } \\
+  \subfloat[][]{
+    \label{fig:conflict-resolution-all}
+    \includegraphics[width=0.48\columnwidth]{conflict-resolution-all.pdf} }
+  \subfloat[][]{
+    \label{fig:conflict-resolution-octree}
+    \includegraphics[width=0.48\columnwidth]{conflict-resolution-octree-samples.pdf} }
+  \caption{
+    \protect\subref{fig:conflict-resolution-x} A conflict cell with two lines from different objects.
+    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least one sample (red dots).
+    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least two samples.
+This ensures that an octree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-octree}) will have no leaf cells that intersect both lines, ensuring that the new octree is locally free of conflict cells.
+  }
+  \label{fig:conflict-resolution}
+\end{figure}
+%    \end{adjustbox}
 
-Step 3 of the algorithm (lines \ref{alg:octree_intersections_begin}-\ref{alg:octree_intersections_end}) runs in $O(L)$ time, where $L$ is the number of lines in all objects.
-Even though the loop is doubly-nested, each line is stored in a unique internal node, so no more than $L$ lines will be visited in the loops.
-In practice, far fewer than $L$ lines will be checked for each leaf cell, because most datasets have lines that are completely contained in internal cells that are reasonably low in the tree.
+%\begin{equation}
+%\label{eqn:p}
+%p(s) = p = p_0 + su
+%\end{equation}
+%\begin{equation}
+%\label{eqn:q}
+%q(s) = q = q_0 + tv
+%\end{equation}
+%\begin{equation}
+%\label{eqn:a_}
+%a(s) = a = |p^x-q^x| = |p^y-q^y|
+%\end{equation}
 
-Steps 1-3 are $O(\log{N}+L) = O(2L) = O(L)$. However, average case is $O(\log{N})$, considering that most lines are contained entirely in a cell at reasonably low depth.
+\begin{align}
+p(s) &= p = p_0 + su \label{eqn:p} \\
+q(t) &= q = q_0 + tv \label{eqn:q}
+%a(s) = a = |p^x-q^x| = |p^y-q^y| \label{eqn:a}
+\end{align}
+For the case that the bisecting line has slope $0 \le m \le 1$, we wish to find $a = q^x-p^x$, which results in a box that touches both lines.
+Given a point $p$, we find $q$ by substituting in equations \eqref{eqn:p} and \eqref{eqn:q} and solving for $t$,
+\begin{equation}
+\label{eqn:abc}
+t = \frac{p^x+p^y-(q_0^x+q_0^y)}{v^x+v^y}
+\end{equation}
+Thus,
+\begin{align} 
+a(s) &= q^x(t) - p^x(s) \nonumber \\
+   &= q_0^x + v^x\frac{p^x(s)+p^y(s)-(q_0^x+q_0^y)}{v^x+v^y} - p^x(s) \nonumber \\
+%   &= \frac{q_0^xv^y-q_0^yv^x+p^yv^x-p^xv^y}{v^x+v^y} \nonumber \\
+   &= \frac{q_0^xv^y-q_0^yv^x+p^y(s)v^x-p^x(s)v^y}{v^x+v^y} \nonumber \\
+   &= \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y+s(u^yv^x-u^xv^y)}{v^x+v^y} \nonumber \\
+   &= s\frac{u^yv^x-u^xv^y}{v^x+v^y}+\frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{v^x+v^y} \label{eqn:a}
+\end{align}
+
+%$a$ is the largest power of two less than or equal to $a'$:
+%\begin{equation}
+%\label{eqn:ghi}
+%a(s) = a = 2^{\lfloor \log_2{a'} \rfloor}
+%\end{equation}
+
+As shown in figure \ref{fig:conflict-resolution}, we desire to build a sequence $S=(s_0, s_1, \dots)$ such that $p^x(s_i) = p^x(s_{i-2}) + a(s_{i-2})$.
+%For convenience, we split the sequence into even and odd components and define $p^x(s_i) = p^x(s_{i-2}) + a(s_{i-2})$.
+That is,
+\begin{align}
+x_i &= x_{i-2} + a(s_{i-2}) \label{eqn:xi_init}
+\end{align}
+Substituting in equation \eqref{eqn:a},
+\begin{equation}
+x_i = x_{i-2} + s_{i-2}\frac{u^yv^x-u^xv^y}{v^x+v^y} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{v^x+v^y}
+\end{equation}
+%\begin{align}
+%x_i &= x_{i-1} + \frac{1}{v^x+v^y}(q_0^xv^y-q_0^yv^x+p^y(s_{i-1})v^x-p^x(s_{i-1})v^y)
+%\end{align}
+%Substituting in equation \eqref{eqn:p} and rearranging,
+%\begin{align}
+%x_i &= x_{i-1} + \frac{1}{v^x+v^y}(q_0^xv^y-q_0^yv^x+(p_0^y+s_{i-1}u^y)v^x-(p_0^x+s_{i-1}u^x)v^y) \\
+%    &= x_{i-1} + \frac{1}{v^x+v^y}(q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y+s_{i-1}(u^yv^x-u^xv^y)) \\
+%    &= x_{i-1} + s_{i-1}\frac{u^yv^x-u^xv^y}{v^x+v^y} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{v^x+v^y}
+%\end{align}
+Substituting in equation \eqref{eqn:p},
+\begin{align}
+x_i &= p_0^x+s_{i-2}u^x + s_{i-2}\frac{u^yv^x-u^xv^y}{v^x+v^y} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{v^x+v^y} \\
+    &= s_{i-2}\frac{v^x(u^x+u^y)}{v^x+v^y} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x+p_0^xv^x}{v^x+v^y} \label{eqn:xi}
+\end{align}
+We're now ready to build the sequence. From \eqref{eqn:p},
+\begin{align}
+s_i &= \frac{x_i-p_0^x}{u^x}
+\end{align}
+Substituting in equation \eqref{eqn:xi},
+\begin{align}
+s_i &= s_{i-2}\frac{v^x(u^x+u^y)}{u^x(v^x+v^y)} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{u^x(v^x+v^y)} \\
+    &= \alpha s_{i-2} + \beta \label{eqn:si_rec}
+\end{align}
+where $\displaystyle \alpha=\frac{v^x(u^x+u^y)}{u^x(v^x+v^y)}$ and $\displaystyle \beta=\frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{u^x(v^x+v^y)}$.
+
+To solve the recurrence relation \eqref{eqn:si_rec},
+\begin{align}
+s_{i+2} &= \alpha s_i+\beta \\
+s_{i+2}-s_i &= \alpha s_i-\alpha s_{i-2} \\
+s_{i+2} &= (\alpha+1) s_i-\alpha s_{i-2}
+\end{align}
+The characteristic polynomial is
+\begin{equation}
+p(t) = t^2-(\alpha+1)t + \alpha
+\end{equation}
+which has roots $\alpha$ and $1$. Thus,
+\begin{equation}
+s_i = k_1 + k_2 \alpha^i
+\end{equation}
+The initial condition when $i$ is even is $s_0=0$.
+When $i$ is odd, the initial condition will be where $s$ is halfway along the first box (it could be anywhere in the interior of the first box, but we use halfway for convenience).
+So from \eqref{eqn:p},
+\begin{align}
+s_1 &= \frac{p^x-p_0^x}{2u^x} \\
+    &= \frac{a(0)}{2u^x}
+\end{align}
+\begin{align}
+k_1^{even} = k_1^{odd} &= \beta\left(1-\frac{\alpha}{\alpha-1}\right) \\
+k_2^{even} &= \frac{\beta}{\alpha-1} \\
+k_2^{odd} &= \frac{a(0)}{2u^x}+\frac{\beta}{\alpha-1}
+\end{align}
+
+It is tempting to formulate the recurrance relation using a single initial condition. This would be done by replacing equation \eqref{eqn:xi_init} with
+\begin{align}
+x_i &= x_{i-1} + \frac{a(s_{i-1})}{2}
+\end{align}
+This formulation can lead to conflict cells as shown in figure \ref{fig:conflict-resolution-half}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.48\columnwidth]{conflict-resolution-half.pdf}
+  \caption{Using a single recurrence solution we can no longer guarantee that every box that intersects both lines contains at least two samples.
+The box in red contains only one sample, so the octree may contain conflict cells.
+  }
+  \label{fig:conflict-resolution-half}
+\end{figure}
+
+%With $a$ in hand, the question becomes, at what value of $s$ does $a(s)$ double?
+We desire to build a sequence $S=(s_0, s_1, \dots)$ such that $a(s_i) = 2a(s_{i-1})$.
+Substituting \eqref{eqn:p} into \eqref{eqn:a_prime} we get
+%\begin{align}
+%\label{eqn:klm}
+%2a(s_{i-1}) &= \frac{q_0^xv^y-q_0^yv^x+(p_0^y+s_iu^y)v^x-(p_0^x+s_iu^x)v^y}{v^x+v^y} \nonumber \\
+%&= \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y+s_i(u^yv^x-u^xv^y)}{v^x+v^y}
+%\end{align}
+%Solving for $s_i$ results in
+%\begin{equation}
+%\label{eqn:klm}
+%s_i = \frac{2(v^x+v^y)a(s_{i-1})-q_0^xv^y+q_0^yv^x-p_0^yv^x+p_0^xv^y}{u^yv^x-u^xv^y}
+%\end{equation}
+%
+%For $i>0$, $a(s_i)=a'(s_i)$, so we can characterize $s_i$ as a linear equation:
+%\begin{align}
+%s_0 &= 0 \nonumber \\
+%s_1 &= 2ma(s_0)+b \nonumber \\
+%s_2 &= 2^2ma(s_0)+b \nonumber \\
+%\cdots \nonumber \\
+%s_i &= 2^ima(s_0)+b
+%\label{eqn:si}
+%\end{align}
+%where 
+%\[m=\frac{v^x+v^y}{u^yv^x-u^xv^y}\]
+%and
+%\[b=\frac{-q_0^xv^y+q_0^yv^x-p_0^yv^x+p_0^xv^y}{u^yv^x-u^xv^y}\]
+%%where $m=(v^x+v^y)/(u^yv^x-u^xv^y)$ and $b=(-q_0^xv^y+q_0^yv^x-p_0^yv^x+p_0^xv^y)/(u^yv^x-u^xv^y)$.
+%
+%The number of elements in sequence $S$ is obtained by setting equation \eqref{eqn:si} less than one and solving for $i$:
+%\begin{equation}
+%i < \log_2{\frac{1-b}{ma(s_0)}}
+%\end{equation}
+
+\subsection{Build octree on vertices}
+
+We first construct an octree on the vertices of the objects, which we call the ``vertex octree''.
+We use Karras' algorithm \cite{karras2012maximizing} which sorts the Morton codes of the vertices in parallel, then constructs the binary radix tree in parallel.
+With the binary radix tree, the octree can be constructed with a single parallel call.
+The strength of this algorithm lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points.
+That is, even if a large number of vertices are clustered in a small area, requiring deep octree subdivision, only a constant number of parallel calls need be made.
+Given enough parallel units, the Karras algorithm runs in $O(\log{N})$ time, where $N$ is the number of vertices.
+
+\subsection{Identify conflict cells}
+
+Our end goal is to construct an octree such that no octree cell intersects more than one object.
+Note that a cell is allowed to intersect more than one facet, but all facets must belong to the same object, or, in other words, all facets must share the same label.
+It is possible, but unlikely, that the vertex octree has this property.
+If so, then we are done.
+Otherwise, we must identify subdivide conflict cells.
+
+One naive algorithm to identify conflict cells is to process each leaf cell $c$ in parallel and store which facets intersect $c$.
+This is $O(N)$.
+Another approach is to process each facet in parallel and add it to every cell that it intersects.
+This is $O(k\log{N})$ where $k$ is maximum number of cells that any facet intersects.
+As we will show, our algorithm is $O(j + \log{N})$ where $j$ is the maximum number of facets that intersect any cell.
+In practice, $\log{N} > j$, making our algorithm $O(\log{N})$.
+
+We identify conflict cells as shown in algorithm \ref{alg:find-conflict-cells}.
+In lines \ref{alg:octree_containing_begin}-\ref{alg:octree_containing_end}, for each internal octree cell $c$, we store all facets for which $c$ is the smallest containing cell.
+Since we are implementing this in a GPGPU environment, we don't have dynamic memory, so each facet must be processed twice.
+The first loop discovers how many facets are to be stored in each cell after which we allocate space for the facets.
+We use parallel prefix sums to determine the amount of space we need to allocate as well as the offsets for each internal cell.
+The second loop actually stores the facets.
+
+The $\container{f}$ procedure finds the smallest octree cell that fully contains the facet $f$.
+A straightforward implementation of $\container{f}$ is to perform a standard octree search on the vertices of $f$ and take the smallest octree cell that contains all of them.
+(Note that the cell is always an internal node, since a post-condition of the Karras algorithm is that no leaf cell contains more than one vertex.) In our implementation however, we take advantage of our existing data structures.
+The octree cell that contains a vertex $v$ is uniquely determined by the D-tuple bits of its morton code.
+For example, if a 2D vertex has morton code $010010$, then the octree is traversed from the root to child $01$ to child $00$ to child $10$.
+To determine $\container{f}$, we find the longest common prefix (\lcp) of the vertices.
+Truncating the length of \lcp\,to a multiple of D, we find the smallest octree cell that contains all vertices of $f$.
+The complexity of $\container{f}$ is $O(\log{N})$ for both implementations.
+Thus, lines \ref{alg:octree_containing_begin}-\ref{alg:octree_containing_end} run in $O(\log{N})$ time.
+
+Lines \ref{alg:octree_intersections_begin}-\ref{alg:octree_intersections_end} of the algorithm identify and store all facets that intersect with a given leaf cell $c$.
+Again, it is done in two steps for memory allocation purposes.
+Each leaf cell $c$ looks at its $O(\log{N})$ ancestors and tests all facets stored in those ancestors for intersection with $c$.
+Any intersecting facets get stored in $c$.
+These lines run in $O(F)$ time, where $F$ is the number of facets in all objects.
+Even though the loop is doubly-nested, each facet is stored in a unique internal node, so no more than $F$ facets will be visited in the loops.
+In practice, far fewer than $F$ facets will be checked for each leaf cell, because most datasets have facets that are completely contained in internal cells that are reasonably low in the tree.
+
+The entire conflict cell detection algorithm runs in $O(\log{N}+L) = O(L)$ because $L > \log{N}$.
+However, average case is $O(\log{N})$, considering that most lines are contained entirely in a cell at reasonably low depth.
 
@@
 A conflict cell is a cell that intersects at least two different objects, or two lines of different labels.
+
+The second procedure we use is $\directAncestors{c}$, which finds all ancestors of octree cell $c$.
 
@@
   \LinesNumbered
-  \KwIn{Objects, Vertices}
+  \KwIn{VertexOctree}
   \BlankLine
-  \tcp{1. Build initial octree IOctree}
-  IOctree := karras(Vertices)\;
-
-  \tcp{2. Find containing internal cells}
-  \ForPar{line l in Objects}{
+  \tcp{Store contained facets}
+  \ForPar{facet $f$ in Objects}{
     \label{alg:octree_containing_begin}
-    c := $\container{l}{IOctree}$\;
-    c.numLines++\;
+    a := $\container{f}$\;
+    a.numFacets := a.numFacets + 1\;
   }
-  Allocate c.lines using parallel prefix sums (scan)\;
-  \ForPar{line l in Objects}{
-    c := $\container{l}{IOctree}$\;
-    c.lines.append(l)\;
+  Allocate space for facets in internal cells\;
+  \ForPar{facet $f$ in Objects}{
+    a := $\container{f}$\;
+    a.facets := a.facets $\cup$ f
   } \label{alg:octree_containing_end}
-  \tcp{3. Find line-leaf cell intersections}
-  \ForPar{leaf cell c in IOctree}{
+  \tcp{Store intersecting facets}
+  \ForPar{leaf cell c in VertexOctree}{
     \label{alg:octree_intersections_begin}
-    \ForEach{cell a in direct\_ancestors(c)}{
-      \ForEach{line l in a.lines}{
-        \If{l intersects c}{
-          c.numLines++\;
+    \ForEach{cell a in \directAncestors{c}}{
+      \ForEach{facet $f$ in a.facets}{
+        \If{$f$ intersects c}{
+          c.numFacets := c.numFacets + 1\;
         }
@@
   }
-  Allocate c.lines using parallel prefix sums (scan)\;
-  \ForPar{leaf cell c in IOctree}{
-    \ForEach{cell a in direct\_ancestors(c)}{
-      \ForEach{line l in a.lines}{
-        \If{l intersects c}{
-          c.lines.append(l)\;
+  Allocate space for facets in leaf cells\;
+  \ForPar{leaf cell c in VertexOctree}{
+    \ForEach{cell a in \directAncestors{c}}{
+      \ForEach{facet $f$ in a.facets}{
+        \If{$f$ intersects c}{
+          c.facets := c.facets $\cup$ f\;
         }
@@
   } \label{alg:octree_intersections_end}
+\caption{FIND\_CONFLICT\_CELLS}
+\label{alg:find-conflict-cells}
+\end{algorithm}
+\algorithmspace
 
+\algorithmspace
+\begin{algorithm}
+  \DontPrintSemicolon
+  \LinesNumbered
+  \KwIn{Octree, conflict\_cells}
+  \BlankLine
   \tcp{4. Octree refinement}
-  \ForPar{leaf cell c in IOctree}{
+  \ForPar{leaf cell c in Octree}{
     c' := c\;
-    \While{c' is conflict cell}{
+    \While{c' $\in$ conflict\_cells}{
       $(c'_0,c'_1,\dots,c'_{2^D-1})$ := subdivide c'\;
@@
   }
-\caption{BUILD\_OCTREE}
-\label{alg:octree}
+\caption{REFINE\_OCTREE}
+\label{alg:refine-octree}
 \end{algorithm}
@@
   \subfloat[][]{
+    \label{fig:octree-cartesian-2}
+    \includegraphics[width=0.48\columnwidth]{octree-cartesian-2.pdf} } \\
+  \subfloat[][]{
+    \label{fig:octree-cartesian-3}
+    \includegraphics[width=0.48\columnwidth]{octree-cartesian-3.pdf} }
+  \subfloat[][]{
+    \label{fig:octree-cartesian-5}
+    \includegraphics[width=0.48\columnwidth]{octree-cartesian-5.pdf} } \\
+  \subfloat[][]{
     \label{fig:octree-hierarchical-1}
-    \includegraphics[width=0.48\columnwidth]{octree-hierarchical-1.pdf} } \\
-  \subfloat[][]{
-    \label{fig:octree-cartesian-2}
-    \includegraphics[width=0.48\columnwidth]{octree-cartesian-2.pdf} }
+    \includegraphics[width=0.48\columnwidth]{octree-hierarchical-1.pdf} }
   \subfloat[][]{
@@
     \includegraphics[width=0.48\columnwidth]{octree-hierarchical-2.pdf} }
-  \caption{The $\container{l}{O}$ function finds the smallest octree cell that completely contains line $l$.
-    \protect\subref{fig:octree-cartesian} Cartesian view of the octree with a red and a blue object.
-    \protect\subref{fig:octree-hierarchical} Hierarchical view of the octree.
+  \caption{
+    \protect\subref{fig:octree-cartesian-1} A red object and a blue object.
+    \protect\subref{fig:octree-cartesian-2} The vertex octree, or octree built on the object vertices using Karras' algorithm.
+    \protect\subref{fig:octree-cartesian-3}, \protect\subref{fig:octree-hierarchical-1} The vertex octree with conflict cells highlighted.
+Note the label of an octree cell in \protect\subref{fig:octree-cartesian-3} is the concatenation of labels from root R to the leaf cell in \protect\subref{fig:octree-hierarchical-1}.
+This value also corresponds to the highest order bits of the morton code of any point in the cell.
+    \protect\subref{fig:octree-cartesian-5}, \protect\subref{fig:octree-hierarchical-2} The octree after resolution of conflict cells.
   }
@@
 %    \toprule
-%    line &  $\container{line}{O}$ \\
+%    line &  $\container{line}$ \\
 %    \midrule

### 2016-08-05 paper.tex
About to refactor

@@
 
+To resolve conflict cells we consider pairs of lines of differing labels that intersect a cell $c$. Figure \ref{fig:conflict-resolution-x} shows two lines
+
 \begin{align}
-p(s) &= p = p_0 + su \label{eqn:p} \\
-q(t) &= q = q_0 + tv \label{eqn:q}
-%a(s) = a = |p^x-q^x| = |p^y-q^y| \label{eqn:a}
+q(t) &= q = q_0 + tv \label{eqn:q} \\
+r(f) &= r = r_0 + fw \label{eqn:r}
 \end{align}
-For the case that the bisecting line has slope $0 \le m \le 1$, we wish to find $a = q^x-p^x$, which results in a box that touches both lines.
-Given a point $p$, we find $q$ by substituting in equations \eqref{eqn:p} and \eqref{eqn:q} and solving for $t$,
+along with their bisector
+\begin{align}
+p(s) &= p = p_0 + su \label{eqn:p}
+\end{align}
+Our strategy will be to sample points $P$ on $p(u)$ (figure \ref{fig:conflict-resolution-octree}) such that an octree built on $V \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
+We do this by ensuring that $P$ is sampled such that every box that intersects both $q$ and $r$ also intersects at least two points in $P$.
+Because Karras' algorithm guarantees that every leaf cell intersects at most one point, we know that no leaf cell will intersect $q$ and $r$ and thus no leaf cell will be a conflict cell.
+
+We consider only bisecting lines with slope $0 \le m \le 1$. All other cases can be transformed to this case. We begin by fitting the smallest box centered at ??? that intersects both $q$ and $r$.
+
+\subsubsection{Case $w^y>0$}
+
+We wish to find $a = q^x-p^x$, which results in a box that touches both lines. Given a point $p$, we find $q$ by substituting in equations \eqref{eqn:q} and \eqref{eqn:r} and solving for $t$,
 \begin{equation}
-\label{eqn:abc}
 t = \frac{p^x+p^y-(q_0^x+q_0^y)}{v^x+v^y}

### 2016-08-26 paper.tex
changes

@@
 \end{align}
-along with their bisector
+along with a line
 \begin{align}
@@
 
-We consider only bisecting lines with slope $0 \le m \le 1$. All other cases can be transformed to this case. We begin by fitting the smallest box centered at ??? that intersects both $q$ and $r$.
+We consider only cases where the slope of $p$ is in the range $0 \le m \le 1$.
+All other cases can be transformed to this case using rotation and reflection.
+We begin by fitting the smallest box centered on a point $p$ that intersects both $q$ and $r$.
 
@@
 
-We wish to find $a = q^x-p^x$, which results in a box that touches both lines. Given a point $p$, we find $q$ by substituting in equations \eqref{eqn:q} and \eqref{eqn:r} and solving for $t$,
+Consider a box with edge length $a_0$ centered on a point $p_0 \in p$.
+We first find $p$, which is the line that is equidistant from $q$ and $r$ along the diagonal vector $\gamma = (1, -1)$.
+Given some point $q(t)$,
 \begin{equation}
-t = \frac{p^x+p^y-(q_0^x+q_0^y)}{v^x+v^y}
+q_0+tv + \lambda\gamma = r_0+fw \\
 \end{equation}
-Thus,
+Solving for $\lambda$,
 \begin{align}
-a(s) &= q^x(t) - p^x(s) \nonumber \\
-   &= q_0^x + v^x\frac{p^x(s)+p^y(s)-(q_0^x+q_0^y)}{v^x+v^y} - p^x(s) \nonumber \\
-%   &= \frac{q_0^xv^y-q_0^yv^x+p^yv^x-p^xv^y}{v^x+v^y} \nonumber \\
-   &= \frac{q_0^xv^y-q_0^yv^x+p^y(s)v^x-p^x(s)v^y}{v^x+v^y} \nonumber \\
-   &= \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y+s(u^yv^x-u^xv^y)}{v^x+v^y} \nonumber \\
-   &= s\frac{u^yv^x-u^xv^y}{v^x+v^y}+\frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{v^x+v^y} \label{eqn:a}
+f &= \frac{1}{w^y}(q_0^y+tv^x-\lambda-r_0^y) \\
+\lambda &= r_0^x + fw^x-q_0^x-tv^x \\
+        &= r_0^x+\frac{w^x}{w^y}(q_0^y+tv^y-\lambda-r_0^y) - q_0^x-tv^x \\
+        &= \frac{1}{w^x+w^y}(r_0^xw^y+q_0^yw^x-r_0^yw^x-q_0^xw^y+t(v^yw^x-v^xw^y)
+\end{align}
+The vector $u = p - p_0$. We obtain $p$ from $q(0)$, by finding $\lambda/2$.
+\begin{align}
+p = q(0) + \frac{\lambda}{2}\gamma
+\end{align}
+We get $p_0$ by finding the intersection between the two lines. If they are parallel then we find two points $p$ using $q(0)$ and $q(1)$.
+
+To find $a$,
+\begin{align} 
+p^x-\frac{a}{2} &= q^x \\
+p^y+\frac{a}{2} &= q^y
+\end{align}
+Substituting equation \eqref{eqn:q},
+\begin{align} 
+t = \frac{p^x-a/2-q_0^x}{v^x} \\
+t = \frac{p^y+a/2-q_0^y}{v^y}
+\end{align}
+which leads to
+\begin{align} 
+\frac{p^x-a/2-q_0^x}{v^x} &= \frac{p^y+a/2-q_0^y}{v^y}
+\end{align}
+Solving for $a$,
+\begin{align}
+\label{eqn:plain_a}
+a &= \frac{2(p^xv^y-q_0^xv^y-p^yv^x+q_0^yv^x)}{v^x+v^y}
+\end{align}
+$a$ is a parametric function in $s$, so using equations \eqref{eqn:plain_a} and \eqref{eqn:p},
+\begin{align}
+a(s) &= \frac{2((p_0^x+su^x)v^y-q_0^xv^y-(p_0^y+su^y)v^x+q_0^yv^x)}{v^x+v^y} \\
+     &= s\frac{2(u^xv^y-u^yv^x)}{v^x+v^y} + \frac{2(p_0^xv^y-q_0^xv^y-p_0^yv^x+q_0^yv^x)}{v^x+v^y} \label{eqn:a}
+\end{align}
+Figure \ref{fig:conflict-resolution-even} shows that we will find a series of points $P=p(s_i)$ such that
+\begin{equation}
+p^x(s_i) = p^x(s_{i-2})+a(s_{i-2}) \label{eqn:p_s}
+\end{equation}
+Using equations \eqref{eqn:p}, \eqref{eqn:a} and \eqref{eqn:p_s},
+\begin{align}
+p_0^x+s_iu^x = p_0^x+s_{i-2}u^x+s_{i-2}\frac{2(u^xv^y-u^yv^x)}{v^x+v^y} + \frac{2(p_0^xv^y-q_0^xv^y-p_0^yv^x+q_0^yv^x)}{v^x+v^y}
+\end{align}
+Solving for $s_i$,
+\begin{equation}
+\label{eqn:si_rec}
+s_i = s_{i-2}\alpha + \beta
+\end{equation}
+where
+\begin{align}
+%\alpha &= \frac{u^xv^x+3u^xv^y-2u^yv^x}{u^x(v^x+v^y)}
+\alpha &= 1+\frac{2(u^xv^y-u^yv^x)}{u^x(v^x+v^y)} \\
+\beta &= \frac{2(p_0^xv^y-q_0^xv^y-p_0^yv^x+q_0^yv^x)}{u^x(v^x+v^y)}
 \end{align}
 
-%$a$ is the largest power of two less than or equal to $a'$:
-%\begin{equation}
-%\label{eqn:ghi}
-%a(s) = a = 2^{\lfloor \log_2{a'} \rfloor}
-%\end{equation}
-
-As shown in figure \ref{fig:conflict-resolution}, we desire to build a sequence $S=(s_0, s_1, \dots)$ such that $p^x(s_i) = p^x(s_{i-2}) + a(s_{i-2})$.
-%For convenience, we split the sequence into even and odd components and define $p^x(s_i) = p^x(s_{i-2}) + a(s_{i-2})$.
-That is,
-\begin{align}
-x_i &= x_{i-2} + a(s_{i-2}) \label{eqn:xi_init}
-\end{align}
-Substituting in equation \eqref{eqn:a},
-\begin{equation}
-x_i = x_{i-2} + s_{i-2}\frac{u^yv^x-u^xv^y}{v^x+v^y} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{v^x+v^y}
-\end{equation}
-%\begin{align}
-%x_i &= x_{i-1} + \frac{1}{v^x+v^y}(q_0^xv^y-q_0^yv^x+p^y(s_{i-1})v^x-p^x(s_{i-1})v^y)
-%\end{align}
-%Substituting in equation \eqref{eqn:p} and rearranging,
-%\begin{align}
-%x_i &= x_{i-1} + \frac{1}{v^x+v^y}(q_0^xv^y-q_0^yv^x+(p_0^y+s_{i-1}u^y)v^x-(p_0^x+s_{i-1}u^x)v^y) \\
-%    &= x_{i-1} + \frac{1}{v^x+v^y}(q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y+s_{i-1}(u^yv^x-u^xv^y)) \\
-%    &= x_{i-1} + s_{i-1}\frac{u^yv^x-u^xv^y}{v^x+v^y} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{v^x+v^y}
-%\end{align}
-Substituting in equation \eqref{eqn:p},
-\begin{align}
-x_i &= p_0^x+s_{i-2}u^x + s_{i-2}\frac{u^yv^x-u^xv^y}{v^x+v^y} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{v^x+v^y} \\
-    &= s_{i-2}\frac{v^x(u^x+u^y)}{v^x+v^y} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x+p_0^xv^x}{v^x+v^y} \label{eqn:xi}
-\end{align}
-We're now ready to build the sequence. From \eqref{eqn:p},
-\begin{align}
-s_i &= \frac{x_i-p_0^x}{u^x}
-\end{align}
-Substituting in equation \eqref{eqn:xi},
-\begin{align}
-s_i &= s_{i-2}\frac{v^x(u^x+u^y)}{u^x(v^x+v^y)} + \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{u^x(v^x+v^y)} \\
-    &= \alpha s_{i-2} + \beta \label{eqn:si_rec}
-\end{align}
-where $\displaystyle \alpha=\frac{v^x(u^x+u^y)}{u^x(v^x+v^y)}$ and $\displaystyle \beta=\frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y}{u^x(v^x+v^y)}$.
-
+\red{TODO: this isn't right}
 To solve the recurrence relation \eqref{eqn:si_rec},
@@
 \end{equation}
-The initial condition when $i$ is even is $s_0=0$.
+The initial condition when $i$ is even is $s_0$, which we take to be the point at which the line $p$ enters the cell. Using
+\begin{align}
+s_2 &= \alpha s_0+\beta &= k_1+k_2\alpha  \label{eqn:s2} \\
+s_4 &= \alpha s_2+\beta \\
+    &= \alpha^2s_0+\alpha \beta + \beta &= k_1+k_2\alpha^2 \label{eqn:s4}
+\end{align}
+Solving \eqref{eqn:s2} and \eqref{eqn:s4} simultaneously,
+\begin{align}
+k_1^{even} &= \frac{\alpha s_0+\beta-s_0}{\alpha-1} \\
+k_2^{even} &= \frac{\alpha^2 s_0 + \alpha\beta-\alpha s_0}{\alpha-1}
+\end{align}
+The result is similar when $i$ is odd:
+\begin{align}
+k_1^{odd} &= \frac{\alpha s_0+\beta-s_0}{\alpha-1} \\
+k_2^{odd} &= \frac{\alpha^2 s_0 + \alpha\beta-\alpha s_0}{\alpha-1}
+\end{align}
+The last step to formulating $P$ for parallel computation is to determine how many samples we will need. Let $p(s_n)$ be the point at which the line $p$ exits the cell.
+\begin{align}
+k_1+k_2\alpha^i < s_n
+\end{align}
+results in
+\begin{align}
+i < \log_{\alpha}\frac{s_n-k_1}{k_2}
+\end{align}
+
+%
+%.
 When $i$ is odd, the initial condition will be where $s$ is halfway along the first box (it could be anywhere in the interior of the first box, but we use halfway for convenience).
 So from \eqref{eqn:p},
-\begin{align}
-s_1 &= \frac{p^x-p_0^x}{2u^x} \\
-    &= \frac{a(0)}{2u^x}
-\end{align}
-\begin{align}
-k_1^{even} = k_1^{odd} &= \beta\left(1-\frac{\alpha}{\alpha-1}\right) \\
-k_2^{even} &= \frac{\beta}{\alpha-1} \\
-k_2^{odd} &= \frac{a(0)}{2u^x}+\frac{\beta}{\alpha-1}
-\end{align}
-
-It is tempting to formulate the recurrance relation using a single initial condition. This would be done by replacing equation \eqref{eqn:xi_init} with
-\begin{align}
-x_i &= x_{i-1} + \frac{a(s_{i-1})}{2}
-\end{align}
-This formulation can lead to conflict cells as shown in figure \ref{fig:conflict-resolution-half}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.48\columnwidth]{conflict-resolution-half.pdf}
-  \caption{Using a single recurrence solution we can no longer guarantee that every box that intersects both lines contains at least two samples.
+%\begin{align}
+%s_1 &= \frac{p^x-p_0^x}{2u^x} \\
+%    &= \frac{a(0)}{2u^x}
+%\end{align}
+%\begin{align}
+%k_1^{even} = k_1^{odd} &= \beta\left(1-\frac{\alpha}{\alpha-1}\right) \\
+%k_2^{even} &= \frac{\beta}{\alpha-1} \\
+%k_2^{odd} &= \frac{a(0)}{2u^x}+\frac{\beta}{\alpha-1}
+%\end{align}
+%
+%It is tempting to formulate the recurrance relation using a single initial condition. This would be done by replacing equation \eqref{eqn:xi_init} with
+%\begin{align}
+%x_i &= x_{i-1} + \frac{a(s_{i-1})}{2}
+%\end{align}
+%This formulation can lead to conflict cells as shown in figure \ref{fig:conflict-resolution-half}.
+%
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=0.48\columnwidth]{conflict-resolution-half.pdf}
+%  \caption{Using a single recurrence solution we can no longer guarantee that every box that intersects both lines contains at least two samples.
 The box in red contains only one sample, so the octree may contain conflict cells.
-  }
-  \label{fig:conflict-resolution-half}
-\end{figure}
+%  }
+%  \label{fig:conflict-resolution-half}
+%\end{figure}
 

### 2016-09-14 paper.tex
Updating conflict resolution

@@
 \title[PGVD]%
-      {Parallel Octree Construction on Collections of Objects}
+      {Parallel Quadtree Construction on Collections of Objects}
 
@@
 \begin{abstract}
-We present a parallel octree algorithm that resolves between geometric objects.
-The octree has the property that no octree cell intersects more than one labeled object.
+We present a parallel quadtree algorithm that resolves between geometric objects.
+The quadtree has the property that no quadtree cell intersects more than one labeled object.
 Previous parallel algorithms either spawn kernels hierarchically, separate points only, or make no hard guarantees of object separation.
@@
 We present an algorithm to compute a GVD approximation on arbitrary datasets, including those with closely spaced objects.
-The approach applies a distance transform over an octree representation of the objects.
-Our octree, its associated data structure, and our distance transform are novel and optimized to GVD approximation.
+The approach applies a distance transform over an quadtree representation of the objects.
+Our quadtree, its associated data structure, and our distance transform are novel and optimized to GVD approximation.
 For the remainder of the paper, ``GVD'' will refer to the approximated Generalized Voronoi Diagram.
@@
 \begin{enumerate}
-\item Construct an octree on object vertices using Karras' algorithm \cite{karras2012maximizing}
-\item Detect octree cells that intersect more than one object, which we call ``conflict cells'' (contribution)
+\item Construct an quadtree on object vertices using Karras' algorithm \cite{karras2012maximizing}
+\item Detect quadtree cells that intersect more than one object, which we call ``conflict cells'' (contribution)
 \item Subdivide conflict cells to resolve objects (contribution)
@@
 Ordinary Voronoi diagrams are well studied and efficient algorithms exist that compute them exactly \cite{de2008computational}, but exact algorithms for the generalized Voronoi diagram are limited to a small set of special cases \cite{lee1982medial,karavelas2004robust}.
-In an early work, Lavender \etal \shortcite{lavender1992voronoi} define and compute GVDs over a set of solid models using an octree.
+In an early work, Lavender \etal \shortcite{lavender1992voronoi} define and compute GVDs over a set of solid models using an quadtree.
 Etzion and Rappoport \shortcite{etzion2002computing} represent the GVD bisector symbolically for lazy evaluation, but are limited to sites that are polyhedra.
@@
 
-\paragraph{Distance fields and octrees}
+\paragraph{Distance fields and quadtrees}
 The GVD is a subset of the locus of distance field critical points, a property that we take advantage of.
@@
 % JME: note that in the Bastos work, ``reconstruction'' is reconstruction
-% of the surface from the octree at rendering time.
-Two seminal works adaptively compute the Adaptive Distance Field (ADF) on octree vertices.
-Strain~\shortcite{strain1999fast} fully resolves the octree everywhere on the object surface, and Frisken \etal~\shortcite{frisken2000adaptively} resolve the octree fully only in areas of small local feature size.
+% of the surface from the quadtree at rendering time.
+Two seminal works adaptively compute the Adaptive Distance Field (ADF) on quadtree vertices.
+Strain~\shortcite{strain1999fast} fully resolves the quadtree everywhere on the object surface, and Frisken \etal~\shortcite{frisken2000adaptively} resolve the quadtree fully only in areas of small local feature size.
 Both approaches are designed to retain features of a single object rather than resolving between multiple objects, as is required for GVD computation.
 Qu \etal \shortcite{qu2004feature} implement an energy-minimizing distance field algorithm that preserves features at the expense of efficiency.
-Many recent works on fast octree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}.
-Most octree approaches that support surfaces \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed} are designed for efficient rendering, and actual construction of the octree is implemented on the CPU.
+Many recent works on fast quadtree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}.
+Most quadtree approaches that support surfaces \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed} are designed for efficient rendering, and actual construction of the quadtree is implemented on the CPU.
 
-Two works \cite{bastos2008gpu,park2010cuda} implement the ADF using GPU parallelism to compute the distance value at sample points, but building the octree itself is done sequentially.
-Yin \etal~\shortcite{yin2011fast} compute the distance field entirely on the GPU using a bottom-up approach by initially subdividing into a complete octree, resulting in memory usage that is no better than using a uniform grid.
-A method by Kim and Liu~\shortcite{kim2014exact} computes the octree and a BVH entirely on the GPU.
-However, octree construction is performed on barycenters of triangles, and so a leaf octree cell can have an arbitrary number of triangle intersections as long as it contains no more than one triangle's barycenter.
-We have found no GPU octree construction method that can resolve between objects.
+Two works \cite{bastos2008gpu,park2010cuda} implement the ADF using GPU parallelism to compute the distance value at sample points, but building the quadtree itself is done sequentially.
+Yin \etal~\shortcite{yin2011fast} compute the distance field entirely on the GPU using a bottom-up approach by initially subdividing into a complete quadtree, resulting in memory usage that is no better than using a uniform grid.
+A method by Kim and Liu~\shortcite{kim2014exact} computes the quadtree and a BVH entirely on the GPU.
+However, quadtree construction is performed on barycenters of triangles, and so a leaf quadtree cell can have an arbitrary number of triangle intersections as long as it contains no more than one triangle's barycenter.
+We have found no GPU quadtree construction method that can resolve between objects.
 
 %-----------------------------------------------------------
-% Build octree
+% Build quadtree
 %-----------------------------------------------------------
@@
 \label{sec:algorithm}
-Our algorithm works in both 2D and 3D.
-Lacking a dimension-independent term, we use ``octree'' as a general term to refer to both quadtrees and octrees.
-We refer to object facets, which are lines in 2D and triangles in 3D.
 
@@
     \label{fig:conflict-resolution-octree}
-    \includegraphics[width=0.48\columnwidth]{conflict-resolution-octree-samples.pdf} }
+    \includegraphics[width=0.48\columnwidth]{conflict-resolution-octree.pdf} }
   \caption{
@@
     \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least two samples.
-This ensures that an octree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-octree}) will have no leaf cells that intersect both lines, ensuring that the new octree is locally free of conflict cells.
+This ensures that an quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-octree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
   }
@@
 
-To resolve conflict cells we consider pairs of lines of differing labels that intersect a cell $c$. Figure \ref{fig:conflict-resolution-x} shows two lines
+A conflict cell is a quadtree cell that intersects at least two different objects.
+To resolve a conflict cell $c$, we consider pairs of lines of differing labels that intersect $c$.
+Figure \ref{fig:conflict-resolution-x} shows two lines
 
@@
 \end{align}
-Our strategy will be to sample points $P$ on $p(u)$ (figure \ref{fig:conflict-resolution-octree}) such that an octree built on $V \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
+that bisects $q$ and $r$.
+Our strategy will be to sample points $P$ on $p(u)$ (figure \ref{fig:conflict-resolution-octree}) such that an quadtree built on $V \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
 We do this by ensuring that $P$ is sampled such that every box that intersects both $q$ and $r$ also intersects at least two points in $P$.
@@
 We begin by fitting the smallest box centered on a point $p$ that intersects both $q$ and $r$.
+We break the problem into two cases: the \textit{opposite} case is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
+The \textit{adjacent} case  is where $w^y < 0$, so the line intersections are adjacent -- the top-left and bottom-left corners of the box.
+Figure \ref{fig:conflict-resolution-octree} shows the \textit{opposite} case.
 
-\subsubsection{Case $w^y>0$}
+\subsubsection{\textit{Opposite} case}
 
@@
 %  \caption{Using a single recurrence solution we can no longer guarantee that every box that intersects both lines contains at least two samples.
-The box in red contains only one sample, so the octree may contain conflict cells.
+The box in red contains only one sample, so the quadtree may contain conflict cells.
 %  }
@@
 
-\subsection{Build octree on vertices}
+\subsection{Build quadtree on vertices}
 
-We first construct an octree on the vertices of the objects, which we call the ``vertex octree''.
+We first construct an quadtree on the vertices of the objects, which we call the ``vertex quadtree''.
 We use Karras' algorithm \cite{karras2012maximizing} which sorts the Morton codes of the vertices in parallel, then constructs the binary radix tree in parallel.
-With the binary radix tree, the octree can be constructed with a single parallel call.
+With the binary radix tree, the quadtree can be constructed with a single parallel call.
 The strength of this algorithm lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points.
-That is, even if a large number of vertices are clustered in a small area, requiring deep octree subdivision, only a constant number of parallel calls need be made.
+That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.
 Given enough parallel units, the Karras algorithm runs in $O(\log{N})$ time, where $N$ is the number of vertices.
@@
 
-Our end goal is to construct an octree such that no octree cell intersects more than one object.
+Our end goal is to construct an quadtree such that no quadtree cell intersects more than one object.
 Note that a cell is allowed to intersect more than one facet, but all facets must belong to the same object, or, in other words, all facets must share the same label.
-It is possible, but unlikely, that the vertex octree has this property.
+It is possible, but unlikely, that the vertex quadtree has this property.
 If so, then we are done.
@@
 We identify conflict cells as shown in algorithm \ref{alg:find-conflict-cells}.
-In lines \ref{alg:octree_containing_begin}-\ref{alg:octree_containing_end}, for each internal octree cell $c$, we store all facets for which $c$ is the smallest containing cell.
+In lines \ref{alg:quadtree_containing_begin}-\ref{alg:quadtree_containing_end}, for each internal quadtree cell $c$, we store all facets for which $c$ is the smallest containing cell.
 Since we are implementing this in a GPGPU environment, we don't have dynamic memory, so each facet must be processed twice.
@@
 
-The $\container{f}$ procedure finds the smallest octree cell that fully contains the facet $f$.
-A straightforward implementation of $\container{f}$ is to perform a standard octree search on the vertices of $f$ and take the smallest octree cell that contains all of them.
+The $\container{f}$ procedure finds the smallest quadtree cell that fully contains the facet $f$.
+A straightforward implementation of $\container{f}$ is to perform a standard quadtree search on the vertices of $f$ and take the smallest quadtree cell that contains all of them.
 (Note that the cell is always an internal node, since a post-condition of the Karras algorithm is that no leaf cell contains more than one vertex.) In our implementation however, we take advantage of our existing data structures.
-The octree cell that contains a vertex $v$ is uniquely determined by the D-tuple bits of its morton code.
-For example, if a 2D vertex has morton code $010010$, then the octree is traversed from the root to child $01$ to child $00$ to child $10$.
+The quadtree cell that contains a vertex $v$ is uniquely determined by the D-tuple bits of its morton code.
+For example, if a 2D vertex has morton code $010010$, then the quadtree is traversed from the root to child $01$ to child $00$ to child $10$.
 To determine $\container{f}$, we find the longest common prefix (\lcp) of the vertices.
-Truncating the length of \lcp\,to a multiple of D, we find the smallest octree cell that contains all vertices of $f$.
+Truncating the length of \lcp\,to a multiple of D, we find the smallest quadtree cell that contains all vertices of $f$.
 The complexity of $\container{f}$ is $O(\log{N})$ for both implementations.
-Thus, lines \ref{alg:octree_containing_begin}-\ref{alg:octree_containing_end} run in $O(\log{N})$ time.
+Thus, lines \ref{alg:quadtree_containing_begin}-\ref{alg:quadtree_containing_end} run in $O(\log{N})$ time.
 
-Lines \ref{alg:octree_intersections_begin}-\ref{alg:octree_intersections_end} of the algorithm identify and store all facets that intersect with a given leaf cell $c$.
+Lines \ref{alg:quadtree_intersections_begin}-\ref{alg:quadtree_intersections_end} of the algorithm identify and store all facets that intersect with a given leaf cell $c$.
 Again, it is done in two steps for memory allocation purposes.
@@
 
-In Step 4, Stack is preallocated to size $M\cdot2^D$ where $M$ is the maximum octree depth and $D$ is the dimension.
+In Step 4, Stack is preallocated to size $M\cdot2^D$ where $M$ is the maximum quadtree depth and $D$ is the dimension.
 A conflict cell is a cell that intersects at least two different objects, or two lines of different labels.
 
-The second procedure we use is $\directAncestors{c}$, which finds all ancestors of octree cell $c$.
+The second procedure we use is $\directAncestors{c}$, which finds all ancestors of quadtree cell $c$.
 
@@
   \LinesNumbered
-  \KwIn{VertexOctree}
+  \KwIn{VertexQuadtree}
   \BlankLine
@@
   \ForPar{facet $f$ in Objects}{
-    \label{alg:octree_containing_begin}
+    \label{alg:quadtree_containing_begin}
     a := $\container{f}$\;
@@
     a.facets := a.facets $\cup$ f
-  } \label{alg:octree_containing_end}
+  } \label{alg:quadtree_containing_end}
   \tcp{Store intersecting facets}
-  \ForPar{leaf cell c in VertexOctree}{
-    \label{alg:octree_intersections_begin}
+  \ForPar{leaf cell c in VertexQuadtree}{
+    \label{alg:quadtree_intersections_begin}
     \ForEach{cell a in \directAncestors{c}}{
@@
   Allocate space for facets in leaf cells\;
-  \ForPar{leaf cell c in VertexOctree}{
+  \ForPar{leaf cell c in VertexQuadtree}{
     \ForEach{cell a in \directAncestors{c}}{
@@
     }
-  } \label{alg:octree_intersections_end}
+  } \label{alg:quadtree_intersections_end}
 \caption{FIND\_CONFLICT\_CELLS}
@@
   \LinesNumbered
-  \KwIn{Octree, conflict\_cells}
+  \KwIn{Quadtree, conflict\_cells}
   \BlankLine
-  \tcp{4. Octree refinement}
-  \ForPar{leaf cell c in Octree}{
+  \tcp{4. Quadtree refinement}
+  \ForPar{leaf cell c in Quadtree}{
     c' := c\;
@@
   }
-\caption{REFINE\_OCTREE}
-\label{alg:refine-octree}
+\caption{REFINE\_QUADTREE}
+\label{alg:refine-quadtree}
 \end{algorithm}
@@
     \protect\subref{fig:octree-cartesian-1} A red object and a blue object.
-    \protect\subref{fig:octree-cartesian-2} The vertex octree, or octree built on the object vertices using Karras' algorithm.
-    \protect\subref{fig:octree-cartesian-3}, \protect\subref{fig:octree-hierarchical-1} The vertex octree with conflict cells highlighted.
-Note the label of an octree cell in \protect\subref{fig:octree-cartesian-3} is the concatenation of labels from root R to the leaf cell in \protect\subref{fig:octree-hierarchical-1}.
+    \protect\subref{fig:octree-cartesian-2} The vertex quadtree, or quadtree built on the object vertices using Karras' algorithm.
+    \protect\subref{fig:octree-cartesian-3}, \protect\subref{fig:octree-hierarchical-1} The vertex quadtree with conflict cells highlighted.
+Note the label of an quadtree cell in \protect\subref{fig:octree-cartesian-3} is the concatenation of labels from root R to the leaf cell in \protect\subref{fig:octree-hierarchical-1}.
 This value also corresponds to the highest order bits of the morton code of any point in the cell.
-    \protect\subref{fig:octree-cartesian-5}, \protect\subref{fig:octree-hierarchical-2} The octree after resolution of conflict cells.
+    \protect\subref{fig:octree-cartesian-5}, \protect\subref{fig:octree-hierarchical-2} The quadtree after resolution of conflict cells.
   }
@@
 
-% # octree vertices
+% # quadtree vertices
 %  52K ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
@@
 
-% memory (54 bytes per octree cell)
+% memory (54 bytes per quadtree cell)
 %  2.8Mb ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
@@
     \toprule
-    dataset & objects & object          & octree   & octree          & octree &
+    dataset & objects & object          & quadtree   & quadtree          & quadtree &
     GVD   & GVD             \\
@@
   \end{tabular}}
-  \caption{Table of octree/GVD computation statistics and timings on datasets that are unmanageable using other methods.
-\red{Columns are: \emph{objects} - the number of objects in the dataset; \emph{object $\Delta$s} - the number of line segments (2D) or triangles (3D) of all objects in the dataset; \emph{octree depth} - required octree depth in order to resolve objects; \emph{octree cells} - total number of leaf octree cells; \emph{octree memory} - amount of memory used by the octree; \emph{GVD (sec)} - seconds to perform all steps of GVD computation; \emph{GVD $\Delta$s} - number of line segments (2D) or triangles (3D) in the GVD.}}
+  \caption{Table of quadtree/GVD computation statistics and timings on datasets that are unmanageable using other methods.
+\red{Columns are: \emph{objects} - the number of objects in the dataset; \emph{object $\Delta$s} - the number of line segments (2D) or triangles (3D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{quadtree cells} - total number of leaf quadtree cells; \emph{quadtree memory} - amount of memory used by the quadtree; \emph{GVD (sec)} - seconds to perform all steps of GVD computation; \emph{GVD $\Delta$s} - number of line segments (2D) or triangles (3D) in the GVD.}}
   \label{tab:timings}

### 2016-09-14 paper.tex
Added new version of conflict resolution equations

@@
     \label{fig:conflict-resolution-octree}
-    \includegraphics[width=0.48\columnwidth]{conflict-resolution-octree.pdf} }
+    \includegraphics[width=0.48\columnwidth]{conflict-resolution-octree.pdf} } \\
+  \subfloat[][]{
+    \label{fig:conflict-resolution-adjacent-even}
+    \includegraphics[width=0.48\columnwidth]{conflict-resolution-adjacent-even.pdf} }
+  \subfloat[][]{
+    \label{fig:conflict-resolution-adjacent-all}
+    \includegraphics[width=0.48\columnwidth]{conflict-resolution-adjacent-all.pdf} }
   \caption{
@@
 that bisects $q$ and $r$.
-Our strategy will be to sample points $P$ on $p(u)$ (figure \ref{fig:conflict-resolution-octree}) such that an quadtree built on $V \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
+Our strategy will be to sample points $P$ on $p(s)$ (figure \ref{fig:conflict-resolution-octree}) such that an quadtree built on $V \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
 We do this by ensuring that $P$ is sampled such that every box that intersects both $q$ and $r$ also intersects at least two points in $P$.
 Because Karras' algorithm guarantees that every leaf cell intersects at most one point, we know that no leaf cell will intersect $q$ and $r$ and thus no leaf cell will be a conflict cell.
+We will find a series of boxes such that each box's left-most intersection with $p(s)$ is a sample point meeting the above criterion.
 
@@
 We begin by fitting the smallest box centered on a point $p$ that intersects both $q$ and $r$.
-We break the problem into two cases: the \textit{opposite} case is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
-The \textit{adjacent} case  is where $w^y < 0$, so the line intersections are adjacent -- the top-left and bottom-left corners of the box.
-Figure \ref{fig:conflict-resolution-octree} shows the \textit{opposite} case.
+We break the problem into two cases: the \textit{opposite} case (see Figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
+The \textit{adjacent} case (see Figure \ref{fig:conflict-resolution-adjacent-even}) is where $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
 
-\subsubsection{\textit{Opposite} case}
+\subsubsection{Finding $a(s)$ -- \textit{opposite} case}
 
-Consider a box with edge length $a_0$ centered on a point $p_0 \in p$.
+Given a point $p(s)$, we wish to find $a=a(s)$, which will give us the starting $x$ coordinate for the next box.
+Consider the top-left corner of the box $q(t(s))=q(t)$ and the bottom-right corner $r(f(s))=r(f)$.
+
+Because $p^x(s)=q^x(t)$,
+\begin{equation}
+t = \frac{p^x(s)-q_0^x}{v^x} = \frac{p_x^x-q_0^x+su^x}{v^x} \label{eqn:t}
+\end{equation}
+Because our boxes are square,
+\begin{equation}
+r(f) = r_0+fw = q_0+tv+a\begin{bmatrix}1\\-1\end{bmatrix} \label{eqn:ro}
+\end{equation}
+From \eqref{eqn:ro},
+\begin{align}
+f &= \frac{1}{w^y}(q_0^y+tv^y-a-r_0^y) \label{eqn:f} \\
+a &= r_0^x+fw^x-q_0^x-tv^x \label{eqn:a1o}
+\end{align}
+Substituting equations \eqref{eqn:t} and \eqref{eqn:f} into equation \eqref{eqn:a1o} and solving for $a$,
+\begin{equation}
+a(s) = \hat{\alpha}_o s + \hat{\beta}_o \label{eqn:ao}
+\end{equation}
+where
+\begin{equation}
+\hat{\alpha}_o = \frac{u^x|w \times v|}{v^x(w^x+w^y)}
+\end{equation}
+and
+\begin{equation}
+\hat{\beta}_o = \frac{|w \times v|(p_0^x-q_0^x) + v^x(|r_0 \times w| + |w \times q_0|)}{v^x(w^x+w^y)}
+\end{equation}
+
+\subsubsection{Finding $a(s)$ -- \textit{adjacent} case}
+
+Consider the top-left corner of the box $q(t(s))=q(t)$ and the bottom-left corner $r(f(s))=r(f)$. $r(f)$ is now defined as
+\begin{equation}
+r(f) = r_0+fw = q_0+tv+a\begin{bmatrix}0\\-1\end{bmatrix} \label{eqn:ra}
+\end{equation}
+Equations \eqref{eqn:t} and \eqref{eqn:f} remain the same while \eqref{eqn:a1o} becomes
+\begin{equation}
+0 = r_0^x+fw^x-q_0^x-tv^x \label{eqn:a1a}
+\end{equation}
+Substituting equations \eqref{eqn:t} and \eqref{eqn:f} into equation \eqref{eqn:a1a} and solving for $a$,
+\begin{equation}
+a(s) = \hat{\alpha}_a s + \hat{\beta}_a \label{eqn:aa}
+\end{equation}
+where
+\begin{equation}
+\hat{\alpha}_a = \frac{u^x}{v^xw^x}
+\end{equation}
+and
+\begin{equation}
+\hat{\beta}_a = \frac{w^x(p_0^x-q_0^x)+|w \times q_0| + |r_0 \times w|}{w^x}
+\end{equation}
+
+%
+%
+%Because $p^x(s)=q^x(t)$,
+%\begin{equation}
+%t = \frac{p^x(s)-q_0^x}{v^x} = \frac{p_x^x-q_0^x+su^x}{v^x} \label{eqn:t}
+%\end{equation}
+%Because our boxes are square,
+%\begin{equation}
+%r(f) = r_0+fw = q_0+tv+a
+%\begin{bmatrix}
+%1\\-1
+%\end{bmatrix}
+%\label{eqn:r}
+%\end{equation}
+%From \eqref{eqn:r},
+%\begin{align}
+%f &= \frac{1}{w^y}(q_0^y+tv^y-a-r_0^y) \label{eqn:f} \\
+%a &= r_0^x+fw^x-q_0^x-tv^x \label{eqn:a1}
+%\end{align}
+%Substituting equations \eqref{eqn:t} and \eqref{eqn:f} into equation \eqref{eqn:a1} and solving for $a$,
+%\begin{equation}
+%a = \alpha s + \beta
+%\end{equation}
+%where
+%\begin{equation}
+%\alpha = \frac{u^x|w \times v|}{v^x(w^x+w^y)}
+%\end{equation}
+%and
+%\begin{equation}
+%\beta = \frac{|w \times v|(p_0^x-q_0^x) + v^x(|r_0 \times w| + |w \times q_0|)}{v^x(w^x+w^y)}
+%\end{equation}
+%
+%Consider a box with edge length $a_0$ centered on a point $p_0 \in p$.
 We first find $p$, which is the line that is equidistant from $q$ and $r$ along the diagonal vector $\gamma = (1, -1)$.
 Given some point $q(t)$,
+%\begin{equation}
+%q_0+tv + \lambda\gamma = r_0+fw \\
+%\end{equation}
+%Solving for $\lambda$,
+%\begin{align}
+%f &= \frac{1}{w^y}(q_0^y+tv^x-\lambda-r_0^y) \\
+%\lambda &= r_0^x + fw^x-q_0^x-tv^x \\
+%        &= r_0^x+\frac{w^x}{w^y}(q_0^y+tv^y-\lambda-r_0^y) - q_0^x-tv^x \\
+%        &= \frac{1}{w^x+w^y}(r_0^xw^y+q_0^yw^x-r_0^yw^x-q_0^xw^y+t(v^yw^x-v^xw^y)
+%\end{align}
+%The vector $u = p - p_0$. We obtain $p$ from $q(0)$, by finding $\lambda/2$.
+%\begin{align}
+%p = q(0) + \frac{\lambda}{2}\gamma
+%\end{align}
+%We get $p_0$ by finding the intersection between the two lines. If they are parallel then we find two points $p$ using $q(0)$ and $q(1)$.
+%
+%To find $a$,
+%\begin{align} 
+%p^x-\frac{a}{2} &= q^x \\
+%p^y+\frac{a}{2} &= q^y
+%\end{align}
+%Substituting equation \eqref{eqn:q},
+%\begin{align} 
+%t = \frac{p^x-a/2-q_0^x}{v^x} \\
+%t = \frac{p^y+a/2-q_0^y}{v^y}
+%\end{align}
+%which leads to
+%\begin{align} 
+%\frac{p^x-a/2-q_0^x}{v^x} &= \frac{p^y+a/2-q_0^y}{v^y}
+%\end{align}
+%Solving for $a$,
+%\begin{align}
+%\label{eqn:plain_a}
+%a &= \frac{2(p^xv^y-q_0^xv^y-p^yv^x+q_0^yv^x)}{v^x+v^y}
+%\end{align}
+%$a$ is a parametric function in $s$, so using equations \eqref{eqn:plain_a} and \eqref{eqn:p},
+%\begin{align}
+%a(s) &= \frac{2((p_0^x+su^x)v^y-q_0^xv^y-(p_0^y+su^y)v^x+q_0^yv^x)}{v^x+v^y} \\
+%     &= s\frac{2(u^xv^y-u^yv^x)}{v^x+v^y} + \frac{2(p_0^xv^y-q_0^xv^y-p_0^yv^x+q_0^yv^x)}{v^x+v^y} \label{eqn:a}
+%\end{align}
+%Figure \ref{fig:conflict-resolution-even} shows that we will find a series of points $P=p(s_i)$ such that
+%\begin{equation}
+%p^x(s_i) = p^x(s_{i-2})+a(s_{i-2}) \label{eqn:p_s}
+%\end{equation}
+%Using equations \eqref{eqn:p}, \eqref{eqn:a} and \eqref{eqn:p_s},
+%\begin{align}
+%p_0^x+s_iu^x = p_0^x+s_{i-2}u^x+s_{i-2}\frac{2(u^xv^y-u^yv^x)}{v^x+v^y} + \frac{2(p_0^xv^y-q_0^xv^y-p_0^yv^x+q_0^yv^x)}{v^x+v^y}
+%\end{align}
+%Solving for $s_i$,
+%\begin{equation}
+%\label{eqn:si_rec}
+%s_i = s_{i-2}\alpha + \beta
+%\end{equation}
+%where
+%\begin{align}
+%%\alpha &= \frac{u^xv^x+3u^xv^y-2u^yv^x}{u^x(v^x+v^y)}
+%\alpha &= 1+\frac{2(u^xv^y-u^yv^x)}{u^x(v^x+v^y)} \\
+%\beta &= \frac{2(p_0^xv^y-q_0^xv^y-p_0^yv^x+q_0^yv^x)}{u^x(v^x+v^y)}
+%\end{align}
+
+\subsubsection{Sampling}
+In both the \textit{opposite} and the \textit{adjacent} cases, $a(s)$ is of the form $a(s) = \hat{\alpha} s + \hat{\beta}$.
+We now use $a(s)$ to construct a sequence of $s$ values $S = \{s_0, s_1, s_2, \dots, s_n\}$ that meet our sampling criterion.
+We first construct the even samples (see Figures \ref{fig:conflict-resolution-even} and \ref{fig:conflict-resolution-adjacent-even}).
+Given a starting point $p(s_0)$,
 \begin{equation}
-q_0+tv + \lambda\gamma = r_0+fw \\
+p^x(s_{i+2}) = p^x(s_i) + a(s_i)
 \end{equation}
-Solving for $\lambda$,
-\begin{align}
-f &= \frac{1}{w^y}(q_0^y+tv^x-\lambda-r_0^y) \\
-\lambda &= r_0^x + fw^x-q_0^x-tv^x \\
-        &= r_0^x+\frac{w^x}{w^y}(q_0^y+tv^y-\lambda-r_0^y) - q_0^x-tv^x \\
-        &= \frac{1}{w^x+w^y}(r_0^xw^y+q_0^yw^x-r_0^yw^x-q_0^xw^y+t(v^yw^x-v^xw^y)
-\end{align}
-The vector $u = p - p_0$. We obtain $p$ from $q(0)$, by finding $\lambda/2$.
-\begin{align}
-p = q(0) + \frac{\lambda}{2}\gamma
-\end{align}
-We get $p_0$ by finding the intersection between the two lines. If they are parallel then we find two points $p$ using $q(0)$ and $q(1)$.
+Substituting in equations \eqref{eqn:p} and \eqref{eqn:ao}/\eqref{eqn:aa},
+\begin{equation}
+p_0^x + s_{i+2}u^x = p_0^x + s_i + \hat{\alpha} s_i + \hat{\beta}
+\end{equation}
+Solving for $s_{i+2}$ gives the recurrence relation
+\begin{equation}
+s_{i+2} = \alpha s_i + \beta \label{eqn:recurrence}
+\end{equation}
+where
+\begin{equation}
+\alpha = 1 + \frac{\hat{\alpha}}{u^x}
+\end{equation}
+and
+\begin{equation}
+\beta = \frac{\hat{\beta}}{u^x}
+\end{equation}
 
-To find $a$,
-\begin{align} 
-p^x-\frac{a}{2} &= q^x \\
-p^y+\frac{a}{2} &= q^y
-\end{align}
-Substituting equation \eqref{eqn:q},
-\begin{align} 
-t = \frac{p^x-a/2-q_0^x}{v^x} \\
-t = \frac{p^y+a/2-q_0^y}{v^y}
-\end{align}
-which leads to
-\begin{align} 
-\frac{p^x-a/2-q_0^x}{v^x} &= \frac{p^y+a/2-q_0^y}{v^y}
-\end{align}
-Solving for $a$,
-\begin{align}
-\label{eqn:plain_a}
-a &= \frac{2(p^xv^y-q_0^xv^y-p^yv^x+q_0^yv^x)}{v^x+v^y}
-\end{align}
-$a$ is a parametric function in $s$, so using equations \eqref{eqn:plain_a} and \eqref{eqn:p},
-\begin{align}
-a(s) &= \frac{2((p_0^x+su^x)v^y-q_0^xv^y-(p_0^y+su^y)v^x+q_0^yv^x)}{v^x+v^y} \\
-     &= s\frac{2(u^xv^y-u^yv^x)}{v^x+v^y} + \frac{2(p_0^xv^y-q_0^xv^y-p_0^yv^x+q_0^yv^x)}{v^x+v^y} \label{eqn:a}
-\end{align}
-Figure \ref{fig:conflict-resolution-even} shows that we will find a series of points $P=p(s_i)$ such that
+Constructing the odd samples is identical, except that we start at
 \begin{equation}
-p^x(s_i) = p^x(s_{i-2})+a(s_{i-2}) \label{eqn:p_s}
+s_1 = \left(1+\frac{\hat{\alpha}}{2u^x}\right)s_0 + \frac{\hat{\beta}}{2}
 \end{equation}
-Using equations \eqref{eqn:p}, \eqref{eqn:a} and \eqref{eqn:p_s},
-\begin{align}
-p_0^x+s_iu^x = p_0^x+s_{i-2}u^x+s_{i-2}\frac{2(u^xv^y-u^yv^x)}{v^x+v^y} + \frac{2(p_0^xv^y-q_0^xv^y-p_0^yv^x+q_0^yv^x)}{v^x+v^y}
-\end{align}
-Solving for $s_i$,
+which is the point in the center of the first box in the x-dimension.
+
+We solve the recurrence relation \eqref{eqn:recurrence} using the characteristic polynomial to yield
 \begin{equation}
-\label{eqn:si_rec}
-s_i = s_{i-2}\alpha + \beta
+s_i = k_1 + k_2 \alpha^i
 \end{equation}
@@
 \begin{align}
-%\alpha &= \frac{u^xv^x+3u^xv^y-2u^yv^x}{u^x(v^x+v^y)}
-\alpha &= 1+\frac{2(u^xv^y-u^yv^x)}{u^x(v^x+v^y)} \\
-\beta &= \frac{2(p_0^xv^y-q_0^xv^y-p_0^yv^x+q_0^yv^x)}{u^x(v^x+v^y)}
+k_1^{even} &= \frac{\beta}{1-\alpha} \\
+k_1^{odd} &= \frac{\beta}{1-\alpha} \\
+k_2^{even} &= \frac{\alpha s_0 + \beta - s_0}{\alpha-1} \\
+k_2^{odd} &= \frac{\alpha s_1 + \beta - s_1}{\alpha-1}
 \end{align}
-
-\red{TODO: this isn't right}
-To solve the recurrence relation \eqref{eqn:si_rec},
+The last step to formulating $P$ for parallel computation is to determine how many samples we will need. Let $p(s_{exit})$ be the point at which the line $p$ exits the cell.
 \begin{align}
-s_{i+2} &= \alpha s_i+\beta \\
-s_{i+2}-s_i &= \alpha s_i-\alpha s_{i-2} \\
-s_{i+2} &= (\alpha+1) s_i-\alpha s_{i-2}
-\end{align}
-The characteristic polynomial is
-\begin{equation}
-p(t) = t^2-(\alpha+1)t + \alpha
-\end{equation}
-which has roots $\alpha$ and $1$. Thus,
-\begin{equation}
-s_i = k_1 + k_2 \alpha^i
-\end{equation}
-The initial condition when $i$ is even is $s_0$, which we take to be the point at which the line $p$ enters the cell. Using
-\begin{align}
-s_2 &= \alpha s_0+\beta &= k_1+k_2\alpha  \label{eqn:s2} \\
-s_4 &= \alpha s_2+\beta \\
-    &= \alpha^2s_0+\alpha \beta + \beta &= k_1+k_2\alpha^2 \label{eqn:s4}
-\end{align}
-Solving \eqref{eqn:s2} and \eqref{eqn:s4} simultaneously,
-\begin{align}
-k_1^{even} &= \frac{\alpha s_0+\beta-s_0}{\alpha-1} \\
-k_2^{even} &= \frac{\alpha^2 s_0 + \alpha\beta-\alpha s_0}{\alpha-1}
-\end{align}
-The result is similar when $i$ is odd:
-\begin{align}
-k_1^{odd} &= \frac{\alpha s_0+\beta-s_0}{\alpha-1} \\
-k_2^{odd} &= \frac{\alpha^2 s_0 + \alpha\beta-\alpha s_0}{\alpha-1}
-\end{align}
-The last step to formulating $P$ for parallel computation is to determine how many samples we will need. Let $p(s_n)$ be the point at which the line $p$ exits the cell.
-\begin{align}
-k_1+k_2\alpha^i < s_n
+k_1+k_2\alpha^i < s_{exit}
 \end{align}
@@
 \begin{align}
-i < \log_{\alpha}\frac{s_n-k_1}{k_2}
+i < \log_{\alpha}\frac{s_{exit}-k_1}{k_2}
 \end{align}

### 2016-09-21 paper.tex
added Nathan's algorithm

@@
 \label{sec:intro}
-Previous paper: \cite{edwards2015approximating}
+Constructing quadtrees on objects in an important task with applications to collision detection, distance fields, robot navigation, object description, and other applications.
+Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects.
+Our work however, centers on using quadtrees to separate, or resolve, collections of closely spaced objects.
+Using a quadtree, we can model the space between objects, the first step in constructing distance fields, detecting collisions, and computing the generalized Voronoi diagram.
+Modeling inter-object spacing is computationally straightforward when inter-object spacing is large compared to the world bounding box.
+Approaches typically involve a uniform grid of the space, which leads to efficient computation using, very often, graphics processors.
 
-The generalized Voronoi diagram (GVD) is an important structure that divides space into a complex of generalized Voronoi cells (GVCs) around objects.
-Similar to the ordinary Voronoi diagram, each GVC contains exactly one object, or site, and every point in the GVC is closer to its contained object than to any other object.
-The generalized Voronoi diagram is the boundary of the cell complex, and thus every point on the GVD is equidistant from two or more closest objects.
-Applications of the GVD range from motion path planning to GIS analysis to mosaicking.
+\gray{Whereas most algorithms are efficient and robust on certain datasets, all algorithms but one \cite{edwards2015approximating} require inordinate amounts of memory on datasets where objects are very closely spaced relative to the size of the domain.
+The failures occur because the space is uniformly gridded.
+In such approaches, voxel size must be small enough to resolve object spacings, and if two objects are very close to each other the number of voxels can become prohibitively large.}
 
-Ordinary Voronoi diagrams have been studied extensively and efficient algorithms exist to compute them, but the GVD is difficult to compute analytically in general \cite{boissonnat2006curved,hoff1999fast} and so the majority of approaches compute an approximation.
-Whereas most algorithms are efficient and robust on certain datasets, all algorithms to our knowledge require inordinate amounts of memory on datasets where objects are very closely spaced relative to the size of the domain.
-The failures occur because the space is uniformly gridded.
-In such approaches, voxel size must be small enough to resolve object spacings, and if two objects are very close to each other the number of voxels can become prohibitively large.
+We present an algorithm to compute a quadtree on arbitrary datasets, including those with closely spaced and intersecting objects. \red{Keep going...}
 
-We present an algorithm to compute a GVD approximation on arbitrary datasets, including those with closely spaced objects.
-The approach applies a distance transform over an quadtree representation of the objects.
+\gray{The approach applies a distance transform over an quadtree representation of the objects.
 Our quadtree, its associated data structure, and our distance transform are novel and optimized to GVD approximation.
-For the remainder of the paper, ``GVD'' will refer to the approximated Generalized Voronoi Diagram.
+For the remainder of the paper, ``GVD'' will refer to the approximated Generalized Voronoi Diagram.}
 
-This paper demonstrates GVD computation on data beyond the computational abilities of previous algorithms, unlocking interesting and important applications.
-Our approach allows GVD-based proximity queries and other applications using a larger class of meaningful datasets.
+\gray{This paper demonstrates GVD computation on data beyond the computational abilities of previous algorithms, unlocking interesting and important applications.
+Our approach allows GVD-based proximity queries and other applications using a larger class of meaningful datasets.}
 
@@
 \end{enumerate}
+
+Each step is done in parallel either on vertices, quadtree cells or object facets (lines).
 
@@
 \label{sec:algorithm}
+
+\subsection{Detecting conflict cells}
+
+Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from Root to $C$.
+The address of the root cell is a special case and is defined as $R$.
+Figure \ref{fig:octree-cartesian-5} shows the addresses of each leaf cell in the quadtree.
+
+We define a smallest containing cell (SCC) to be the smallest internal quadtree node which entirely contains a given facet.
+Given a facet defined by two endpoints $p$ and $q$, the quadtree address of the SCC is the longest common prefix of the Morton codes of $p$ and $q$.
+In figure \ref{fig:octree-cartesian-3}, the address of the SCC of facet D is 10.
+The SCC of facet E is $R$.
+
+We construct an array of facet-SCC pairs, which is the same size as the number of facets. After allocation, SCC discovery for each facet is done in parallel. We call the array the SCC array.
+
+Next we sort the SCC array, which is demonstrated in figure \ref{fig:scc-sort}.
+We first sort using a radix lexicographical sort, then sort by SCC address size.
+This sort is done in parallel with Karras octree construction.
+
+\begin{figure}
+  \begin{minipage}{0.3\columnwidth}
+    \uline{raw} \\
+    10110 \\
+    11 \\
+    0001 \\
+    10110 \\
+    10 \\
+    010110 \\
+    10110 \\
+    00101 \\
+    0110100 \\
+    10 \\
+    11
+  \end{minipage}%
+  \begin{minipage}{0.3\columnwidth}
+    \uline{lexsort} \\
+    0001 \\
+    00101 \\
+    010110 \\
+    0110100 \\
+    10 \\
+    10 \\
+    10110 \\
+    10110 \\
+    10110 \\
+    11 \\
+    11
+  \end{minipage}%
+  \begin{minipage}{0.3\columnwidth}
+    \uline{lensort} \\
+    10 \\
+    10 \\
+    11 \\
+    11 \\
+    0001 \\
+    00101 \\
+    10110 \\
+    10110 \\
+    10110 \\
+    010110 \\
+    0110100
+  \end{minipage}
+  \caption{Sorting the SCC-facet pairs.
+We first sort lexicographically (\textit{lexsort}), then by length (\textit{lensort}), using a parallel radix sort.
+In the \textit{lexsort}, we implicitly pad with zeros on the right so that all values are the same length.
+\red{How exactly is the \textit{lensort} done?}}
+  \label{fig:scc-sort}
+\end{figure}
+
+Then we use the SCC array to find the conflict cells using algorithm \ref{alg:find-conflict-cells}.
+
+\algorithmspace
+\begin{algorithm}
+  \DontPrintSemicolon
+  \LinesNumbered
+  \KwIn{VertexQuadtree}
+  \BlankLine
+  \ForPar{leaf cell $L$}{
+    $L$.color = -1\;
+    \ForEach{cell $A$ in \directAncestors{$L$}}{
+      id := compute\_cell\_index($A$)\;
+      \ForEach{cell2facet in identical elements of cells2facets[id]}{
+        $f$ := cell2facet.facet\;
+        \If{$f$ intersects $L$}{
+          \If{$L$.color == -1} {
+            \tcp{First facet found}
+            \tcp{that intersects $L$}
+            $L$.color = $f$.color\;
+            $L$.facet[0] = $f$\;
+          }
+          \ElseIf{$L$.color != $f$.color} {
+            \tcp{Cell $L$ is ambiguous}
+            $L$.color = -2\;
+            $L$.facet[1] = $f$\;
+          }
+        }
+      }
+    }
+  } \label{alg:quadtree_intersections_end}
+\caption{FIND\_CONFLICT\_CELLS}
+\label{alg:find-conflict-cells}
+\end{algorithm}
+\algorithmspace
 

### 2016-09-30 paper.tex
Added conflict find figures

@@
-% ---------------------------------------------------------------------------
-% Author guideline and sample document for EG publication using LaTeX2e input
-% D.Fellner, v1.13, Jul 31, 2008
+\documentclass[submission]{gmp2017}
+%\documentclass{gmp2017}
 
-\documentclass{egpubl}
-\usepackage{eg2015}
+\title{Parallel Quadtree Construction on Collections of Objects}
 
-% --- for  Annual CONFERENCE
-%\ConferenceSubmission   % uncomment for Conference submission
-\ConferencePaper        % uncomment for (final) Conference Paper
-% \STAR                   % uncomment for STAR contribution
-% \Tutorial               % uncomment for Tutorial contribution
-% \ShortPresentation      % uncomment for (final) Short Conference Presentation
-% \Areas                  % uncomment for Areas contribution
-% \MedicalPrize           % uncomment for Medical Prize contribution
-% \Education              % uncomment for Education contribution
-%
-% --- for  CGF Journal
-% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
-% \JournalPaper         % uncomment for final version of Journal Paper
-%
-% --- for  CGF Journal: special issue
-% \SpecialIssueSubmission    % uncomment for submission to Computer Graphics Forum, special issue
-% \SpecialIssuePaper         % uncomment for final version of Journal Paper, special issue
-%
-% --- for  EG Workshop Proceedings
-% \WsSubmission    % uncomment for submission to EG Workshop
-% \WsPaper         % uncomment for final version of EG Workshop contribution
-%
- \electronicVersion % can be used both for the printed and electronic version
+% Put your submission number here.
+% This number will appear instead of authors and affiliations if the
+% class option "submission" is active.
+\SubNumber{142}
 
-% !! *please* don't change anything above
-% !! unless you REALLY know what you are doing
-% ------------------------------------------------------------------------
+% Put the author names here.
+% Use the second argument as a reference to the list of affiliations.
+% No authors and affiliations will appear if the class option "submission"
+% is active.
+\author{John Edwards}{1}
+\author{Nathan Vollmer}{1}
+\author{Nicholas Harrison}{1}
 
-% for including postscript figures
-% mind: package option 'draft' will replace PS figure by a filname within a frame
-%\ifpdf \usepackage[pdftex]{graphicx} \pdfcompresslevel=9
-%\else \usepackage[dvips]{graphicx} \fi
+% Put the affiliations of the authors here.
+\affiliation{1}{Department of Informatics and Computer Science, Idaho State University, USA}
 
-\PrintedOrElectronic
-
-% prepare for electronic version of your document
-\usepackage{t1enc,dfadobe}
-
-\usepackage{egweblnk}
-\usepackage{cite}
-\usepackage{balance}
-\usepackage{color}
-
-% For backwards compatibility to old LaTeX type font selection.
-% Uncomment if your document adheres to LaTeX2e recommendations.
-% \let\rm=\rmfamily    \let\sf=\sffamily    \let\tt=\ttfamily
-% \let\it=\itshape     \let\sl=\slshape     \let\sc=\scshape
-% \let\bf=\bfseries
-
-% end of prologue
-
-%\input{EGauthorGuidelines-body.inc}
-% ---------------------------------------------------------------------
-% EG author guidelines plus sample file for EG publication using LaTeX2e input
-% D.Fellner, v1.17, Sep 23, 2010
-
-\title[PGVD]%
-      {Parallel Quadtree Construction on Collections of Objects}
-
-% for anonymous conference submission please enter your SUBMISSION ID
-% instead of the author's name (and leave the affiliation blank) !!
-% For Computer Graphics Forum: Please use the abbreviation of your first name.
-
-%% author example
-%\author[D. Fellner \& S. Behnke]
-%       {D.\,W. Fellner\thanks{Chairman Eurographics Publications Board}$^{1,2}$
-%        and S. Behnke$^{2}$
-%        S. Spencer$^2$\thanks{Chairman Siggraph Publications Board} \\
-%         $^1$TU Darmstadt \& Fraunhofer IGD, Germany \\
-%         $^2$Institut f{\"u}r ComputerGraphik \& Wissensvisualisierung, TU Graz, Austria
-%       }
-
-\author[J. Edwards \& N. Vollmer \& N. Harrison]
-%       {John Edwards\thanks{jedwards@sci.utah.edu}$^1$ and
-%       {John Edwards\\
-       {John Edwards and
-         Nathan Vollmer and
-         Nicholas Harrison \\
-         \parbox[t]{10cm}{\centering Idaho State University}
-%         $^2$Google, Inc. \\
-%         \footnotesize{$^3$The University of Texas at Austin}}
-       }
-%\author[paper1040]
-%       {paper1040}
-
-% ------------------------------------------------------------------------
-
-% if the Editors-in-Chief have given you the data, you may uncomment
-% the following five lines and insert it here
-%
-% \volume{27}   % the volume in which the issue will be published;
-% \issue{1}     % the issue number of the publication
-% \pStartPage{1}      % set starting page
-
-\renewcommand{\paragraph}[1]{\noindent \textbf{#1}}
 \input{common}
@@
 
-\begin{classification} % according to http://www.acm.org/class/1998/
-  \CCScat{I.3.5}{Computer Graphics}{Computational Geometry and Object Modeling}{Boundary representations}
-  \CCScat{I.3.6}{Computer Graphics}{Methodology and Techniques}{Graphics data structures and data types}
-  %\CCScat{Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
-\end{classification}
+%\begin{classification} % according to http://www.acm.org/class/1998/
+%  \CCScat{I.3.5}{Computer Graphics}{Computational Geometry and Object Modeling}{Boundary representations}
+%  \CCScat{I.3.6}{Computer Graphics}{Methodology and Techniques}{Graphics data structures and data types}
+%  %\CCScat{Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
+%\end{classification}
 
@@
 
-\subsection{Detecting conflict cells}
+Our algorithm proceeds in three main steps:
+
+\begin{tightenumerate}
+\item Build initial octree on vertices $V$
+\item Detect conflict cells
+  \begin{tightenumerate}
+  \item Build smallest common cell (SCC) array
+  \item Sort SCC array
+  \item Find intersecting facets for each leaf cell using SCC array
+  \end{tightenumerate}
+\item Resolve conflict cells
+  \begin{tightenumerate}
+  \item Compute sample points $S$
+  \item Build new octree on $V \cup S$
+  \end{tightenumerate}
+\end{tightenumerate}
+
+Steps 1 and 2 are independent of dimension, and so our descriptions will use dimension-independent terms. In this paper, step 3 is described in 2D, with a 3D extension left to future work.
+
+\subsection{Build initial octree}
+Our first step is to build an octree on the vertices using Karras' algorithm. This step gives us our first approximation to our final octree.
+
+\subsection{Detect conflict cells}
 
@@
 The address of the root cell is a special case and is defined as $R$.
-Figure \ref{fig:octree-cartesian-5} shows the addresses of each leaf cell in the quadtree.
+Figure \ref{fig:conflict-find-4} shows the address of each leaf cell in the quadtree.
 
 We define a smallest containing cell (SCC) to be the smallest internal quadtree node which entirely contains a given facet.
-Given a facet defined by two endpoints $p$ and $q$, the quadtree address of the SCC is the longest common prefix of the Morton codes of $p$ and $q$.
-In figure \ref{fig:octree-cartesian-3}, the address of the SCC of facet D is 10.
-The SCC of facet E is $R$.
+Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the SCC is the longest common prefix of the Morton codes of the points in $P$.
+Figure \ref{fig:scc-sort-1} gives the addresses of the SCCs of the facets in figure \ref{fig:conflict-find-4}.
 
-We construct an array of facet-SCC pairs, which is the same size as the number of facets. After allocation, SCC discovery for each facet is done in parallel. We call the array the SCC array.
+We begin by constructing an array of facet-SCC pairs (see figure \ref{fig:scc-sort-1}), which is done in parallel over all facets.
 
@@
 \begin{figure}
-  \begin{minipage}{0.3\columnwidth}
-    \uline{raw} \\
-    10110 \\
-    11 \\
-    0001 \\
-    10110 \\
-    10 \\
-    010110 \\
-    10110 \\
-    00101 \\
-    0110100 \\
-    10 \\
-    11
-  \end{minipage}%
-  \begin{minipage}{0.3\columnwidth}
-    \uline{lexsort} \\
-    0001 \\
-    00101 \\
-    010110 \\
-    0110100 \\
-    10 \\
-    10 \\
-    10110 \\
-    10110 \\
-    10110 \\
-    11 \\
-    11
-  \end{minipage}%
-  \begin{minipage}{0.3\columnwidth}
-    \uline{lensort} \\
-    10 \\
-    10 \\
-    11 \\
-    11 \\
-    0001 \\
-    00101 \\
-    10110 \\
-    10110 \\
-    10110 \\
-    010110 \\
-    0110100
-  \end{minipage}
-  \caption{Sorting the SCC-facet pairs.
-We first sort lexicographically (\textit{lexsort}), then by length (\textit{lensort}), using a parallel radix sort.
+  \centering
+  \subfloat[][]{
+    \label{fig:conflict-find-1}
+    \includegraphics[width=0.24\columnwidth]{conflict-find-1.pdf} }
+  \subfloat[][]{
+    \label{fig:conflict-find-2}
+    \includegraphics[width=0.24\columnwidth]{conflict-find-2.pdf} }
+  \subfloat[][]{
+    \label{fig:conflict-find-4}
+    \includegraphics[width=0.24\columnwidth]{conflict-find-4.pdf} }
+  \subfloat[][]{
+    \label{fig:conflict-find-5}
+    \includegraphics[width=0.24\columnwidth]{conflict-find-5.pdf} } \\
+  \subfloat[][]{
+    \label{fig:conflict-find-6}
+    \includegraphics[width=0.24\columnwidth]{conflict-find-6.pdf} }
+  \subfloat[][]{
+    \label{fig:scc-sort-1}
+  \begin{minipage}{0.24\columnwidth}
+    \centering
+    raw \\
+    \vspace{0.1cm}
+    \begin{tabular}{|l|l|}
+      \hline
+      facet & SCC \\
+      \hline
+      A & 10 \\
+      B & 100110 \\
+      C & 1001 \\
+      D &  \\
+      E & 1010 \\
+      F & 10 \\
+      G & 10 \\
+      H & 10 \\
+      I & 011010 \\
+      \hline
+    \end{tabular}
+  \end{minipage} }
+  \subfloat[][]{
+    \label{fig:scc-sort-2}
+  \begin{minipage}{0.24\columnwidth}
+    \centering
+    lexsort \\
+    \vspace{0.1cm}
+    \begin{tabular}{|l|l|}
+      \hline
+      facet & SCC \\
+      \hline
+      D &  \\
+      I & 011010 \\
+      A & 10 \\
+      F & 10 \\
+      G & 10 \\
+      H & 10 \\
+      C & 1001 \\
+      B & 100110 \\
+      E & 1010 \\
+      \hline
+    \end{tabular}
+  \end{minipage} }
+  \subfloat[][]{
+    \label{fig:scc-sort-3}
+  \begin{minipage}{0.24\columnwidth}
+    \centering
+    lensort \\
+    \vspace{0.1cm}
+    \begin{tabular}{|l|l|}
+      \hline
+      facet & SCC \\
+      \hline
+      D &  \\
+      A & 10 \\
+      F & 10 \\
+      G & 10 \\
+      H & 10 \\
+      C & 1001 \\
+      E & 1010 \\
+      I & 011010 \\
+      B & 100110 \\
+      \hline
+    \end{tabular}
+  \end{minipage} }
+  \caption{
+    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and orange with facets labeled A-I.
+    \protect\subref{fig:conflict-find-2} We construct an initial quadtree on the vertices using Karras' algorithm.
+    \protect\subref{fig:conflict-find-4} We then compute the smallest common cell (SCC) for each facet. These pairs are given in figure \protect\subref{fig:scc-sort-1}.
+    \protect\subref{fig:conflict-find-5} Conflict cells, which intersect more than one object, are highlighted.
+    \protect\subref{fig:conflict-find-6} The new quadtree after conflict resolution.
+    \protect\subref{fig:scc-sort-1} The SCC-facet pairs are stored in an array initially sorted on facet index (letters are used here for clarity).
+We then sort the pairs using two stable, parallel radix sorts on SCC address for fast indexed access.
+    The first sort is \textit{lexsort}, a lexicographical sort \protect\subref{fig:scc-sort-2},
+    then comes \textit{lensort}, a sort by length \protect\subref{fig:scc-sort-3}.
 In the \textit{lexsort}, we implicitly pad with zeros on the right so that all values are the same length.
-\red{How exactly is the \textit{lensort} done?}}
+\red{How exactly is the \textit{lensort} done?}
+  }
+  \label{fig:conflict-find}
+\end{figure}
+
+\begin{figure}
   \label{fig:scc-sort}
@@
 
-\subsection{Resolving conflict cells}
+\subsection{Resolve conflict cells}
 
@@
     \label{fig:conflict-resolution-x}
-    \includegraphics[width=0.48\columnwidth]{conflict-resolution-x.pdf} }
+    \includegraphics[width=0.24\columnwidth]{conflict-resolution-x.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-even}
-    \includegraphics[width=0.48\columnwidth]{conflict-resolution-even.pdf} } \\
+    \includegraphics[width=0.24\columnwidth]{conflict-resolution-even.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-all}
-    \includegraphics[width=0.48\columnwidth]{conflict-resolution-all.pdf} }
+    \includegraphics[width=0.24\columnwidth]{conflict-resolution-all.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-octree}
-    \includegraphics[width=0.48\columnwidth]{conflict-resolution-octree.pdf} } \\
+    \includegraphics[width=0.24\columnwidth]{conflict-resolution-octree.pdf} } \\
   \subfloat[][]{
     \label{fig:conflict-resolution-adjacent-even}
-    \includegraphics[width=0.48\columnwidth]{conflict-resolution-adjacent-even.pdf} }
+    \includegraphics[width=0.24\columnwidth]{conflict-resolution-adjacent-even.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-adjacent-all}
-    \includegraphics[width=0.48\columnwidth]{conflict-resolution-adjacent-all.pdf} }
+    \includegraphics[width=0.24\columnwidth]{conflict-resolution-adjacent-all.pdf} }
   \caption{
@@
     \label{fig:octree-cartesian-1}
-    \includegraphics[width=0.48\columnwidth]{octree-cartesian-1.pdf} }
+    \includegraphics[width=0.24\columnwidth]{octree-cartesian-1.pdf} }
   \subfloat[][]{
     \label{fig:octree-cartesian-2}
-    \includegraphics[width=0.48\columnwidth]{octree-cartesian-2.pdf} } \\
+    \includegraphics[width=0.24\columnwidth]{octree-cartesian-2.pdf} }
   \subfloat[][]{
     \label{fig:octree-cartesian-3}
-    \includegraphics[width=0.48\columnwidth]{octree-cartesian-3.pdf} }
+    \includegraphics[width=0.24\columnwidth]{octree-cartesian-3.pdf} }
   \subfloat[][]{
     \label{fig:octree-cartesian-5}
-    \includegraphics[width=0.48\columnwidth]{octree-cartesian-5.pdf} } \\
+    \includegraphics[width=0.24\columnwidth]{octree-cartesian-5.pdf} } \\
   \subfloat[][]{
     \label{fig:octree-hierarchical-1}
-    \includegraphics[width=0.48\columnwidth]{octree-hierarchical-1.pdf} }
+    \includegraphics[width=0.24\columnwidth]{octree-hierarchical-1.pdf} }
   \subfloat[][]{
     \label{fig:octree-hierarchical-2}
-    \includegraphics[width=0.48\columnwidth]{octree-hierarchical-2.pdf} }
+    \includegraphics[width=0.24\columnwidth]{octree-hierarchical-2.pdf} }
   \caption{

### 2016-10-10 paper.tex
Added new data structure figures

@@
 \input{common}
+
+\usepackage[inline]{showlabels}
 
@@
 Modeling inter-object spacing is computationally straightforward when inter-object spacing is large compared to the world bounding box.
-Approaches typically involve a uniform grid of the space, which leads to efficient computation using, very often, graphics processors.
+Approaches typically involve a uniform grid of the space, which leads to efficient computation that often uses graphics processors.
 
-\gray{Whereas most algorithms are efficient and robust on certain datasets, all algorithms but one \cite{edwards2015approximating} require inordinate amounts of memory on datasets where objects are very closely spaced relative to the size of the domain.
-The failures occur because the space is uniformly gridded.
-In such approaches, voxel size must be small enough to resolve object spacings, and if two objects are very close to each other the number of voxels can become prohibitively large.}
+Difficulties arise when spacing between objects is small relative to the size of the domain.
+An approach using a uniform grid would have excessive memory requirements in order to resolve between objects.
+\red{What does ``resolve'' mean?} In these cases, teh uniformly sized grid cell must be small enough to fit between objects at every location in the domain.
+To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory.
 
-We present an algorithm to compute a quadtree on arbitrary datasets, including those with closely spaced and intersecting objects. \red{Keep going...}
-
-\gray{The approach applies a distance transform over an quadtree representation of the objects.
-Our quadtree, its associated data structure, and our distance transform are novel and optimized to GVD approximation.
-For the remainder of the paper, ``GVD'' will refer to the approximated Generalized Voronoi Diagram.}
-
-\gray{This paper demonstrates GVD computation on data beyond the computational abilities of previous algorithms, unlocking interesting and important applications.
-Our approach allows GVD-based proximity queries and other applications using a larger class of meaningful datasets.}
+We present an algorithm that builds a quadtree on arbitrarily spaced objects in parallel.
+This work extends the work done by Edwards et al \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm targeted for the GPU.
+Our algorithm performs an order of magnitude faster than the previous work and will be an excellent base for later distance transform and generalized Voronoi diagram computation.
 
@@
 
-Each step is done in parallel either on vertices, quadtree cells or object facets (lines).
+Each step is done in parallel either on vertices, quadtree cells or object facets.
 
@@
 \section{Related work}
-Related work falls into two categories: algorithms that compute the GVD and algorithms that compute distance fields, many of which are adaptive.
-
-\paragraph{Generalized Voronoi diagrams}
-A theoretical framework for generalized Voronoi diagrams can be found in Boissonnat \etal \shortcite{boissonnat2006curved}.
-Ordinary Voronoi diagrams are well studied and efficient algorithms exist that compute them exactly \cite{de2008computational}, but exact algorithms for the generalized Voronoi diagram are limited to a small set of special cases \cite{lee1982medial,karavelas2004robust}.
-In an early work, Lavender \etal \shortcite{lavender1992voronoi} define and compute GVDs over a set of solid models using an quadtree.
-Etzion and Rappoport \shortcite{etzion2002computing} represent the GVD bisector symbolically for lazy evaluation, but are limited to sites that are polyhedra.
-Boada \etal \shortcite{boada2002voronoi,boada2008approximations} use an adaptive approach to GVD computation, but their algorithm is restricted to GVDs with connected regions and is inefficient for polyhedral objects with many facets.
-Two other works are adaptive \cite{teichmann1997polygonal,vleugels1998approximating} but are computationally expensive and are restricted to convex sites.
-
-In recent years Voronoi diagram algorithms that take advantage of fast graphics hardware have become more common \cite{cao2010parallel,fischer2006fast,hsieh2005simple,rong2007variants,sud2006interactive,sud2006fast,hoff1999fast,wu2008gpu}.
-These algorithms are efficient and generalize well to the GVD, but most are limited to a subset of site types.
-More importantly, all of them use uniform grids and require an extraordinary number of voxels to resolve closely spaced objects (for example, Figs.
-\ref{fig:knife} and \ref{fig:path} would require $2^{36}$ and $2^{48}$ voxels, respectively).
-To our knowledge, ours is the first fully adaptive algorithm that computes the generalized Voronoi diagram for arbitrary datasets.
-
-\paragraph{Distance fields and quadtrees}
-The GVD is a subset of the locus of distance field critical points, a property that we take advantage of.
-In that light, the GVD could be a post-processing step to any method that computes a distance field.
-Distance transforms compute a distance field, but most are uniformly gridded \cite{jones20063d} and are thus no more suitable than GVD algorithms that use the GPU.
+In an early work, Lavender \etal \shortcite{lavender1992voronoi} define and compute quadtrees over a set of solid models.
+\red{Boada \etal \shortcite{boada2002voronoi,boada2008approximations} use an adaptive approach to GVD computation, but their algorithm is restricted to GVDs with connected regions and is inefficient for polyhedral objects with many facets.
+Two other works are adaptive \cite{teichmann1997polygonal,vleugels1998approximating} but are computationally expensive and are restricted to convex sites.}
 
@@
 % of the surface from the quadtree at rendering time.
-Two seminal works adaptively compute the Adaptive Distance Field (ADF) on quadtree vertices.
+Two seminal works build octrees on objects in order to compute the Adaptive Distance Field (ADF) on octree vertices.
 Strain~\shortcite{strain1999fast} fully resolves the quadtree everywhere on the object surface, and Frisken \etal~\shortcite{frisken2000adaptively} resolve the quadtree fully only in areas of small local feature size.
 Both approaches are designed to retain features of a single object rather than resolving between multiple objects, as is required for GVD computation.
-Qu \etal \shortcite{qu2004feature} implement an energy-minimizing distance field algorithm that preserves features at the expense of efficiency.
 Many recent works on fast quadtree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}.
+Kim and Liu's work~\shortcite{kim2014exact} is similar, computing the quadtree on the barycenters of triangles, giving an approximation of an object-resolving quadtree.
 Most quadtree approaches that support surfaces \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed} are designed for efficient rendering, and actual construction of the quadtree is implemented on the CPU.
 
-Two works \cite{bastos2008gpu,park2010cuda} implement the ADF using GPU parallelism to compute the distance value at sample points, but building the quadtree itself is done sequentially.
-Yin \etal~\shortcite{yin2011fast} compute the distance field entirely on the GPU using a bottom-up approach by initially subdividing into a complete quadtree, resulting in memory usage that is no better than using a uniform grid.
-A method by Kim and Liu~\shortcite{kim2014exact} computes the quadtree and a BVH entirely on the GPU.
-However, quadtree construction is performed on barycenters of triangles, and so a leaf quadtree cell can have an arbitrary number of triangle intersections as long as it contains no more than one triangle's barycenter.
-We have found no GPU quadtree construction method that can resolve between objects.
+Two works \cite{bastos2008gpu,park2010cuda} implement Adaptive Distance Fields in parallel on quadtrees but building the quadtree itself is done sequentially.
+Yin \etal~\shortcite{yin2011fast} compute the octree entirely on the GPU using a bottom-up approach by initially subdividing into a complete quadtree, resulting in memory usage that is no better than using a uniform grid.
+We have found no GPU quadtree construction method that is fully adaptive and can resolve between objects.
 
@@
 
+We refer to quadtree leaf cells that intersect two or more objects as ``conflict cells.'' A necessary and sufficient condition for a quadtree to resolve objects is to have no conflict cells.
+Our approach to computing such a quadtree is to first detexct conflict cells in parallel and then resolve them.
+
 Our algorithm proceeds in three main steps:
@@
 \begin{tightenumerate}
-\item Build initial octree on vertices $V$
+\item Build initial octree on vertices $V$ \cite{karras2012maximizing}
 \item Detect conflict cells
   \begin{tightenumerate}
-  \item Build smallest common cell (SCC) array
-  \item Sort SCC array
-  \item Find intersecting facets for each leaf cell using SCC array
+  \item Build bounding cells (BCells) array (Figure \ref{fig:data-structures-1})
+  \item Sort BCells array (Figure \ref{fig:data-structures-2})
+  \item Index BCells with octree data structure (Figure \ref{fig:data-structures-3})
+  \item Find facets that intersect each leaf cell
   \end{tightenumerate}
@@
   \end{tightenumerate}
+\item Iterate
 \end{tightenumerate}
@@
 \subsection{Build initial octree}
-Our first step is to build an octree on the vertices using Karras' algorithm. This step gives us our first approximation to our final octree.
+Our first step is to build an octree on the vertices using Karras' algorithm \cite{karras2012maximizing}. This step gives us our first approximation to our final octree.
 
@@
 The address of the root cell is a special case and is defined as $R$.
-Figure \ref{fig:conflict-find-4} shows the address of each leaf cell in the quadtree.
+Figure \ref{fig:conflict-find-4} shows the address of each leaf cell in a quadtree.
 
-We define a smallest containing cell (SCC) to be the smallest internal quadtree node which entirely contains a given facet.
-Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the SCC is the longest common prefix of the Morton codes of the points in $P$.
-Figure \ref{fig:scc-sort-1} gives the addresses of the SCCs of the facets in figure \ref{fig:conflict-find-4}.
+We define a bounding cell (BCell) to be the smallest internal quadtree node which entirely contains a given facet.
+Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the BCell is the longest common prefix of the Morton codes of the points in $P$.
+Figure \ref{fig:scc-sort-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.
 
-We begin by constructing an array of facet-SCC pairs (see figure \ref{fig:scc-sort-1}), which is done in parallel over all facets.
+We begin by constructing an array of BCells (see figure \ref{fig:data-structures-1}), which is done in parallel over all facets.
+Each facet $f$ computes the longest common prefix of its vertices and stores the result in \texttt{BCells[$f$].Cell\_ID}.
 
-Next we sort the SCC array, which is demonstrated in figure \ref{fig:scc-sort}.
-We first sort using a radix lexicographical sort, then sort by SCC address size.
-This sort is done in parallel with Karras octree construction.
+Next we sort the \texttt{BCells} array on the \texttt{Cell\_ID} field using a parallel radix lexicographical sort (Figure \ref{fig:data-structures-2}).
+\texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.
 
@@
     \label{fig:conflict-find-1}
-    \includegraphics[width=0.24\columnwidth]{conflict-find-1.pdf} }
+    \includegraphics[width=0.22\columnwidth]{conflict-find-1.pdf} } \hfill
   \subfloat[][]{
     \label{fig:conflict-find-2}
-    \includegraphics[width=0.24\columnwidth]{conflict-find-2.pdf} }
+    \includegraphics[width=0.22\columnwidth]{conflict-find-2.pdf} } \hfill
   \subfloat[][]{
     \label{fig:conflict-find-4}
-    \includegraphics[width=0.24\columnwidth]{conflict-find-4.pdf} }
+    \includegraphics[width=0.22\columnwidth]{conflict-find-4.pdf} } \hfill
   \subfloat[][]{
     \label{fig:conflict-find-5}
-    \includegraphics[width=0.24\columnwidth]{conflict-find-5.pdf} } \\
+    \includegraphics[width=0.22\columnwidth]{conflict-find-5.pdf} } \\
   \subfloat[][]{
     \label{fig:conflict-find-6}
-    \includegraphics[width=0.24\columnwidth]{conflict-find-6.pdf} }
+    \includegraphics[width=0.22\columnwidth]{conflict-find-6.pdf} } \hfill
   \subfloat[][]{
-    \label{fig:scc-sort-1}
-  \begin{minipage}{0.24\columnwidth}
-    \centering
-    raw \\
-    \vspace{0.1cm}
-    \begin{tabular}{|l|l|}
-      \hline
-      facet & SCC \\
-      \hline
-      A & 10 \\
-      B & 100110 \\
-      C & 1001 \\
-      D &  \\
-      E & 1010 \\
-      F & 10 \\
-      G & 10 \\
-      H & 10 \\
-      I & 011010 \\
-      \hline
-    \end{tabular}
-  \end{minipage} }
+    \label{fig:data-structures-1}
+    \includegraphics[width=0.22\columnwidth]{data-structures-1.pdf} } \hfill
   \subfloat[][]{
-    \label{fig:scc-sort-2}
-  \begin{minipage}{0.24\columnwidth}
-    \centering
-    lexsort \\
-    \vspace{0.1cm}
-    \begin{tabular}{|l|l|}
-      \hline
-      facet & SCC \\
-      \hline
-      D &  \\
-      I & 011010 \\
-      A & 10 \\
-      F & 10 \\
-      G & 10 \\
-      H & 10 \\
-      C & 1001 \\
-      B & 100110 \\
-      E & 1010 \\
-      \hline
-    \end{tabular}
-  \end{minipage} }
+    \label{fig:data-structures-2}
+    \includegraphics[width=0.22\columnwidth]{data-structures-2.pdf} } \hfill
   \subfloat[][]{
-    \label{fig:scc-sort-3}
-  \begin{minipage}{0.24\columnwidth}
-    \centering
-    lensort \\
-    \vspace{0.1cm}
-    \begin{tabular}{|l|l|}
-      \hline
-      facet & SCC \\
-      \hline
-      D &  \\
-      A & 10 \\
-      F & 10 \\
-      G & 10 \\
-      H & 10 \\
-      C & 1001 \\
-      E & 1010 \\
-      I & 011010 \\
-      B & 100110 \\
-      \hline
-    \end{tabular}
-  \end{minipage} }
+    \label{fig:data-structures-3}
+    \includegraphics[width=0.22\columnwidth]{data-structures-3.pdf} }
   \caption{
@@
     \protect\subref{fig:conflict-find-6} The new quadtree after conflict resolution.
-    \protect\subref{fig:scc-sort-1} The SCC-facet pairs are stored in an array initially sorted on facet index (letters are used here for clarity).
-We then sort the pairs using two stable, parallel radix sorts on SCC address for fast indexed access.
-    The first sort is \textit{lexsort}, a lexicographical sort \protect\subref{fig:scc-sort-2},
-    then comes \textit{lensort}, a sort by length \protect\subref{fig:scc-sort-3}.
-In the \textit{lexsort}, we implicitly pad with zeros on the right so that all values are the same length.
-\red{How exactly is the \textit{lensort} done?}
+    \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity).
+    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access.
+    \protect\subref{fig:data-structures-3} We then, in parallel on each element of BCells, store the index of the quadtree cell corresponding to the Cell\_ID and, if the previous element in the BCells array does not have an identical Cell\_ID, store the BCell index in the quadtree Facets field.
+\red{I don't think we need \texttt{BCells.Cell\_idx}.
+I don't use it in the algorithm description.}
   }
@@
 
-Then we use the SCC array to find the conflict cells using algorithm \ref{alg:find-conflict-cells}.
+Then we use the \texttt{BCells} array and octree data structure to find the conflict cells using algorithm \ref{alg:find-conflict-cells}.
+We process each leaf cell $L$ in parallel.
+We set $L$'s color to $-1$ (uninitialized).
+We then investigate each ancestor $A$ of $L$.
+We find the ancestors using the \texttt{Parent} field in the octree data structure.
+Using the \texttt{Facets} field, we find the first of possibly multiple facets bounded by $A$.
+The first facet's index is found using \texttt{BCells[Quadtree[$A$].Facets].Facet\_idx}.
+Any facet $f$ for which $A$ is the bounding cell could potentially intersect the leaf cell $L$.
+We test for intersection between $f$ and $L$ and \red{continue on here...}.
 

### 2016-10-13 paper.tex
Added text and complexity analysis

@@
 Previous parallel algorithms either spawn kernels hierarchically, separate points only, or make no hard guarantees of object separation.
-Our algorithm runs in \red{complexity?} in the average case and has excellent results in practice.
-We demonstrate with results on 2D and 3D datasets.
+We derive the time complexity and demonstrate excellent results in practice.
 
@@
 An approach using a uniform grid would have excessive memory requirements in order to resolve between objects.
-\red{What does ``resolve'' mean?} In these cases, teh uniformly sized grid cell must be small enough to fit between objects at every location in the domain.
+\red{What does ``resolve'' mean?} In these cases, the uniformly sized grid cell must be small enough to fit between objects at every location in the domain.
 To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory.
@@
 
-Our algorithm has three steps:
+Our algorithm has three main components:
 
@@
 We refer to quadtree leaf cells that intersect two or more objects as ``conflict cells.'' A necessary and sufficient condition for a quadtree to resolve objects is to have no conflict cells.
-Our approach to computing such a quadtree is to first detexct conflict cells in parallel and then resolve them.
+Our approach to computing such a quadtree is to first build an initial quadtree using a set $S$ of point samples.
+We initialize $S$ to be the object vertices.
+We then detect conflict cells in parallel, followed by augmenting $S$ with sample points such that a subsequent quadtree built on $S$ resolves conflict cells.
+If $S$ changed, then we iterate.
 
-Our algorithm proceeds in three main steps:
-
-\begin{tightenumerate}
-\item Build initial octree on vertices $V$ \cite{karras2012maximizing}
-\item Detect conflict cells
-  \begin{tightenumerate}
-  \item Build bounding cells (BCells) array (Figure \ref{fig:data-structures-1})
-  \item Sort BCells array (Figure \ref{fig:data-structures-2})
-  \item Index BCells with octree data structure (Figure \ref{fig:data-structures-3})
-  \item Find facets that intersect each leaf cell
-  \end{tightenumerate}
-\item Resolve conflict cells
-  \begin{tightenumerate}
-  \item Compute sample points $S$
-  \item Build new octree on $V \cup S$
-  \end{tightenumerate}
-\item Iterate
-\end{tightenumerate}
-
-Steps 1 and 2 are independent of dimension, and so our descriptions will use dimension-independent terms. In this paper, step 3 is described in 2D, with a 3D extension left to future work.
+Most of our algorithm is independent of dimension, and so we will use terms ``facet'' and ``octree'' (which lacks a dimension-independent term).
+The process of finding sample points to resolve conflict cells is in 2D in this paper, so for that section we will use ``line'' and ``quadtree''.
 
 \subsection{Build initial octree}
-Our first step is to build an octree on the vertices using Karras' algorithm \cite{karras2012maximizing}. This step gives us our first approximation to our final octree.
+\label{sec:build-initial-octree}
+Our first step is to build an octree on the vertices.
+This step gives us our first approximation to our final octree.
+We use Karras' algorithm \cite{karras2012maximizing} which sorts the Morton codes of the vertices in parallel, then constructs the binary radix tree in parallel.
+With the binary radix tree, the quadtree can be constructed with a single parallel call.
+The strength of this algorithm lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points.
+That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.
 
@@
 
-Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from Root to $C$.
-The address of the root cell is a special case and is defined as $R$.
+Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from Root to $C$, where the local address is the Morton code of a $2^{DIM}$ block of voxels.
+The address of the root cell is defined as the empty string.
 Figure \ref{fig:conflict-find-4} shows the address of each leaf cell in a quadtree.
@@
 Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the BCell is the longest common prefix of the Morton codes of the points in $P$.
-Figure \ref{fig:scc-sort-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.
+Figure \ref{fig:data-structures-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.
 
-We begin by constructing an array of BCells (see figure \ref{fig:data-structures-1}), which is done in parallel over all facets.
-Each facet $f$ computes the longest common prefix of its vertices and stores the result in \texttt{BCells[$f$].Cell\_ID}.
+We begin by constructing an array \texttt{BCells} and sibling array \texttt{FacetMap} (see figure \ref{fig:data-structures-1}), which is done in parallel over all facets.
+Each facet $f$ computes the longest common prefix of its vertices and stores the result in \texttt{BCells[$f$]}.
 
-Next we sort the \texttt{BCells} array on the \texttt{Cell\_ID} field using a parallel radix lexicographical sort (Figure \ref{fig:data-structures-2}).
+Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (Figure \ref{fig:data-structures-2}).
 \texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.
@@
     \label{fig:data-structures-2}
-    \includegraphics[width=0.22\columnwidth]{data-structures-2.pdf} } \hfill
+    \includegraphics[width=0.22\columnwidth]{data-structures-3.pdf} } \hfill
   \subfloat[][]{
     \label{fig:data-structures-3}
-    \includegraphics[width=0.22\columnwidth]{data-structures-3.pdf} }
+    \includegraphics[width=0.22\columnwidth]{data-structures-4.pdf} }
   \caption{
@@
     \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity).
+The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure) and start and end facets (``SFacet'' and ``EFacet'').
     \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access.
     \protect\subref{fig:data-structures-3} We then, in parallel on each element of BCells, store the index of the quadtree cell corresponding to the Cell\_ID and, if the previous element in the BCells array does not have an identical Cell\_ID, store the BCell index in the quadtree Facets field.
-\red{I don't think we need \texttt{BCells.Cell\_idx}.
-I don't use it in the algorithm description.}
   }
@@
 Then we use the \texttt{BCells} array and octree data structure to find the conflict cells using algorithm \ref{alg:find-conflict-cells}.
-We process each leaf cell $L$ in parallel.
+We process each leaf cell $L$ in parallel (line \ref{alg:L}).
 We set $L$'s color to $-1$ (uninitialized).
-We then investigate each ancestor $A$ of $L$.
+We then investigate each ancestor $A$ of $L$ (line \ref{alg:A}).
 We find the ancestors using the \texttt{Parent} field in the octree data structure.
-Using the \texttt{Facets} field, we find the first of possibly multiple facets bounded by $A$.
-The first facet's index is found using \texttt{BCells[Quadtree[$A$].Facets].Facet\_idx}.
+Using the \texttt{SFacet} and \texttt{EFacet} fields, we find, respectively, the first and (inclusive) last of possibly multiple facets bounded by $A$ (line \ref{alg:i}).
+We use the \texttt{FacetMap} array to find all facets bounded by bounding cell $A$ (line \ref{alg:f}).
 Any facet $f$ for which $A$ is the bounding cell could potentially intersect the leaf cell $L$.
-We test for intersection between $f$ and $L$ and \red{continue on here...}.
+We test for intersection between $f$ and $L$ and store the first two facets of differing color (lines \ref{alg:7}-\ref{alg:16}).
+If at the conclusion of execution $L$.color is equal to $-2$ then $L$ is a conflict cell and must be resolved.
 
@@
   \LinesNumbered
-  \KwIn{VertexQuadtree}
+  \KwIn{Quadtree}
   \BlankLine
-  \ForPar{leaf cell $L$}{
+  \ForPar{leaf cell $L$}{ \label{alg:L}
     $L$.color = -1\;
-    \ForEach{cell $A$ in \directAncestors{$L$}}{
-      id := compute\_cell\_index($A$)\;
-      \ForEach{cell2facet in identical elements of cells2facets[id]}{
-        $f$ := cell2facet.facet\;
-        \If{$f$ intersects $L$}{
+    \ForEach{cell $A$ in \directAncestors{$L$}}{  \label{alg:A}
+      \ForEach{i in \{SFacets[A]$\dots$EFacets[A]\}}{
+%      i := Quadtree[$A$].Facets\; \label{alg:i}
+%      \While{BCells[i].Cell\_ID = Address($A$)}{
+%        $f$ := Facets[BCells[i].Facet\_idx]\; \label{alg:f}
+        $f$ := Facets[FacetMap[i]]\; \label{alg:f}
+        \If{$f$ intersects $L$}{ \label{alg:7}
           \If{$L$.color == -1} {
-            \tcp{First facet found}
-            \tcp{that intersects $L$}
             $L$.color = $f$.color\;
@@
           }
-          \ElseIf{$L$.color != $f$.color} {
-            \tcp{Cell $L$ is ambiguous}
+          \ElseIf{$L$.color $\ne$ $f$.color} {
             $L$.color = -2\;
@@
           }
-        }
+        } \label{alg:16}
       }
@@
 
-A conflict cell is a quadtree cell that intersects at least two different objects.
-To resolve a conflict cell $c$, we consider pairs of lines of differing labels that intersect $c$.
-Figure \ref{fig:conflict-resolution-x} shows two lines
+To resolve a conflict cell $C$, we consider pairs of lines of differing labels that intersect $C$. Figure \ref{fig:conflict-resolution-x} shows two lines
 
@@
 that bisects $q$ and $r$.
-Our strategy will be to sample points $P$ on $p(s)$ (figure \ref{fig:conflict-resolution-octree}) such that an quadtree built on $V \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
+Our strategy will be to sample points $P$ on $p(s)$ (figure \ref{fig:conflict-resolution-octree}) such that an quadtree built on $S \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
 We do this by ensuring that $P$ is sampled such that every box that intersects both $q$ and $r$ also intersects at least two points in $P$.
@@
 We consider only cases where the slope of $p$ is in the range $0 \le m \le 1$.
-All other cases can be transformed to this case using rotation and reflection.
+All other instances can be transformed to this case using rotation and reflection.
 We begin by fitting the smallest box centered on a point $p$ that intersects both $q$ and $r$.
@@
 In both the \textit{opposite} and the \textit{adjacent} cases, $a(s)$ is of the form $a(s) = \hat{\alpha} s + \hat{\beta}$.
-We now use $a(s)$ to construct a sequence of $s$ values $S = \{s_0, s_1, s_2, \dots, s_n\}$ that meet our sampling criterion.
+We now use $a(s)$ to construct a sequence of values $S = \{s_0, s_1, s_2, \dots, s_n\}$ that meet our sampling criterion.
 We first construct the even samples (see Figures \ref{fig:conflict-resolution-even} and \ref{fig:conflict-resolution-adjacent-even}).
@@
 \begin{equation}
-s_i = k_1 + k_2 \alpha^i
+s_i = k_1 + k_2 \alpha^i \label{eqn:sample}
 \end{equation}
-where
+where the $k$ variables are split into those for even values of $i$ and those for odd values of $i$, and are given as
 \begin{align}
@@
 \begin{align}
-i < \log_{\alpha}\frac{s_{exit}-k_1}{k_2}
+i < \log_{\alpha}\frac{s_{exit}-k_1}{k_2} \label{eqn:num-samples}
 \end{align}
 
-%
-%.
-When $i$ is odd, the initial condition will be where $s$ is halfway along the first box (it could be anywhere in the interior of the first box, but we use halfway for convenience).
-So from \eqref{eqn:p},
-%\begin{align}
-%s_1 &= \frac{p^x-p_0^x}{2u^x} \\
-%    &= \frac{a(0)}{2u^x}
-%\end{align}
-%\begin{align}
-%k_1^{even} = k_1^{odd} &= \beta\left(1-\frac{\alpha}{\alpha-1}\right) \\
-%k_2^{even} &= \frac{\beta}{\alpha-1} \\
-%k_2^{odd} &= \frac{a(0)}{2u^x}+\frac{\beta}{\alpha-1}
-%\end{align}
-%
-%It is tempting to formulate the recurrance relation using a single initial condition. This would be done by replacing equation \eqref{eqn:xi_init} with
-%\begin{align}
-%x_i &= x_{i-1} + \frac{a(s_{i-1})}{2}
-%\end{align}
-%This formulation can lead to conflict cells as shown in figure \ref{fig:conflict-resolution-half}.
-%
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=0.48\columnwidth]{conflict-resolution-half.pdf}
-%  \caption{Using a single recurrence solution we can no longer guarantee that every box that intersects both lines contains at least two samples.
-The box in red contains only one sample, so the quadtree may contain conflict cells.
-%  }
-%  \label{fig:conflict-resolution-half}
-%\end{figure}
+In short, we first use equation \eqref{eqn:num-samples} compute the number of samples required to resolve the cell.
+The second step is to compute the samples themselves, which is done in parallel over all new samples to be computed, using equation \eqref{eqn:sample}.
 
-%With $a$ in hand, the question becomes, at what value of $s$ does $a(s)$ double?
-We desire to build a sequence $S=(s_0, s_1, \dots)$ such that $a(s_i) = 2a(s_{i-1})$.
-Substituting \eqref{eqn:p} into \eqref{eqn:a_prime} we get
-%\begin{align}
-%\label{eqn:klm}
-%2a(s_{i-1}) &= \frac{q_0^xv^y-q_0^yv^x+(p_0^y+s_iu^y)v^x-(p_0^x+s_iu^x)v^y}{v^x+v^y} \nonumber \\
-%&= \frac{q_0^xv^y-q_0^yv^x+p_0^yv^x-p_0^xv^y+s_i(u^yv^x-u^xv^y)}{v^x+v^y}
-%\end{align}
-%Solving for $s_i$ results in
-%\begin{equation}
-%\label{eqn:klm}
-%s_i = \frac{2(v^x+v^y)a(s_{i-1})-q_0^xv^y+q_0^yv^x-p_0^yv^x+p_0^xv^y}{u^yv^x-u^xv^y}
-%\end{equation}
-%
-%For $i>0$, $a(s_i)=a'(s_i)$, so we can characterize $s_i$ as a linear equation:
-%\begin{align}
-%s_0 &= 0 \nonumber \\
-%s_1 &= 2ma(s_0)+b \nonumber \\
-%s_2 &= 2^2ma(s_0)+b \nonumber \\
-%\cdots \nonumber \\
-%s_i &= 2^ima(s_0)+b
-%\label{eqn:si}
-%\end{align}
-%where 
-%\[m=\frac{v^x+v^y}{u^yv^x-u^xv^y}\]
-%and
-%\[b=\frac{-q_0^xv^y+q_0^yv^x-p_0^yv^x+p_0^xv^y}{u^yv^x-u^xv^y}\]
-%%where $m=(v^x+v^y)/(u^yv^x-u^xv^y)$ and $b=(-q_0^xv^y+q_0^yv^x-p_0^yv^x+p_0^xv^y)/(u^yv^x-u^xv^y)$.
-%
-%The number of elements in sequence $S$ is obtained by setting equation \eqref{eqn:si} less than one and solving for $i$:
-%\begin{equation}
-%i < \log_2{\frac{1-b}{ma(s_0)}}
-%\end{equation}
+\subsubsection{Iterate}
 
-\subsection{Build quadtree on vertices}
+Because conflict cell resolution only considers two facets at a time, we may have to iterate multiple times if more than two facets intersect a given cell.
+If new sample points were found then we add them to the current set $S$ of sample points and return to building the octree from points (section \ref{sec:build-initial-octree}).
+We finish when no new sample points are found.
 
-We first construct an quadtree on the vertices of the objects, which we call the ``vertex quadtree''.
-We use Karras' algorithm \cite{karras2012maximizing} which sorts the Morton codes of the vertices in parallel, then constructs the binary radix tree in parallel.
-With the binary radix tree, the quadtree can be constructed with a single parallel call.
-The strength of this algorithm lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points.
-That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.
-Given enough parallel units, the Karras algorithm runs in $O(\log{N})$ time, where $N$ is the number of vertices.
+\subsection{Time complexity analysis}
 
-\subsection{Identify conflict cells}
+Let $M = |F|$ and $N = |V|$, where $F$ are the object facets and $V$ are the object vertices.
+Let $D$ be the depth of the octree.
+In this analysis we assume sufficient parallel units to maximize parallelization.
 
-Our end goal is to construct an quadtree such that no quadtree cell intersects more than one object.
-Note that a cell is allowed to intersect more than one facet, but all facets must belong to the same object, or, in other words, all facets must share the same label.
-It is possible, but unlikely, that the vertex quadtree has this property.
-If so, then we are done.
-Otherwise, we must identify subdivide conflict cells.
+\begin{tightenumerate}
+\item Build octree using Karras' algorithm \cite{karras2012maximizing} - $O(D)$.
+\item Detect conflict cells
+  \begin{tightenumerate}
+  \item Build BCells array - $O(D)$.
+Building of the array runs in parallel for each facet $f$.
+The facet looks at each vertex (we assume simplices with a constant number of dimensions), computes Morton codes and finds the longest common prefix among vertices.
+This requires looking at each bit, of which there are $O(D)$.
+  \item Sort BCells array - $O(\log{M})$. The array has $M$ elements, and we use a parallel radix sort with $\log$ complexity.
+  \item Index BCells with octree data structure - $O(D)$. This runs in parallel on leaf cell IDs and each kernel requires a search of the octree for a given cell ID, taking at most $D$ steps.
+  \item Find facets that intersect each leaf cell - Worst case $O(M+D)$, average case $O(D)$.
+In unusual datasets, a single leaf cell will be intersected by $O(M)$ facets.
+On average, however, leaf cells intersect a small number of facets, and thus this step is dominated by the depth $D$ of the octree due to visiting each ancestor of the leaf cell.
+  \end{tightenumerate}
+\item Resolve conflict cells
+  \begin{tightenumerate}
+  \item Compute new sample points - $O(1)$.
+The first step computes, in parallel over conflict cells, the number of samples required to resolve the cell using equation \eqref{eqn:num-samples}.
+The second step is to compute the samples themselves, which is done in parallel over all new samples to be computed, using equation \eqref{eqn:sample}.
+  \item $S \leftarrow S \cup S'$ - $O(1)$.
+  \end{tightenumerate}
+\item Iterate - $O(Q)$ iterations.
+In the worst case, all facets intersect a single cell, requiring potentially $Q=O(M^2)$ iterations.
+In our testing, we have never seen $Q$ exceed \red{fill in with some value here}.
+\end{tightenumerate}
 
-One naive algorithm to identify conflict cells is to process each leaf cell $c$ in parallel and store which facets intersect $c$.
-This is $O(N)$.
-Another approach is to process each facet in parallel and add it to every cell that it intersects.
-This is $O(k\log{N})$ where $k$ is maximum number of cells that any facet intersects.
-As we will show, our algorithm is $O(j + \log{N})$ where $j$ is the maximum number of facets that intersect any cell.
-In practice, $\log{N} > j$, making our algorithm $O(\log{N})$.
-
-We identify conflict cells as shown in algorithm \ref{alg:find-conflict-cells}.
-In lines \ref{alg:quadtree_containing_begin}-\ref{alg:quadtree_containing_end}, for each internal quadtree cell $c$, we store all facets for which $c$ is the smallest containing cell.
-Since we are implementing this in a GPGPU environment, we don't have dynamic memory, so each facet must be processed twice.
-The first loop discovers how many facets are to be stored in each cell after which we allocate space for the facets.
-We use parallel prefix sums to determine the amount of space we need to allocate as well as the offsets for each internal cell.
-The second loop actually stores the facets.
-
-The $\container{f}$ procedure finds the smallest quadtree cell that fully contains the facet $f$.
-A straightforward implementation of $\container{f}$ is to perform a standard quadtree search on the vertices of $f$ and take the smallest quadtree cell that contains all of them.
-(Note that the cell is always an internal node, since a post-condition of the Karras algorithm is that no leaf cell contains more than one vertex.) In our implementation however, we take advantage of our existing data structures.
-The quadtree cell that contains a vertex $v$ is uniquely determined by the D-tuple bits of its morton code.
-For example, if a 2D vertex has morton code $010010$, then the quadtree is traversed from the root to child $01$ to child $00$ to child $10$.
-To determine $\container{f}$, we find the longest common prefix (\lcp) of the vertices.
-Truncating the length of \lcp\,to a multiple of D, we find the smallest quadtree cell that contains all vertices of $f$.
-The complexity of $\container{f}$ is $O(\log{N})$ for both implementations.
-Thus, lines \ref{alg:quadtree_containing_begin}-\ref{alg:quadtree_containing_end} run in $O(\log{N})$ time.
-
-Lines \ref{alg:quadtree_intersections_begin}-\ref{alg:quadtree_intersections_end} of the algorithm identify and store all facets that intersect with a given leaf cell $c$.
-Again, it is done in two steps for memory allocation purposes.
-Each leaf cell $c$ looks at its $O(\log{N})$ ancestors and tests all facets stored in those ancestors for intersection with $c$.
-Any intersecting facets get stored in $c$.
-These lines run in $O(F)$ time, where $F$ is the number of facets in all objects.
-Even though the loop is doubly-nested, each facet is stored in a unique internal node, so no more than $F$ facets will be visited in the loops.
-In practice, far fewer than $F$ facets will be checked for each leaf cell, because most datasets have facets that are completely contained in internal cells that are reasonably low in the tree.
-
-The entire conflict cell detection algorithm runs in $O(\log{N}+L) = O(L)$ because $L > \log{N}$.
-However, average case is $O(\log{N})$, considering that most lines are contained entirely in a cell at reasonably low depth.
-
-In Step 4, Stack is preallocated to size $M\cdot2^D$ where $M$ is the maximum quadtree depth and $D$ is the dimension.
-A conflict cell is a cell that intersects at least two different objects, or two lines of different labels.
-
-The second procedure we use is $\directAncestors{c}$, which finds all ancestors of quadtree cell $c$.
-
-%\SetKwFor{ForPar}{for}{do in parallel}{endfpar}
-
-\algorithmspace
-\begin{algorithm}
-  \DontPrintSemicolon
-  \LinesNumbered
-  \KwIn{VertexQuadtree}
-  \BlankLine
-  \tcp{Store contained facets}
-  \ForPar{facet $f$ in Objects}{
-    \label{alg:quadtree_containing_begin}
-    a := $\container{f}$\;
-    a.numFacets := a.numFacets + 1\;
-  }
-  Allocate space for facets in internal cells\;
-  \ForPar{facet $f$ in Objects}{
-    a := $\container{f}$\;
-    a.facets := a.facets $\cup$ f
-  } \label{alg:quadtree_containing_end}
-  \tcp{Store intersecting facets}
-  \ForPar{leaf cell c in VertexQuadtree}{
-    \label{alg:quadtree_intersections_begin}
-    \ForEach{cell a in \directAncestors{c}}{
-      \ForEach{facet $f$ in a.facets}{
-        \If{$f$ intersects c}{
-          c.numFacets := c.numFacets + 1\;
-        }
-      }
-    }
-  }
-  Allocate space for facets in leaf cells\;
-  \ForPar{leaf cell c in VertexQuadtree}{
-    \ForEach{cell a in \directAncestors{c}}{
-      \ForEach{facet $f$ in a.facets}{
-        \If{$f$ intersects c}{
-          c.facets := c.facets $\cup$ f\;
-        }
-      }
-    }
-  } \label{alg:quadtree_intersections_end}
-\caption{FIND\_CONFLICT\_CELLS}
-\label{alg:find-conflict-cells}
-\end{algorithm}
-\algorithmspace
-
-\algorithmspace
-\begin{algorithm}
-  \DontPrintSemicolon
-  \LinesNumbered
-  \KwIn{Quadtree, conflict\_cells}
-  \BlankLine
-  \tcp{4. Quadtree refinement}
-  \ForPar{leaf cell c in Quadtree}{
-    c' := c\;
-    \While{c' $\in$ conflict\_cells}{
-      $(c'_0,c'_1,\dots,c'_{2^D-1})$ := subdivide c'\;
-      push $(c'_0,c'_1,\dots,c'_{2^D-1})$ onto Stack\;
-      c' := Stack.pop
-    }
-  }
-\caption{REFINE\_QUADTREE}
-\label{alg:refine-quadtree}
-\end{algorithm}
-\algorithmspace
-
-\begin{figure}
-  \centering
-  \subfloat[][]{
-    \label{fig:octree-cartesian-1}
-    \includegraphics[width=0.24\columnwidth]{octree-cartesian-1.pdf} }
-  \subfloat[][]{
-    \label{fig:octree-cartesian-2}
-    \includegraphics[width=0.24\columnwidth]{octree-cartesian-2.pdf} }
-  \subfloat[][]{
-    \label{fig:octree-cartesian-3}
-    \includegraphics[width=0.24\columnwidth]{octree-cartesian-3.pdf} }
-  \subfloat[][]{
-    \label{fig:octree-cartesian-5}
-    \includegraphics[width=0.24\columnwidth]{octree-cartesian-5.pdf} } \\
-  \subfloat[][]{
-    \label{fig:octree-hierarchical-1}
-    \includegraphics[width=0.24\columnwidth]{octree-hierarchical-1.pdf} }
-  \subfloat[][]{
-    \label{fig:octree-hierarchical-2}
-    \includegraphics[width=0.24\columnwidth]{octree-hierarchical-2.pdf} }
-  \caption{
-    \protect\subref{fig:octree-cartesian-1} A red object and a blue object.
-    \protect\subref{fig:octree-cartesian-2} The vertex quadtree, or quadtree built on the object vertices using Karras' algorithm.
-    \protect\subref{fig:octree-cartesian-3}, \protect\subref{fig:octree-hierarchical-1} The vertex quadtree with conflict cells highlighted.
-Note the label of an quadtree cell in \protect\subref{fig:octree-cartesian-3} is the concatenation of labels from root R to the leaf cell in \protect\subref{fig:octree-hierarchical-1}.
-This value also corresponds to the highest order bits of the morton code of any point in the cell.
-    \protect\subref{fig:octree-cartesian-5}, \protect\subref{fig:octree-hierarchical-2} The quadtree after resolution of conflict cells.
-  }
-  \label{fig:steps}
-\end{figure}
-
-In Fig.
-\ref{fig:steps}, R (Root) is the smallest containing cell for lines A, B, and C, cell 20 contains line D, and cell 2 contains lines E and F.
-After Step 3 of the algorithm, line A is stored in leaf cells 202, 203, 21, and 3.
-Conflict cells, which are the only cells that are subdivided, are 203 and 21.
-
-%\begin{table}
-%  \centering
-%  \begin{tabular}{ll}
-%    \toprule
-%    line &  $\container{line}$ \\
-%    \midrule
-%    A & R \\
-%    B & R \\
-%    C & R \\
-%    D & 20 \\
-%    E & 2 \\
-%    F & 2 \\
-%    \bottomrule
-%  \end{tabular}
-%\end{table}
-
-%-----------------------------------------------------------
-% Compute GVD surface
-%-----------------------------------------------------------
-\section{Compute GVD surface}
-\label{sec:bisector}
+The final complexity of each iteration is $O(M+D)$ worst case and $O(\log{M}+D)$ average case.
+In practice we must fix the depth of the octree to a constant value in order to use a predetermined integer size for the Morton codes, which brings the average case complexity to $O(\log{M})$.
+Taking iteration into account, the final complexity is $(Q\log{M})$ average case.
 
@@
 
-%\bibliographystyle{eg-alpha}
-\bibliographystyle{eg-alpha-doi}
+\bibliographystyle{abbrv}
 \balance

### 2016-10-17 paper.tex
Figures, space complexity, abstract and related work

@@
 We present a parallel quadtree algorithm that resolves between geometric objects.
-The quadtree has the property that no quadtree cell intersects more than one labeled object.
-Previous parallel algorithms either spawn kernels hierarchically, separate points only, or make no hard guarantees of object separation.
-We derive the time complexity and demonstrate excellent results in practice.
+The quadtree has the property that no cell intersects more than one labeled object.
+A popular technique for discretizing space is to impose a uniform grid, an approach that is easily parallelizable, but this approach either fails because object separation isn't known a priori, or fails because the number of cells required to resolve closely spaced objects exceeds available memory.
+Previous parallel algorithms that are spatially adaptive, or that discretize finely in only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation.
+Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
+We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D.
+Our results show an order of magnitude improvement over the current state of the art.
 
@@
 \label{sec:intro}
-Constructing quadtrees on objects in an important task with applications to collision detection, distance fields, robot navigation, object description, and other applications.
+Constructing quadtrees on objects in an important task with applications in collision detection, distance fields, robot navigation, object description, and other applications.
 Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects.
-Our work however, centers on using quadtrees to separate, or resolve, collections of closely spaced objects.
+Our work however, centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object.
 Using a quadtree, we can model the space between objects, the first step in constructing distance fields, detecting collisions, and computing the generalized Voronoi diagram.
@@
 Difficulties arise when spacing between objects is small relative to the size of the domain.
-An approach using a uniform grid would have excessive memory requirements in order to resolve between objects.
-\red{What does ``resolve'' mean?} In these cases, the uniformly sized grid cell must be small enough to fit between objects at every location in the domain.
-To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory.
+An approach using a uniform grid would have excessive memory requirements in order to resolve between objects because the uniformly sized grid cell must be small enough to fit between objects at every location in the domain.
+Thus, an adaptive approach must be used for datasets of closely spaced objects.
+To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory but does so in serial, with expected performance liabilities.
 
-We present an algorithm that builds a quadtree on arbitrarily spaced objects in parallel.
 This work extends the work done by Edwards et al \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm targeted for the GPU.
@@
 
-Each step is done in parallel either on vertices, quadtree cells or object facets.
+Each step is done in parallel either on object vertices, object facets or quadtree cells.
 
@@
 \section{Related work}
-In an early work, Lavender \etal \shortcite{lavender1992voronoi} define and compute quadtrees over a set of solid models.
-\red{Boada \etal \shortcite{boada2002voronoi,boada2008approximations} use an adaptive approach to GVD computation, but their algorithm is restricted to GVDs with connected regions and is inefficient for polyhedral objects with many facets.
-Two other works are adaptive \cite{teichmann1997polygonal,vleugels1998approximating} but are computationally expensive and are restricted to convex sites.}
 
+\paragraph{Serial} In an early work, Lavender \etal \shortcite{lavender1992voronoi} define and compute octrees over a set of solid models. 
 % JME: note that in the Bastos work, ``reconstruction'' is reconstruction
@@
 Both approaches are designed to retain features of a single object rather than resolving between multiple objects, as is required for GVD computation.
-Many recent works on fast quadtree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}.
+Boada \etal \shortcite{boada2002voronoi,boada2008approximations} use an adaptive approach to GVD computation, but their algorithm is restricted to GVDs with connected regions and is inefficient for polyhedral objects with many facets.
+Two other works are adaptive \cite{teichmann1997polygonal,vleugels1998approximating} but are computationally expensive and are restricted to convex sites.
+
+\paragraph{Parallel} Many recent works on fast quadtree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}.
 Kim and Liu's work~\shortcite{kim2014exact} is similar, computing the quadtree on the barycenters of triangles, giving an approximation of an object-resolving quadtree.
 Most quadtree approaches that support surfaces \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed} are designed for efficient rendering, and actual construction of the quadtree is implemented on the CPU.
-
 Two works \cite{bastos2008gpu,park2010cuda} implement Adaptive Distance Fields in parallel on quadtrees but building the quadtree itself is done sequentially.
@@
 Most of our algorithm is independent of dimension, and so we will use terms ``facet'' and ``octree'' (which lacks a dimension-independent term).
-The process of finding sample points to resolve conflict cells is in 2D in this paper, so for that section we will use ``line'' and ``quadtree''.
+The process of finding sample points to resolve conflict cells is limited to 2D in this paper, so for that section we will use ``line'' and ``quadtree''.
 
@@
   \centering
+  \begin{tikzpicture}
+    \node[anchor=south west,inner sep=0,outer sep=0] at (0,0) {
       \subfloat[][]{
@@
         \label{fig:conflict-find-5}
-    \includegraphics[width=0.22\columnwidth]{conflict-find-5.pdf} } \\
+        \includegraphics[width=0.22\columnwidth]{conflict-find-5.pdf} }
+    };
+    %\draw[red,ultra thick] (4.0,1.7) rectangle (6.55,4.2);
+    \draw[red,thick] (4.0,1.7) rectangle (6.55,4.2);
+
+    % lines
+    % upper-right
+    \draw[red,thin] (6.55,4.2) -- (11.8,4.2);
+    % lower-left
+    \draw[red,thin] (4.0,1.7) -- (8.0,0.3);
+
+    %% curves
+    %% upper-left
+    %%\draw[red,thin] (4.0,4.2) to[out=10,in=170] (8.0,4.2);
+    %% upper-right
+    %\draw[red,thin] (6.55,4.2) to[out=10,in=170] (11.8,4.2);
+    %% lower-left
+    %\draw[red,thin] (4.0,1.7) to[out=-40,in=175] (8.0,0.3);
+
+    \draw[red,thick] (8.0,0.3) rectangle (11.8,4.2);
+  \end{tikzpicture} \\
       \subfloat[][]{
@@
   \caption{
-    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and orange with facets labeled A-I.
+    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and green with facets labeled A-I.
     \protect\subref{fig:conflict-find-2} We construct an initial quadtree on the vertices using Karras' algorithm.
-    \protect\subref{fig:conflict-find-4} We then compute the smallest common cell (SCC) for each facet. These pairs are given in figure \protect\subref{fig:scc-sort-1}.
+    \protect\subref{fig:conflict-find-4} Zoomed-in view of the boundary cell (BCell) computation for each facet. These pairs are given in figure \protect\subref{fig:data-structures-1}.
     \protect\subref{fig:conflict-find-5} Conflict cells, which intersect more than one object, are highlighted.
@@
     \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity).
-The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure) and start and end facets (``SFacet'' and ``EFacet'').
+The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure).
     \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access.
-    \protect\subref{fig:data-structures-3} We then, in parallel on each element of BCells, store the index of the quadtree cell corresponding to the Cell\_ID and, if the previous element in the BCells array does not have an identical Cell\_ID, store the BCell index in the quadtree Facets field.
+We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given octree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
+\protect\subref{fig:data-structures-3} For a given octree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
   }
@@
 We find the ancestors using the \texttt{Parent} field in the octree data structure.
-Using the \texttt{SFacet} and \texttt{EFacet} fields, we find, respectively, the first and (inclusive) last of possibly multiple facets bounded by $A$ (line \ref{alg:i}).
+Using the \texttt{FFacet} and \texttt{LFacet} fields, we find, respectively, the first and (inclusive) last of possibly multiple facets bounded by $A$ (line \ref{alg:i}).
 We use the \texttt{FacetMap} array to find all facets bounded by bounding cell $A$ (line \ref{alg:f}).
@@
     \ForEach{cell $A$ in \directAncestors{$L$}}{  \label{alg:A}
-      \ForEach{i in \{SFacets[A]$\dots$EFacets[A]\}}{
+      \ForEach{i in \{FFacet[A]$\dots$LFacet[A]\}}{ \label{alg:i}
 %      i := Quadtree[$A$].Facets\; \label{alg:i}
@@
     }
-  } \label{alg:quadtree_intersections_end}
+  }
 \caption{FIND\_CONFLICT\_CELLS}
@@
 
-%\begin{adjustbox}{valign=t}
 \begin{figure}
@@
 \end{figure}
-%    \end{adjustbox}
 
-%\begin{equation}
-%\label{eqn:p}
-%p(s) = p = p_0 + su
-%\end{equation}
-%\begin{equation}
-%\label{eqn:q}
-%q(s) = q = q_0 + tv
-%\end{equation}
-%\begin{equation}
-%\label{eqn:a_}
-%a(s) = a = |p^x-q^x| = |p^y-q^y|
-%\end{equation}
+We present a conflict cell resolution algorithm for pairs of lines in 2D.
+For a conflict cell $C$ our approach is to find sample points inside the cell such that no leaf cells in a quadtree constructed over the sample points intersect both lines.
+In this section we derive equation \eqref{eqn:num-samples} which computes the number of samples required to resolve the cell, and equation \eqref{eqn:sample} which computes the samples themselves.
+The power of our approach lies in the fact that both expressions are closed-form and neither one is iterative, so we can evaluate the first in parallel over leaf cells and the second in parallel over all samples to compute.
 
@@
 
-In short, we first use equation \eqref{eqn:num-samples} compute the number of samples required to resolve the cell.
-The second step is to compute the samples themselves, which is done in parallel over all new samples to be computed, using equation \eqref{eqn:sample}.
-
 \subsubsection{Iterate}
@@
 
-\subsection{Time complexity analysis}
+\subsection{Complexity analysis}
 
@@
 In this analysis we assume sufficient parallel units to maximize parallelization.
+
+\vspace{2mm}
+\paragraph{Time complexity}
 
@@
 Taking iteration into account, the final complexity is $(Q\log{M})$ average case.
+
+\vspace{2mm}
+\paragraph{Space complexity}
+
+The primary data structures are shown in figure \ref{fig:data-structures-1}.
+The quadtree data structure is size $O(|S|)$ and the remaining arrays are of size $M$.
+As $|S| \ge M$, our final space complexity is $O(|S|)$.
+The number of samples in $S$ depends on the dataset.
+In 2D, in the worst case, the facets can form an arrangement of maximum number of intersections, which is $M(M-1)/2 = O(M^2)$.
+If this is the case then we subdivide to the maximum octree depth at each intersection, causing an octree of size $O(DM^2)$.
+In our testing, our octrees have never exceeded a size of \red{fill this in with experimental results}.
 

### 2016-11-13 paper.tex
Changes
Various changes, include results section

@@
 % class option "submission" is active.
-\SubNumber{142}
+\SubNumber{19}
 
@@
 % is active.
+\author{Nathan Vollmer}{1}
 \author{John Edwards}{1}
-\author{Nathan Vollmer}{1}
-\author{Nicholas Harrison}{1}
 
@@
 
-\usepackage[inline]{showlabels}
+%\usepackage[inline]{showlabels}
 
@@
 We present a parallel quadtree algorithm that resolves between geometric objects.
-The quadtree has the property that no cell intersects more than one labeled object.
-A popular technique for discretizing space is to impose a uniform grid, an approach that is easily parallelizable, but this approach either fails because object separation isn't known a priori, or fails because the number of cells required to resolve closely spaced objects exceeds available memory.
+Our quadtree has the property that no cell intersects more than one labeled object.
+A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori, because the number of cells required to resolve closely spaced objects exceeds available memory.
 Previous parallel algorithms that are spatially adaptive, or that discretize finely in only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation.
@@
 \label{sec:intro}
-Constructing quadtrees on objects in an important task with applications in collision detection, distance fields, robot navigation, object description, and other applications.
+Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, robot navigation, object description, and other applications.
 Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects.
-Our work however, centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object.
+However, our work centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object.
 Using a quadtree, we can model the space between objects, the first step in constructing distance fields, detecting collisions, and computing the generalized Voronoi diagram.
@@
 
-Each step is done in parallel either on object vertices, object facets or quadtree cells.
+Each step is done in parallel either on object vertices, object facets, or quadtree cells.
 
@@
 We refer to quadtree leaf cells that intersect two or more objects as ``conflict cells.'' A necessary and sufficient condition for a quadtree to resolve objects is to have no conflict cells.
-Our approach to computing such a quadtree is to first build an initial quadtree using a set $S$ of point samples.
+Our approach to computing such a quadtree is to first build an initial quadtree, called the ``vertex quadtree,'' using a set $S$ of point samples.
 We initialize $S$ to be the object vertices.
 We then detect conflict cells in parallel, followed by augmenting $S$ with sample points such that a subsequent quadtree built on $S$ resolves conflict cells.
-If $S$ changed, then we iterate.
+If $S$ changed, then we iterate (see section \ref{sec:iterate}).
 
@@
 
-Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (Figure \ref{fig:data-structures-2}).
-\texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.
+Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (Figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.
 
@@
 We process each leaf cell $L$ in parallel (line \ref{alg:L}).
-We set $L$'s color to $-1$ (uninitialized).
-We then investigate each ancestor $A$ of $L$ (line \ref{alg:A}).
-We find the ancestors using the \texttt{Parent} field in the octree data structure.
+First, we set $L$'s color to $-1$ (uninitialized).
+We then investigate each ancestor $A$ of $L$ (line \ref{alg:A}) by using the \texttt{Parent} field in the octree data structure.
 Using the \texttt{FFacet} and \texttt{LFacet} fields, we find, respectively, the first and (inclusive) last of possibly multiple facets bounded by $A$ (line \ref{alg:i}).
-We use the \texttt{FacetMap} array to find all facets bounded by bounding cell $A$ (line \ref{alg:f}).
+The \texttt{FacetMap} array is used to find all facets bounded by bounding cell $A$ (line \ref{alg:f}).
 Any facet $f$ for which $A$ is the bounding cell could potentially intersect the leaf cell $L$.
@@
 We present a conflict cell resolution algorithm for pairs of lines in 2D.
-For a conflict cell $C$ our approach is to find sample points inside the cell such that no leaf cells in a quadtree constructed over the sample points intersect both lines.
-In this section we derive equation \eqref{eqn:num-samples} which computes the number of samples required to resolve the cell, and equation \eqref{eqn:sample} which computes the samples themselves.
+For a conflict cell $C$, our approach is to find sample points inside the cell such that no leaf cells in a quadtree constructed over the sample points intersect both lines.
+In this section we derive equation \eqref{eqn:num-samples} which computes the number of samples required to resolve the cell.
+We also derive equation \eqref{eqn:sample} which computes the samples themselves.
 The power of our approach lies in the fact that both expressions are closed-form and neither one is iterative, so we can evaluate the first in parallel over leaf cells and the second in parallel over all samples to compute.
@@
 We begin by fitting the smallest box centered on a point $p$ that intersects both $q$ and $r$.
-We break the problem into two cases: the \textit{opposite} case (see Figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
-The \textit{adjacent} case (see Figure \ref{fig:conflict-resolution-adjacent-even}) is where $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
+We break the problem into two cases:
+\begin{tightenumerate}
+ \item The \textit{opposite} case (see Figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
+ \item In the \textit{adjacent} case (see Figure \ref{fig:conflict-resolution-adjacent-even}), $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
+\end{tightenumerate}
 
@@
 \subsubsection{Iterate}
+\label{sec:iterate}
 
@@
 If new sample points were found then we add them to the current set $S$ of sample points and return to building the octree from points (section \ref{sec:build-initial-octree}).
-We finish when no new sample points are found.
+We finish when the only conflicts identified are at the maximum depth, i.e., unresolvable.
 
@@
   \end{tightenumerate}
-\item Iterate - $O(Q)$ iterations.
-In the worst case, all facets intersect a single cell, requiring potentially $Q=O(M^2)$ iterations.
-In our testing, we have never seen $Q$ exceed \red{fill in with some value here}.
+\item Iterate - $O(Q)$ iterations. In the worst case, all facets intersect a single cell, requiring potentially $Q=O(M^2)$ iterations. In our testing, $Q$ has not exceeded 3.
 \end{tightenumerate}
@@
 If this is the case then we subdivide to the maximum octree depth at each intersection, causing an octree of size $O(DM^2)$.
-In our testing, our octrees have never exceeded a size of \red{fill this in with experimental results}.
+%In our testing, our octrees have never exceeded a size of \red{fill this in with experimental results}.
 
@@
 %-------------------------------------------------------------------------------
-\section{Results and applications}
-Our implementation\footnote{Source code is available at \url{http://cedmav.org/research/project/33-gvds.html}.} of the algorithm supports \red{polygons and} triangulated objects, and our wavefront initialization step is implemented on the GPU using OpenCL.
-All tests were run on a MacBook Pro laptop with a dual-core 2.9 GHz processor, 8 GB memory, and Intel HD 4000 graphics card.
-Figure \ref{fig:bunny} shows our implementation of the GVD computation pipeline, and Figure \ref{fig:pipes} shows the computed GVD on a more challenging dataset.
-We compare our method with other work and then show examples in three application settings: path planning, proximity queries, and exploded diagrams.
+\section{Results and conclusions}
+Our implementation\footnote{
+Source code will be made available at our website. %is available at \url{http://cedmav.org/research/project/33-gvds.html}.
+}
+of the algorithm supports polygons and polylines which needn't be manifold or connected.
+All tests were run on a MacBook Pro laptop with a dual-core 2.9 GHz processor, 16 GB memory, and Intel Iris Graphics 6100 graphics card.
+Figure \ref{fig:results} shows results on four datasets: a simple toy dataset showing conflict cell detection and resolution (\ref{fig:simple-conflicts}-\ref{fig:simple-samples}); a more complex maze dataset (\ref{fig:glines}), and a complex dataset with many objects at very different scales (\ref{fig:ut1-samples}-\ref{fig:ut5}).
+Table \ref{tab:timings} shows timings for our implementation compared to the previous state-of-the-art.
+Our implementation is significantly faster and also generates fewer quadtree cells.
 
-\subsection{Comparison to other methods}
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:simple-conflicts}
+    \includegraphics[width=0.24\columnwidth]{simple-conflict.png} }
+  \subfloat[][]{
+    \label{fig:simple-samples}
+    \includegraphics[width=0.24\columnwidth]{simple-samples.png} }
+%  \subfloat[][]{
+%    \label{fig:simple}
+%    \includegraphics[width=0.24\columnwidth]{simple.png} }
+  \subfloat[][]{
+    \label{fig:glines}
+    \includegraphics[width=0.24\columnwidth]{glines.png} }
+
+  \subfloat[][]{
+    \label{fig:ut1-samples}
+    \includegraphics[width=0.24\columnwidth]{ut1-samples.png} }
+  \subfloat[][1x zoom]{
+    \label{fig:ut1}
+    \includegraphics[width=0.24\columnwidth]{ut1.png} }
+  \subfloat[][16x zoom]{
+    \label{fig:ut2}
+    \includegraphics[width=0.24\columnwidth]{ut2.png} }
+
+  \subfloat[][256x zoom]{
+    \label{fig:ut3}
+    \includegraphics[width=0.24\columnwidth]{ut3.png} }
+  \subfloat[][4,000x zoom]{
+    \label{fig:ut4}
+    \includegraphics[width=0.24\columnwidth]{ut4.png} }
+  \subfloat[][60,000x zoom]{
+    \label{fig:ut5}
+    \includegraphics[width=0.24\columnwidth]{ut5.png} }
+
+  \caption{
+    \protect\subref{fig:simple-conflicts} A toy dataset showing conflict cells after building the quadtree from object vertices.
+    \protect\subref{fig:simple-samples} The toy dataset showing how samples are collected.
+%    \protect\subref{fig:simple} def
+    \protect\subref{fig:glines} A maze test dataset.
+    \protect\subref{fig:ut1-samples} A complex dataset with 470 objects at vastly different scales in object size and spacing.
+    \protect\subref{fig:ut1}-\protect\subref{fig:ut5} Complex dataset at different zoom levels up to 60K magnification. This shows the importance of an adaptive method such as a quadtree.
+  }
+  \label{fig:results}
+\end{figure}
 
@@
   \footnotesize{
-  \begin{tabular}{l c c c c c c c}
+%  \begin{tabular}{lcc|cc|cc|cc|cc}
+  \begin{tabular}{lcc|cc|cc|cc}
     \toprule
-    dataset & objects & object          & quadtree   & quadtree          & quadtree &
-    GVD   & GVD             \\
-            &         & $\Delta$s       & depth    & cells           & memory &
-    (sec) & $\Delta$s       \\
-            &         & ($\times 10^3$) &          & ($\times 10^3$) & (Mb)   &
-          & ($\times 10^3$) \\
-    \midrule
-    % ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
-    Fig. \ref{fig:gears} & 3 & 7 & 8 & 54 & 3 & 0.9 & 83\\
-    % ./viewer3 -l 12 ~/data/knife/knife-holder.obj ~/data/knife/knife[2-4].obj
-    Fig. \ref{fig:knife} & 4 & 15 & 12 & 146 & 9 & 3.9 & 232 \\
-    % ./viewer2 -l 24 ../data2/filled_ut/*.dat
-    Fig. \ref{fig:path} & 470 & 5 & 24 & 158 & 8 & 2.0 & 151 \\
-    % ./viewer3 --uniform-colors -a 1 -l 8 ~/data/neuron/improved/a*.obj
-    Fig. \ref{fig:axons} & 448 & 4015 & 8 & 2716 & 151 & 195 & 8100 \\
-    % ./viewer3 -l 8 ~/data/rice-dwarf/decimated2/n1UF2a-*.obj
-    Fig. \ref{fig:mol-explode} & 35 & 1500 & 8 & 496 & 70 & 19 & 2700 \\
+    dataset & objects & object          & \multicolumn{2}{c|}{quadtree}
+    & \multicolumn{2}{c|}{time} &  \multicolumn{2}{c}{quad cells} 
+    \\
+%    & \multicolumn{2}{c}{quad memory}\\
+            &         & facets          & \multicolumn{2}{c|}{depth}
+    & \multicolumn{2}{c|}{(millisec)} & \multicolumn{2}{c}{($\times 10^3$)}
+    \\
+%    & \multicolumn{2}{c}{(Mb)}    \\
+%            &         & ($\times 10^3$) & Ours & Prev
+            &         &  & Ours & Prev
+%    & Ours & Prev & Ours & Prev & Ours & Prev \\
+    & Ours & Prev & Ours & Prev \\
+    %\midrule
+    \hline
+
+    % ./gvd-viewer2 -l 9 ../data2/test1*.dat
+    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/test1*.dat
+    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 10 & 9 & 54 & 3 & 177 & 1168  \\
+
+    % ./gvd-viewer2 -l 24 ../data2/maze/*.dat
+    % ./QUADTREE --disable-conflict-color -d 1 -m 3 -l 24 data/maze/poly*.dat
+    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 24 & 128 & 465 & 38 & 157  \\
+
+    % ./gvd-viewer2 -l 8 ../data2/glines/100x100/*.dat
+    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 9 data/glines/100x100/*
+    Fig. \ref{fig:glines} & 2 & 27,998 & 9 & 8 & 148 & 429 & 43 & 66  \\
+
+    % ./gvd-viewer2 -l 9 ../data2/glines/200x200/*.dat
+    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/glines/200x200/*
+    Fig. \ref{fig:glines} x2 & 2 & 113,084 & 10 & 9 & 414 & 1778 & 125 & 262  \\
+
     \bottomrule
   \end{tabular}}
-  \caption{Table of quadtree/GVD computation statistics and timings on datasets that are unmanageable using other methods.
-\red{Columns are: \emph{objects} - the number of objects in the dataset; \emph{object $\Delta$s} - the number of line segments (2D) or triangles (3D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{quadtree cells} - total number of leaf quadtree cells; \emph{quadtree memory} - amount of memory used by the quadtree; \emph{GVD (sec)} - seconds to perform all steps of GVD computation; \emph{GVD $\Delta$s} - number of line segments (2D) or triangles (3D) in the GVD.}}
+  \caption{Table of quadtree computation statistics and timings on datasets that are unmanageable using other methods.
+Columns are: \emph{objects} - the number of objects in the dataset; \emph{object facets} - the number of line segments (2D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{time (ms)} - milliseconds to build the quadtree; \emph{quad cells} - number of quadtree cells.
+Dataset ``\ref{fig:glines} x2'' is a maze dataset increased in size by a factor of two in each dimension from \ref{fig:glines}.
+}
   \label{tab:timings}
 \end{table*}
+
+As can be seen in table \ref{tab:timings}, there is overhead with our approach: running our algorithm on small datasets yields smaller gains.
+In fact, our approach actually performs worse on the toy dataset.
+The power of our algorithm becomes more obvious on large, complex datasets, where our performance time gains are significant.
+
+We are in the process of integrating our algorithm with animated systems, generating quadtrees in real-time for collision detection, distance transforms, and generalized Voronoi diagram computation.
+Our implementation continues to be refined and optimized, and we expect to shortly have a version with an order of magnitude improvement over the state of the art.
 
@@
 %-------------------------------------------------------------------------------
-\section{Conclusions}
+%\section{Conclusions}
 

### 2017-03-28 paper-orig.tex
converted to Elsevier format

--- 2016-11-13 paper.tex
@@
 \begin{abstract}
-We present a parallel quadtree algorithm that resolves between geometric objects.
+%We present a parallel quadtree algorithm that resolves between geometric objects.
 Our quadtree has the property that no cell intersects more than one labeled object.
@@
 We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D.
-Our results show an order of magnitude improvement over the current state of the art.
+Our results show significant improvement over the current state of the art.
+We present a parallel quadtree algorithm that resolves between geometric objects, modeling space between objects rather than the objects themselves.
+Our quadtree has the property that no cell intersects more than one labeled object.
+A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori or because the number of cells required to resolve closely spaced objects exceeds available memory.
+Previous parallel algorithms that are spatially adaptive, or that discretize finely in only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation.
+Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
+We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D.
+Our results show significant improvement over the current state of the art.
 
@@
 \subsection{Resolve conflict cells}
+\label{sec:resolve}
 
@@
 Our implementation continues to be refined and optimized, and we expect to shortly have a version with an order of magnitude improvement over the state of the art.
+Importantly, we are also working on an extension to 3D.
+Every step in our method has a straightforward extension to 3D with the exception of point sampling for conflict resolution (see section \ref{sec:resolve}), which is where we are focusing our efforts.
 

### 2017-04-01 paper.tex
First major text revision addressing reviewers' comments and adding figures

@@
-\documentclass[submission]{gmp2017}
+\documentclass[final,3p,times,twocolumn]{elsarticle}
+%\documentclass[submission]{gmp2017}
 %\documentclass{gmp2017}
-
-\title{Parallel Quadtree Construction on Collections of Objects}
 
@@
 % class option "submission" is active.
-\SubNumber{19}
+%\SubNumber{19}
 
@@
 % is active.
-\author{Nathan Vollmer}{1}
-\author{John Edwards}{1}
+%\author{Nathan Vollmer}{1}
+%\author{John Edwards}{1}
 
 % Put the affiliations of the authors here.
-\affiliation{1}{Department of Informatics and Computer Science, Idaho State University, USA}
+%\affiliation{1}{Department of Informatics and Computer Science, Idaho State University, USA}
+
+%% The numcompress package shorten the last page in references.
+%% `nodots' option removes dots from firstnames in references.
+\usepackage[nodots]{numcompress}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers after \end{frontmatter}.
+\usepackage{lineno}
+
+%% Avoids linenumbers to collide with text for 5p format:
+\setlength\linenumbersep{3pt}
 
@@
 \begin{document}
+
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%%
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+%\title{Parallel Quadtree Construction on Collections of Objects}
+\title{Fast Construction of Inter-Object Spacing Representations}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
+
+\author{}
+
+\address{}
 
@@
 %
-\maketitle
+%\maketitle
 
 \begin{abstract}
-We present a parallel quadtree algorithm that resolves between geometric objects.
+%We present a parallel quadtree algorithm that resolves between geometric objects.
 Our quadtree has the property that no cell intersects more than one labeled object.
@@
 We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D.
-Our results show an order of magnitude improvement over the current state of the art.
+Our results show significant improvement over the current state of the art.
+We present a parallel quadtree algorithm that resolves between geometric objects, modeling space between objects rather than the objects themselves.
+Our quadtree has the property that no cell intersects more than one labeled object.
+A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori or because the number of cells required to resolve closely spaced objects exceeds available memory.
+Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation.
+Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
+We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D.
+Our results show significant improvement over the current state of the art.
 
@@
 
+\end{frontmatter}
+
+%%
+%% Start line numbering here if you want
+%%
+\linenumbers
+
 %-------------------------------------------------------------------------
@@
 \label{sec:intro}
-Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, robot navigation, object description, and other applications.
+Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, robot navigation, shape modeling, object description, and other applications.
 Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects.
 However, our work centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object.
-Using a quadtree, we can model the space between objects, the first step in constructing distance fields, detecting collisions, and computing the generalized Voronoi diagram.
+Using a quadtree, we can model the space between objects.
+Our quadtree representation is not purposed for fast retrieval as is often the case in hierarchical subdivisions of point data.
+Rather, the quadtrees are used for constructing distance fields, detecting collisions and computing the GVD \cite{edwards2015approximating}.
+
+\red{I just added this paragraph.
+What do you think?} Object separation, such as is treated here, is of some use in 2D (e.g.
+path planning), but becomes a very important problem in 3D.
+Hierarchically subdividing non-point data in a principled parallel way is surprisingly complex, and this paper lays the groundwork for our continuing work in 3D.
+
 Modeling inter-object spacing is computationally straightforward when inter-object spacing is large compared to the world bounding box.
@@
 
-Difficulties arise when spacing between objects is small relative to the size of the domain.
+Difficulties arise when objects are close together relative to the size of the domain.
 An approach using a uniform grid would have excessive memory requirements in order to resolve between objects because the uniformly sized grid cell must be small enough to fit between objects at every location in the domain.
 Thus, an adaptive approach must be used for datasets of closely spaced objects.
-To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory but does so in serial, with expected performance liabilities.
+To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory, but it does so in serial, with expected performance liabilities.
+A naive approach to parallelizing quadtree computation would be to assign all available compute units according to a course grid, then run the serial algorithm on each compute unit.
+While simple, there is potential for serious load imbalancing if the close object spacings are not uniformly distributed.
 
-This work extends the work done by Edwards et al \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm targeted for the GPU.
-Our algorithm performs an order of magnitude faster than the previous work and will be an excellent base for later distance transform and generalized Voronoi diagram computation.
+This work extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution.
+Our algorithm, which is targeted for the GPU, performs an order of magnitude faster than the previous work and will be an important base for later distance transform and generalized Voronoi diagram computation.
 
@@
 \begin{enumerate}
-\item Construct an quadtree on object vertices using Karras' algorithm \cite{karras2012maximizing}
+\item Construct a quadtree on object vertices using the Karras algorithm \cite{karras2012maximizing}
 \item Detect quadtree cells that intersect more than one object, which we call ``conflict cells'' (contribution)
@@
 Two seminal works build octrees on objects in order to compute the Adaptive Distance Field (ADF) on octree vertices.
-Strain~\shortcite{strain1999fast} fully resolves the quadtree everywhere on the object surface, and Frisken \etal~\shortcite{frisken2000adaptively} resolve the quadtree fully only in areas of small local feature size.
+Strain \shortcite{strain1999fast} fully resolves the quadtree everywhere on the object surface, and Frisken \etal \shortcite{frisken2000adaptively} resolve the quadtree fully only in areas of small local feature size.
 Both approaches are designed to retain features of a single object rather than resolving between multiple objects, as is required for GVD computation.
@@
 
-Most of our algorithm is independent of dimension, and so we will use terms ``facet'' and ``octree'' (which lacks a dimension-independent term).
+%Most of our algorithm is independent of dimension, and so we will use terms ``facet'' and ``octree'' (which lacks a dimension-independent term).
 The process of finding sample points to resolve conflict cells is limited to 2D in this paper, so for that section we will use ``line'' and ``quadtree''.
+Each step of our algorithm, with the exception of resolving conflict cells, is independent of dimension and can be used for 3D octree applications.
+But since point sampling for conflict cell resolution is 2D we will use the term quadtree throught the algorithm description for consistency.
+Our algorithm assumes the objects are faceted where the facets are simplices.
 
@@
 \label{sec:build-initial-octree}
-Our first step is to build an octree on the vertices.
+Our first step is to build an octree on the given set of vertices.
+We use the Karras algorithm \cite{karras2012maximizing} which starts by sorting the Morton codes of the given vertices.
+Our implementation uses an efficient parallel radix sorter described by Ha \etal \shortcite{ha2009fast}.
+Once the vertices are sorted, a binary radix tree, and then an initial octree can be constructed in parallel.
+The strength of this approach lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points.
+That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.
+
+%Our first step is to build an octree on the vertices.
 This step gives us our first approximation to our final octree.
@@
 
+\subsection{Pruning the octree}
+Assume we have a vertex labeling, such that each vertex is labeled with the object it belongs to.
+The initial octree resolves between vertices, regardless of whether they are labeled differently from each other or not, so each leaf node is guaranteed to contain no more than one vertex.
+Since our objective is to resolve between objects of different labels, we can proactively prune the initial vertex octree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label.
+\red{Clarify this passage: We remove internal binary radix tree nodes seperating identically labeled vertices.
+For each binary radix tree leaf in parallel, we first look at the label of the leaf's corresponding vertices.
+If the colors match, we apply that color to the leaf.
+Otherwise, we mark that leaf as "required".} This initial coloring can be done immediately after the Karras binary radix tree construction without the need to invoke an additional kernel.
+Before generating the binary radix tree's corresponding quadtree, we propagate the binary radix leaf colors up the tree in parallel, marking internal nodes as "required" when the colors of two children nodes mismatch.
+Finally, we generate quadtree nodes from only the required internal binary radix tree nodes.
+By pruning the initial octree, we simplify conflict detection and reduce our memory footprint.
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:pruning-1}
+    \includegraphics[width=0.35\columnwidth]{pruned-vs-unpruned/4.PNG} }
+  \subfloat[][]{
+    \label{fig:pruning-2}
+    \includegraphics[width=0.35\columnwidth]{pruned-vs-unpruned/5.PNG} }
+  \caption{
+    \protect\subref{fig:pruning-1} The initial quadtree built on the object vertices, in which no quadtree cell contains more than one vertex, can be far more complex than needed to resolve between objects.
+    \protect\subref{fig:pruning-2} After pruning the quadtree. Quadtree cells can contain multiple vertices as long as they all have the same label.
+  }
+  \label{fig:pruning}
+\end{figure}
+
 \subsection{Detect conflict cells}
 
-Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from Root to $C$, where the local address is the Morton code of a $2^{DIM}$ block of voxels.
+Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from Root to $C$, where the local address is a 2-bit (3-bit in 3D) Morton code.
 The address of the root cell is defined as the empty string.
@@
 
-Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (Figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.
+Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.
 
@@
     \node[anchor=south west,inner sep=0,outer sep=0] at (0,0) {
-      \subfloat[][]{
-        \label{fig:conflict-find-1}
-        \includegraphics[width=0.22\columnwidth]{conflict-find-1.pdf} } \hfill
+%      \subfloat[][]{
+%        \label{fig:conflict-find-1}
+%        \includegraphics[width=0.44\columnwidth]{conflict-find-1.pdf} } \hfill
       \subfloat[][]{
         \label{fig:conflict-find-2}
-        \includegraphics[width=0.22\columnwidth]{conflict-find-2.pdf} } \hfill
+        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
       \subfloat[][]{
         \label{fig:conflict-find-4}
-        \includegraphics[width=0.22\columnwidth]{conflict-find-4.pdf} } \hfill
+        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
+%      \subfloat[][]{
+%        \label{fig:conflict-find-5}
+%        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
+%      \subfloat[][]{
+%        \label{fig:conflict-find-6}
+%        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }
+    };
+
+    \node[anchor=south west,inner sep=0,outer sep=0] at (0,-4) {
+%      \subfloat[][]{
+%        \label{fig:conflict-find-2}
+%        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
+%      \subfloat[][]{
+%        \label{fig:conflict-find-4}
+%        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
       \subfloat[][]{
         \label{fig:conflict-find-5}
-        \includegraphics[width=0.22\columnwidth]{conflict-find-5.pdf} }
-    };
-    %\draw[red,ultra thick] (4.0,1.7) rectangle (6.55,4.2);
-    \draw[red,thick] (4.0,1.7) rectangle (6.55,4.2);
-
-    % lines
-    % upper-right
-    \draw[red,thin] (6.55,4.2) -- (11.8,4.2);
-    % lower-left
-    \draw[red,thin] (4.0,1.7) -- (8.0,0.3);
-
-    %% curves
-    %% upper-left
-    %%\draw[red,thin] (4.0,4.2) to[out=10,in=170] (8.0,4.2);
-    %% upper-right
-    %\draw[red,thin] (6.55,4.2) to[out=10,in=170] (11.8,4.2);
-    %% lower-left
-    %\draw[red,thin] (4.0,1.7) to[out=-40,in=175] (8.0,0.3);
-
-    \draw[red,thick] (8.0,0.3) rectangle (11.8,4.2);
-  \end{tikzpicture} \\
+        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
       \subfloat[][]{
         \label{fig:conflict-find-6}
-        \includegraphics[width=0.22\columnwidth]{conflict-find-6.pdf} } \hfill
+        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }
+    };
+
+    %% single column
+    %% small rectangle
+    %\draw[red,thick] (4.0,1.7) rectangle (6.55,4.2);
+    %% upper-right line
+    %\draw[red,thin] (6.55,4.2) -- (11.8,4.2);
+    %% lower-left line
+    %\draw[red,thin] (4.0,1.7) -- (8.0,0.3);
+    %% large rectangle
+    %\draw[red,thick] (8.0,0.3) rectangle (11.8,4.2);
+
+%    % spanning column, 0.44 size
+%    % small rectangle
+%    \draw[red,thick] (1.0,1.7) rectangle (6.55,4.2);
+%    % upper-right line
+%    \draw[red,thin] (6.55,4.2) -- (11.8,4.2);
+%    % lower-left line
+%    \draw[red,thin] (4.0,1.7) -- (8.0,0.3);
+%    % large rectangle
+%    \draw[red,thick] (8.0,0.3) rectangle (11.8,4.2);
+
+%    % two column, .22 size
+%    % small rectangle
+%    \draw[red,thick] (2.0,1.05) rectangle (3.27,2.25);
+%    % upper-right line
+%    \draw[red,thin] (3.27,2.25) -- (3.9,2.25);
+%    % lower-left line
+%    \draw[red,thin] (2.0,1.05) -- (3.9,0.4);
+%    % large rectangle
+%    \draw[red,thick] (3.9,0.4) rectangle (5.8,2.25);
+
+    % two column, .4 size
+    % small rectangle
+    \draw[red,thick] (0.0,1.65) rectangle (2.33,4.0);
+%    % upper-right line
+%    \draw[red,thin] (3.27,2.25) -- (3.9,2.25);
+%    % lower-left line
+%    \draw[red,thin] (2.0,1.05) -- (3.9,0.4);
+%    % large rectangle
+%    \draw[red,thick] (3.9,0.4) rectangle (5.8,2.25);
+  \end{tikzpicture}
+%      \subfloat[][]{
+%        \label{fig:data-structures-1}
+%        \includegraphics[width=0.38\columnwidth]{data-structures-1.pdf} } \hfill
+%      \subfloat[][]{
+%        \label{fig:data-structures-2}
+%        \includegraphics[width=0.38\columnwidth]{data-structures-3.pdf} } \hfill
+%      \subfloat[][]{
+%        \label{fig:data-structures-3}
+%        \includegraphics[width=0.38\columnwidth]{data-structures-4.pdf} }
+  \caption{
+%    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and green with facets labeled A-I.
+    We have three objects, blue, red, and green with facets labeled A-I.
+    \protect\subref{fig:conflict-find-2} Initial vertex quadtree.
+    \protect\subref{fig:conflict-find-4} Zoomed-in to the region outlined by red in \protect\subref{fig:conflict-find-2} showing the boundary cell (BCell) computation for each facet.
+%These pairs are given in figure \protect\subref{fig:data-structures-1}.
+    \protect\subref{fig:conflict-find-5} Conflict cells, which intersect more than one object, are highlighted.
+    \protect\subref{fig:conflict-find-6} The new quadtree after conflict resolution.
+%    \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity).
+The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure).
+%    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access.
+We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given octree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
+%\protect\subref{fig:data-structures-3} For a given octree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
+  }
+  \label{fig:conflict-find}
+\end{figure}
+
+\begin{figure}
+  \centering
       \subfloat[][]{
         \label{fig:data-structures-1}
-        \includegraphics[width=0.22\columnwidth]{data-structures-1.pdf} } \hfill
+        \includegraphics[width=0.32\columnwidth]{data-structures-1.pdf} }
       \subfloat[][]{
         \label{fig:data-structures-2}
-        \includegraphics[width=0.22\columnwidth]{data-structures-3.pdf} } \hfill
+        \includegraphics[width=0.32\columnwidth]{data-structures-3.pdf} }
       \subfloat[][]{
         \label{fig:data-structures-3}
-        \includegraphics[width=0.22\columnwidth]{data-structures-4.pdf} }
+        \includegraphics[width=0.32\columnwidth]{data-structures-4.pdf} }
   \caption{
-    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and green with facets labeled A-I.
-    \protect\subref{fig:conflict-find-2} We construct an initial quadtree on the vertices using Karras' algorithm.
-    \protect\subref{fig:conflict-find-4} Zoomed-in view of the boundary cell (BCell) computation for each facet. These pairs are given in figure \protect\subref{fig:data-structures-1}.
-    \protect\subref{fig:conflict-find-5} Conflict cells, which intersect more than one object, are highlighted.
-    \protect\subref{fig:conflict-find-6} The new quadtree after conflict resolution.
     \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity).
@@
   }
-  \label{fig:conflict-find}
-\end{figure}
-
-\begin{figure}
-  \label{fig:scc-sort}
+  \label{fig:data-structures}
 \end{figure}
@@
     \label{fig:conflict-resolution-x}
-    \includegraphics[width=0.24\columnwidth]{conflict-resolution-x.pdf} }
+    \includegraphics[width=0.35\columnwidth]{conflict-resolution-x.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-even}
-    \includegraphics[width=0.24\columnwidth]{conflict-resolution-even.pdf} }
+    \includegraphics[width=0.35\columnwidth]{conflict-resolution-even.pdf} }\\
   \subfloat[][]{
     \label{fig:conflict-resolution-all}
-    \includegraphics[width=0.24\columnwidth]{conflict-resolution-all.pdf} }
+    \includegraphics[width=0.35\columnwidth]{conflict-resolution-all.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-octree}
-    \includegraphics[width=0.24\columnwidth]{conflict-resolution-octree.pdf} } \\
+    \includegraphics[width=0.35\columnwidth]{conflict-resolution-octree.pdf} }\\
   \subfloat[][]{
     \label{fig:conflict-resolution-adjacent-even}
-    \includegraphics[width=0.24\columnwidth]{conflict-resolution-adjacent-even.pdf} }
+    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-even.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-adjacent-all}
-    \includegraphics[width=0.24\columnwidth]{conflict-resolution-adjacent-all.pdf} }
+    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-all.pdf} }
   \caption{
@@
     \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least two samples.
-This ensures that an quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-octree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
+This ensures that a quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-octree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
   }
@@
 We also derive equation \eqref{eqn:sample} which computes the samples themselves.
-The power of our approach lies in the fact that both expressions are closed-form and neither one is iterative, so we can evaluate the first in parallel over leaf cells and the second in parallel over all samples to compute.
+The power of our approach lies in the fact that both expressions are closed-form and neither one is iterative, so we can evaluate the first in parallel over leaf cells and the second in parallel over all samples that we need to compute.
 
@@
 that bisects $q$ and $r$.
-Our strategy will be to sample points $P$ on $p(s)$ (figure \ref{fig:conflict-resolution-octree}) such that an quadtree built on $S \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
+Our strategy will be to sample points $P$ on $p(s)$ (figure \ref{fig:conflict-resolution-octree}) such that a quadtree built on $S \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
 We do this by ensuring that $P$ is sampled such that every box that intersects both $q$ and $r$ also intersects at least two points in $P$.
@@
 We will find a series of boxes such that each box's left-most intersection with $p(s)$ is a sample point meeting the above criterion.
+In the following discussion, $p^x$ and $p^y$ refer to the $x$ and $y$ coordinates of point $p$, respectively.
 
@@
 \begin{tightenumerate}
- \item The \textit{opposite} case (see Figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
- \item In the \textit{adjacent} case (see Figure \ref{fig:conflict-resolution-adjacent-even}), $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
+ \item The \textit{opposite} case (see figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
+ \item In the \textit{adjacent} case (see figure \ref{fig:conflict-resolution-adjacent-even}), $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
 \end{tightenumerate}
@@
 We now use $a(s)$ to construct a sequence of values $S = \{s_0, s_1, s_2, \dots, s_n\}$ that meet our sampling criterion.
-We first construct the even samples (see Figures \ref{fig:conflict-resolution-even} and \ref{fig:conflict-resolution-adjacent-even}).
+We first construct the even samples (see figures \ref{fig:conflict-resolution-even} and \ref{fig:conflict-resolution-adjacent-even}).
 Given a starting point $p(s_0)$,
@@
 
-\subsubsection{Iterate}
+\subsubsection{Iteration}
 \label{sec:iterate}
@@
 If new sample points were found then we add them to the current set $S$ of sample points and return to building the octree from points (section \ref{sec:build-initial-octree}).
-We finish when the only conflicts identified are at the maximum depth, i.e., unresolvable.
+We finish when the only conflicts identified are at the maximum depth.
 
@@
   \end{tightenumerate}
-\item Iterate - $O(Q)$ iterations. In the worst case, all facets intersect a single cell, requiring potentially $Q=O(M^2)$ iterations. In our testing, $Q$ has not exceeded 3.
+\item Iterate - $O(Q)$ iterations. In the worst case, all facets intersect a single cell, requiring potentially $Q=O(M^2)$ iterations. In our testing, $Q$ has not exceeded 4.
 \end{tightenumerate}
@@
 of the algorithm supports polygons and polylines which needn't be manifold or connected.
-All tests were run on a MacBook Pro laptop with a dual-core 2.9 GHz processor, 16 GB memory, and Intel Iris Graphics 6100 graphics card.
-Figure \ref{fig:results} shows results on four datasets: a simple toy dataset showing conflict cell detection and resolution (\ref{fig:simple-conflicts}-\ref{fig:simple-samples}); a more complex maze dataset (\ref{fig:glines}), and a complex dataset with many objects at very different scales (\ref{fig:ut1-samples}-\ref{fig:ut5}).
+All tests were run on a \red{need specs of your computer here.
+something like ``MacBook Pro laptop with a dual-core 2.9 GHz processor, 16 GB memory, and Intel Iris Graphics 6100 graphics card''.} Figure \ref{fig:results} shows results on four datasets: a simple toy dataset showing conflict cell detection and resolution (\ref{fig:simple-conflicts}-\ref{fig:simple-samples}); a more complex maze dataset (\ref{fig:glines}), and a complex dataset with many objects at very different scales (\ref{fig:ut1-samples}-\ref{fig:ut5}).
 Table \ref{tab:timings} shows timings for our implementation compared to the previous state-of-the-art.
@@
     \label{fig:simple-conflicts}
-    \includegraphics[width=0.24\columnwidth]{simple-conflict.png} }
+    \includegraphics[width=0.31\columnwidth]{simple-conflict.png} }
   \subfloat[][]{
     \label{fig:simple-samples}
-    \includegraphics[width=0.24\columnwidth]{simple-samples.png} }
-%  \subfloat[][]{
-%    \label{fig:simple}
-%    \includegraphics[width=0.24\columnwidth]{simple.png} }
-  \subfloat[][]{
-    \label{fig:glines}
-    \includegraphics[width=0.24\columnwidth]{glines.png} }
-
-  \subfloat[][]{
-    \label{fig:ut1-samples}
-    \includegraphics[width=0.24\columnwidth]{ut1-samples.png} }
-  \subfloat[][1x zoom]{
-    \label{fig:ut1}
-    \includegraphics[width=0.24\columnwidth]{ut1.png} }
-  \subfloat[][16x zoom]{
-    \label{fig:ut2}
-    \includegraphics[width=0.24\columnwidth]{ut2.png} }
-
-  \subfloat[][256x zoom]{
-    \label{fig:ut3}
-    \includegraphics[width=0.24\columnwidth]{ut3.png} }
-  \subfloat[][4,000x zoom]{
-    \label{fig:ut4}
-    \includegraphics[width=0.24\columnwidth]{ut4.png} }
-  \subfloat[][60,000x zoom]{
-    \label{fig:ut5}
-    \includegraphics[width=0.24\columnwidth]{ut5.png} }
+    \includegraphics[width=0.31\columnwidth]{simple-samples.png} }
 
@@
     \protect\subref{fig:simple-samples} The toy dataset showing how samples are collected.
-%    \protect\subref{fig:simple} def
-    \protect\subref{fig:glines} A maze test dataset.
+  }
+  \label{fig:results-toy}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:vascular-1}
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/04-Pruned.PNG} }
+  \subfloat[][]{
+    \label{fig:vascular-2}
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/06-Initial-Conflicts.PNG} }
+  \subfloat[][]{
+    \label{fig:vascular-3}
+%    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/08-Resolved-Quadtree.PNG} }
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/13-Resolved-Quadtree.PNG} }
+  \caption{\red{Can you give some details on this dataset -- what it is etc}
+    \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.
+    \protect\subref{fig:vascular-2} All conflict cells of the initial quadtree.
+    \protect\subref{fig:vascular-3} After conflict cell resolution.
+No quadtree cell intersects more than one object.
+\red{Can you get a similar figure to this that has closed polygons and is maybe zoomed in a little more two a few nice-looking polygons?}
+  }
+  \label{fig:vascular}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:gears-1}
+    \includegraphics[width=0.31\columnwidth]{Nathan/gears/2.PNG} }
+  \subfloat[][]{
+    \label{fig:gears-2}
+    \includegraphics[width=0.31\columnwidth]{Nathan/gears/3.PNG} }
+  \subfloat[][]{
+    \label{fig:gears-3}
+    \includegraphics[width=0.31\columnwidth]{Nathan/gears/5.PNG} }
+  \caption{
+    \protect\subref{fig:gears-1} A dataset of gears with close tolerance. The resolved quadtree with sampled points is shown.
+    \protect\subref{fig:gears-2} Showing just the quadtree and sample points.
+    \protect\subref{fig:gears-3} A zoomed-in image showing the close object spacing compared to the large domain.
+  }
+  \label{fig:gears}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:ut1-samples}
+    \includegraphics[width=0.31\columnwidth]{ut1-samples.png} }
+  \subfloat[][1x zoom]{
+    \label{fig:ut1}
+    \includegraphics[width=0.31\columnwidth]{ut1.png} }
+  \subfloat[][16x zoom]{
+    \label{fig:ut2}
+    \includegraphics[width=0.31\columnwidth]{ut2.png} } \\
+  \subfloat[][256x zoom]{
+    \label{fig:ut3}
+    \includegraphics[width=0.31\columnwidth]{ut3.png} }
+  \subfloat[][4,000x zoom]{
+    \label{fig:ut4}
+    \includegraphics[width=0.31\columnwidth]{ut4.png} }
+  \subfloat[][60,000x zoom]{
+    \label{fig:ut5}
+    \includegraphics[width=0.31\columnwidth]{ut5.png} }
+  \caption{
     \protect\subref{fig:ut1-samples} A complex dataset with 470 objects at vastly different scales in object size and spacing.
@@
   }
-  \label{fig:results}
+  \label{fig:results-maze}
 \end{figure}
@@
   \end{tabular}}
-  \caption{Table of quadtree computation statistics and timings on datasets that are unmanageable using other methods.
+  \caption{\red{Nate, please send me an updated table in whatever format (I can throw it into Latex).
+We should have timings on simple, maze, vascular and gears comparing GVD to PGVD with and without pruning.
+Not sure if we'll end up including the non-pruning results.
+We should also include the \# objects, \# facets and \# quadtree cells, as are included in this table.} Table of quadtree computation statistics and timings on datasets that are unmanageable using other methods.
 Columns are: \emph{objects} - the number of objects in the dataset; \emph{object facets} - the number of line segments (2D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{time (ms)} - milliseconds to build the quadtree; \emph{quad cells} - number of quadtree cells.
@@
 
-\bibliographystyle{abbrv}
-\balance
+\bibliographystyle{model3-num-names}
+%\bibliographystyle{abbrv}
+%\balance
 \bibliography{paper}

### 2017-04-02 paper.tex
Modified small thing in abstract, introduction, clarified prune section, corrected detect conflict cells, added pruning complexity analysis

@@
 \documentclass[final,3p,times,twocolumn]{elsarticle}
-%\documentclass[submission]{gmp2017}
-%\documentclass{gmp2017}
-
-% Put your submission number here.
-% This number will appear instead of authors and affiliations if the
-% class option "submission" is active.
-%\SubNumber{19}
-
-% Put the author names here.
-% Use the second argument as a reference to the list of affiliations.
-% No authors and affiliations will appear if the class option "submission"
-% is active.
-%\author{Nathan Vollmer}{1}
-%\author{John Edwards}{1}
-
-% Put the affiliations of the authors here.
-%\affiliation{1}{Department of Informatics and Computer Science, Idaho State University, USA}
 
@@
 \input{common}
-
-%\usepackage[inline]{showlabels}
 
@@
 
-%\title{Parallel Quadtree Construction on Collections of Objects}
 \title{Fast Construction of Inter-Object Spacing Representations}
-
+\author{Nathan Morrical}
 %% use optional labels to link authors explicitly to addresses:
@@
 
-\author{}
+\author{John Edwards}
 
-\address{}
+\address{Department of Informatics and Computer Science, Idaho State University, USA}
 
@@
 \begin{abstract}
-%We present a parallel quadtree algorithm that resolves between geometric objects.
-Our quadtree has the property that no cell intersects more than one labeled object.
-A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori, because the number of cells required to resolve closely spaced objects exceeds available memory.
-Previous parallel algorithms that are spatially adaptive, or that discretize finely in only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation.
-Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
-We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D.
-Our results show significant improvement over the current state of the art.
 We present a parallel quadtree algorithm that resolves between geometric objects, modeling space between objects rather than the objects themselves.
@@
 A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori or because the number of cells required to resolve closely spaced objects exceeds available memory.
-Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation.
+Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, \red{Hierarchical kernel envokation isn't necessarily bad.
+We might move to that to remove some of the Q linear complexity overhead.} either separate points only, or make no guarantees of object separation.
 Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
@@
 \label{sec:intro}
-Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, robot navigation, shape modeling, object description, and other applications.
+\red{Reworded this paragraph, since the last sentence contained duplicate information provided in the first sentence.} Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, generalized Voronoi Diagram (GVD) construction \cite{edwards2015approximating}, robot navigation, shape modeling, object description, and other applications.
 Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects.
+In other cases, quadtrees are purposed for fast retrieval as is often the case in hierarchical subdivisions of point data.
 However, our work centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object.
-Using a quadtree, we can model the space between objects.
-Our quadtree representation is not purposed for fast retrieval as is often the case in hierarchical subdivisions of point data.
-Rather, the quadtrees are used for constructing distance fields, detecting collisions and computing the GVD \cite{edwards2015approximating}.
 
-\red{I just added this paragraph.
-What do you think?} Object separation, such as is treated here, is of some use in 2D (e.g.
+\red{reworded the first sentence a bit.} This object separation is of some use in 2D (e.g.
 path planning), but becomes a very important problem in 3D.
@@
 
-Modeling inter-object spacing is computationally straightforward when inter-object spacing is large compared to the world bounding box.
+Modeling inter-object spacing is computationally straightforward when that spacing is large compared to the world bounding box.
 Approaches typically involve a uniform grid of the space, which leads to efficient computation that often uses graphics processors.
@@
 Thus, an adaptive approach must be used for datasets of closely spaced objects.
+\red{added carriage return here}
+
 To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory, but it does so in serial, with expected performance liabilities.
@@
 
-This work extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution.
+This paper extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution.
 Our algorithm, which is targeted for the GPU, performs an order of magnitude faster than the previous work and will be an important base for later distance transform and generalized Voronoi diagram computation.
@@
 
-\subsection{Build initial octree}
-\label{sec:build-initial-octree}
-Our first step is to build an octree on the given set of vertices.
+\subsection{Build initial quadtree}
+\label{sec:build-initial-quadtree}
+Our first step is to build a quadtree on the given set of vertices.
 We use the Karras algorithm \cite{karras2012maximizing} which starts by sorting the Morton codes of the given vertices.
 Our implementation uses an efficient parallel radix sorter described by Ha \etal \shortcite{ha2009fast}.
-Once the vertices are sorted, a binary radix tree, and then an initial octree can be constructed in parallel.
+Once the vertices are sorted, a binary radix tree, and then an initial quadtree can be constructed in parallel.
 The strength of this approach lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points.
@@
 
-%Our first step is to build an octree on the vertices.
-This step gives us our first approximation to our final octree.
-We use Karras' algorithm \cite{karras2012maximizing} which sorts the Morton codes of the vertices in parallel, then constructs the binary radix tree in parallel.
-With the binary radix tree, the quadtree can be constructed with a single parallel call.
-The strength of this algorithm lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points.
-That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.
+\subsection{Prune the octree}
+During the initial quadtree construction, we can indirectly prune the initial quadtree to simplify conflict detection and reduce our memory footprint.
+Assume we have a numeric vertex labeling such that each vertex is labeled to match the object it belongs to.
+The initial binary radix tree (BRT) provided by Karras serves as a bounding volume hierarchy, and is used to generate the initial quadtree by seperating vertices regardless of their label.
+Since our objective is to resolve between objects of different labels, we can proactively prune the initial BRT, and subsequently the initial quadtree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label.
 
-\subsection{Pruning the octree}
-Assume we have a vertex labeling, such that each vertex is labeled with the object it belongs to.
-The initial octree resolves between vertices, regardless of whether they are labeled differently from each other or not, so each leaf node is guaranteed to contain no more than one vertex.
-Since our objective is to resolve between objects of different labels, we can proactively prune the initial vertex octree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label.
-\red{Clarify this passage: We remove internal binary radix tree nodes seperating identically labeled vertices.
-For each binary radix tree leaf in parallel, we first look at the label of the leaf's corresponding vertices.
-If the colors match, we apply that color to the leaf.
-Otherwise, we mark that leaf as "required".} This initial coloring can be done immediately after the Karras binary radix tree construction without the need to invoke an additional kernel.
-Before generating the binary radix tree's corresponding quadtree, we propagate the binary radix leaf colors up the tree in parallel, marking internal nodes as "required" when the colors of two children nodes mismatch.
+\red{I tried to be a bit more verbose here and in the paragraph above.
+Let me know if anything in particular needs more clarification.} To prune the initial BRT efficiently, we label each BRT node using the following criterion: if the current node is a leaf node which seperates two vertices with identical labels, label the current node to match the label of the vertices being seperated.
+If the current node is a leaf node that seperates two vertices having mismatching colors, label the current node as -2, i.e.
+"required".
+Lastly, if the current node is an internal node, i.e.
+the current node is an ancestor of another BRT node, mark the current node as -1, i.e.
+"unknown''.
+This initial step can be done immediately after Karras' BRT construction without the need to invoke an additional kernel.
+
+We then propagate the BRT labels up the tree in parallel, marking "unknown'' nodes as "required" when the labels of the current node's two children nodes mismatch.
 Finally, we generate quadtree nodes from only the required internal binary radix tree nodes.
-By pruning the initial octree, we simplify conflict detection and reduce our memory footprint.
 
@@
 We define a bounding cell (BCell) to be the smallest internal quadtree node which entirely contains a given facet.
-Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the BCell is the longest common prefix of the Morton codes of the points in $P$.
-Figure \ref{fig:data-structures-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.
+Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the BCell is contained in the longest common prefix (LCP) of the Morton codes of the points in $P$.
+If a given LCP is more specific than any quadtree node, ie the LCP lies within a quadtree leaf, we simply take the quadtree address of the leaf that the LCP lies within.
+This is often the case with pruned quadtrees where entire facets may lie within a quadtree leaf.
+\red{This paragraph had some incorrect conclusions, so I tried to clarify a bit more.} Figure \ref{fig:data-structures-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.
 
@@
     We have three objects, blue, red, and green with facets labeled A-I.
-    \protect\subref{fig:conflict-find-2} Initial vertex quadtree.
+    \protect\subref{fig:conflict-find-2} Initial vertex quadtree. \red{This vertex quadtree is unpruned. Can we prune it for consistency?}
     \protect\subref{fig:conflict-find-4} Zoomed-in to the region outlined by red in \protect\subref{fig:conflict-find-2} showing the boundary cell (BCell) computation for each facet.
@@
 First, we set $L$'s color to $-1$ (uninitialized).
-We then investigate each ancestor $A$ of $L$ (line \ref{alg:A}) by using the \texttt{Parent} field in the octree data structure.
+We then traverse each ancestor $A$ of $L$ (line \ref{alg:A}) by using the \texttt{Parent} field in the octree data structure.
 Using the \texttt{FFacet} and \texttt{LFacet} fields, we find, respectively, the first and (inclusive) last of possibly multiple facets bounded by $A$ (line \ref{alg:i}).
@@
 \item Build octree using Karras' algorithm \cite{karras2012maximizing} - $O(D)$.
+\item Prune the Binary Radix Tree from Karras \cite{karras2012maximizing} - $O(D)$ \red{Does this sound right?}
 \item Detect conflict cells
@@
 This requires looking at each bit, of which there are $O(D)$.
-  \item Sort BCells array - $O(\log{M})$. The array has $M$ elements, and we use a parallel radix sort with $\log$ complexity.
+  \item Sort BCells array - $O(\log{M})$.
+\red{Shouldn't the big o here be O(n), where n is the number of levels in the quadtree?
+} The array has $M$ elements, and we use a parallel radix sort with $\log$ complexity.
   \item Index BCells with octree data structure - $O(D)$. This runs in parallel on leaf cell IDs and each kernel requires a search of the octree for a given cell ID, taking at most $D$ steps.
@@
 of the algorithm supports polygons and polylines which needn't be manifold or connected.
-All tests were run on a \red{need specs of your computer here.
-something like ``MacBook Pro laptop with a dual-core 2.9 GHz processor, 16 GB memory, and Intel Iris Graphics 6100 graphics card''.} Figure \ref{fig:results} shows results on four datasets: a simple toy dataset showing conflict cell detection and resolution (\ref{fig:simple-conflicts}-\ref{fig:simple-samples}); a more complex maze dataset (\ref{fig:glines}), and a complex dataset with many objects at very different scales (\ref{fig:ut1-samples}-\ref{fig:ut5}).
+All tests were run on a Razer Blade Stealth with an Intel i7 6500u 3.10 ghz dual core processor, 8GB of memory, and an Nvidia GTX 1070 graphics card.
+Figure \ref{fig:results} shows results on four datasets: a simple toy dataset showing conflict cell detection and resolution (\ref{fig:simple-conflicts}-\ref{fig:simple-samples}); a more complex maze dataset (\ref{fig:glines}), and a complex dataset with many objects at very different scales (\ref{fig:ut1-samples}-\ref{fig:ut5}).
 Table \ref{tab:timings} shows timings for our implementation compared to the previous state-of-the-art.
@@
   \caption{
-    \protect\subref{fig:simple-conflicts} A toy dataset showing conflict cells after building the quadtree from object vertices.
+    \protect\subref{fig:simple-conflicts} \red{On the bottom right, the box in the 3rd quadrant is not a conflict, and shouldn't be colored red.} A toy dataset showing conflict cells after building the quadtree from object vertices.
     \protect\subref{fig:simple-samples} The toy dataset showing how samples are collected.

### 2017-04-03 paper.tex
updated figure 6

@@
 %    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/08-Resolved-Quadtree.PNG} }
-    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/13-Resolved-Quadtree.PNG} }
-  \caption{\red{Can you give some details on this dataset -- what it is etc}
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/14-Resolved-Quadtree.PNG} }
+  \caption{A large set of uniquely labeled polygons constructed from connected component analysis on a photograph of vascular cambium, a type of plant tissue.
     \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.
     \protect\subref{fig:vascular-2} All conflict cells of the initial quadtree.
-    \protect\subref{fig:vascular-3} After conflict cell resolution.
-No quadtree cell intersects more than one object.
-\red{Can you get a similar figure to this that has closed polygons and is maybe zoomed in a little more two a few nice-looking polygons?}
+    \protect\subref{fig:vascular-3} After conflict cell resolution. No quadtree cell intersects more than one object.
   }

### 2017-04-03 paper.tex
updated figure 5.

@@
   \caption{
-    \protect\subref{fig:simple-conflicts} \red{On the bottom right, the box in the 3rd quadrant is not a conflict, and shouldn't be colored red.} A toy dataset showing conflict cells after building the quadtree from object vertices.
+    \protect\subref{fig:simple-conflicts} A toy dataset showing conflict cells after building the quadtree from object vertices.
     \protect\subref{fig:simple-samples} The toy dataset showing how samples are collected.

### 2017-04-03 paper.tex
Optimality, Nate changes

--- 2017-04-02 paper.tex
@@
 \documentclass[final,3p,times,twocolumn]{elsarticle}
+%\documentclass[submission]{gmp2017}
+%\documentclass{gmp2017}
+
+% Put your submission number here.
+% This number will appear instead of authors and affiliations if the
+% class option "submission" is active.
+%\SubNumber{19}
+
+% Put the author names here.
+% Use the second argument as a reference to the list of affiliations.
+% No authors and affiliations will appear if the class option "submission"
+% is active.
+%\author{Nathan Vollmer}{1}
+%\author{John Edwards}{1}
+
+% Put the affiliations of the authors here.
+%\affiliation{1}{Department of Informatics and Computer Science, Idaho State University, USA}
 
@@
 \input{common}
+
+%\usepackage[inline]{showlabels}
 
@@
 
-\title{Fast Construction of Inter-Object Spacing Representations}
-\author{Nathan Morrical}
+%\title{Parallel Quadtree Construction on Collections of Objects}
+\title{Fast Construction of Inter-Object Space Representations}
+
 %% use optional labels to link authors explicitly to addresses:
@@
 
-\author{John Edwards}
-
-\address{Department of Informatics and Computer Science, Idaho State University, USA}
+\author{}
+
+\address{}
 
@@
 \begin{abstract}
+%We present a parallel quadtree algorithm that resolves between geometric objects.
+Our quadtree has the property that no cell intersects more than one labeled object.
+A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori, because the number of cells required to resolve closely spaced objects exceeds available memory.
+Previous parallel algorithms that are spatially adaptive, or that discretize finely in only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation.
+Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
+We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D.
+Our results show significant improvement over the current state of the art.
 We present a parallel quadtree algorithm that resolves between geometric objects, modeling space between objects rather than the objects themselves.
@@
 A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori or because the number of cells required to resolve closely spaced objects exceeds available memory.
-Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, \red{Hierarchical kernel envokation isn't necessarily bad.
-We might move to that to remove some of the Q linear complexity overhead.} either separate points only, or make no guarantees of object separation.
+Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation.
 Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
@@
 \label{sec:intro}
-\red{Reworded this paragraph, since the last sentence contained duplicate information provided in the first sentence.} Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, generalized Voronoi Diagram (GVD) construction \cite{edwards2015approximating}, robot navigation, shape modeling, object description, and other applications.
+Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, robot navigation, shape modeling, object description, and other applications.
 Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects.
-In other cases, quadtrees are purposed for fast retrieval as is often the case in hierarchical subdivisions of point data.
 However, our work centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object.
-
-\red{reworded the first sentence a bit.} This object separation is of some use in 2D (e.g.
-path planning), but becomes a very important problem in 3D.
-Hierarchically subdividing non-point data in a principled parallel way is surprisingly complex, and this paper lays the groundwork for our continuing work in 3D.
-
-Modeling inter-object spacing is computationally straightforward when that spacing is large compared to the world bounding box.
+Such quadtrees can be thought of as modeling the space between objects.
+% Our quadtree representation is not purposed for fast retrieval as is often the case in hierarchical subdivisions of point data.
+% Rather, the quadtrees are used for constructing distance fields, detecting collisions and computing the GVD \cite{edwards2015approximating}.
+
+Modeling inter-object spacing is computationally straightforward when the spacing is large compared to the world bounding box.
 Approaches typically involve a uniform grid of the space, which leads to efficient computation that often uses graphics processors.
@@
 Thus, an adaptive approach must be used for datasets of closely spaced objects.
-\red{added carriage return here}
 
@@
 
-This paper extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution.
+This work extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution.
 Our algorithm, which is targeted for the GPU, performs an order of magnitude faster than the previous work and will be an important base for later distance transform and generalized Voronoi diagram computation.
@@
 Each step is done in parallel either on object vertices, object facets, or quadtree cells.
+
+Modeling object separation is of some use in 2D (e.g.
+path planning), but it is a very important problem in many 3D applications.
+Hierarchically subdividing space between faceted objects in a principled parallel way is complex, and this paper lays the groundwork for our continuing efforts in 3D.
 
@@
 \paragraph{Parallel} Many recent works on fast quadtree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}.
-Kim and Liu's work~\shortcite{kim2014exact} is similar, computing the quadtree on the barycenters of triangles, giving an approximation of an object-resolving quadtree.
+%Kim and Liu's work~\shortcite{kim2014exact} is similar, computing the quadtree on the barycenters of triangles, giving an approximation of an object-resolving quadtree.
  Most quadtree approaches that support surfaces \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed} are designed for efficient rendering, and actual construction of the quadtree is implemented on the CPU.
@@
 Yin \etal~\shortcite{yin2011fast} compute the octree entirely on the GPU using a bottom-up approach by initially subdividing into a complete quadtree, resulting in memory usage that is no better than using a uniform grid.
+The most similar work to what we do here is Kim and Liu's method~\shortcite{kim2014exact}, which computes the quadtree on the barycenters of triangles, giving an approximation of our quadtree, but without fully resolving between objects.
 We have found no GPU quadtree construction method that is fully adaptive and can resolve between objects.
@@
 
-\subsection{Prune the octree}
-During the initial quadtree construction, we can indirectly prune the initial quadtree to simplify conflict detection and reduce our memory footprint.
+%Our first step is to build a quadtree on the vertices.
+This step gives us our first approximation to our final quadtree.
+We use Karras' algorithm \cite{karras2012maximizing} which sorts the Morton codes of the vertices in parallel, then constructs the binary radix tree in parallel.
+With the binary radix tree, the quadtree can be constructed with a single parallel call.
+The strength of this algorithm lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points.
+That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.
+
+\subsection{Pruning the quadtree}
+\label{sec:pruning}
+During initial quadtree construction, we can prune the quadtree to simplify conflict detection and reduce our memory footprint.
 Assume we have a numeric vertex labeling such that each vertex is labeled to match the object it belongs to.
-The initial binary radix tree (BRT) provided by Karras serves as a bounding volume hierarchy, and is used to generate the initial quadtree by seperating vertices regardless of their label.
+The binary radix tree (BRT) provided by Karras serves as a bounding volume hierarchy and is used to generate the initial quadtree by seperating vertices regardless of their label.
 Since our objective is to resolve between objects of different labels, we can proactively prune the initial BRT, and subsequently the initial quadtree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label.
 
-\red{I tried to be a bit more verbose here and in the paragraph above.
-Let me know if anything in particular needs more clarification.} To prune the initial BRT efficiently, we label each BRT node using the following criterion: if the current node is a leaf node which seperates two vertices with identical labels, label the current node to match the label of the vertices being seperated.
-If the current node is a leaf node that seperates two vertices having mismatching colors, label the current node as -2, i.e.
-"required".
-Lastly, if the current node is an internal node, i.e.
-the current node is an ancestor of another BRT node, mark the current node as -1, i.e.
-"unknown''.
-This initial step can be done immediately after Karras' BRT construction without the need to invoke an additional kernel.
-
-We then propagate the BRT labels up the tree in parallel, marking "unknown'' nodes as "required" when the labels of the current node's two children nodes mismatch.
+%The initial quadtree resolves between vertices, regardless of whether they are labeled differently from each other or not, so each leaf node is guaranteed to contain no more than one vertex.
+Since our objective is to resolve between objects of different labels, we can proactively prune the initial vertex quadtree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label.
+By pruning the initial quadtree we simplify conflict detection and reduce our memory footprint.
+
+%\red{Clarify this passage: We remove internal binary radix tree nodes seperating identically labeled vertices.
+For each binary radix tree leaf in parallel, we first look at the label of the leaf's corresponding vertices.
+If the colors match, we apply that color to the leaf.
+Otherwise, we mark that leaf as "required".} This initial coloring can be done immediately after the Karras binary radix tree construction without the need to invoke an additional kernel.
+Before generating the binary radix tree's corresponding quadtree, we propagate the binary radix leaf colors up the tree in parallel, marking internal nodes as "required" when the colors of two children nodes mismatch.
+Finally, we generate quadtree nodes from only the required internal binary radix tree nodes.
+By pruning the initial quadtree, we simplify conflict detection and reduce our memory footprint.
+
+\red{Talk to Nate}
+To prune the initial BRT efficiently, we label each BRT node $C$ using the following criterion: if $C$ is a leaf node that separates two vertices with identical labels, label $C$ to match the label of the vertices being separated.
+If $C$ is a leaf node that separates two vertices having mismatched colors, label $C$ as ``required''.
+Lastly, if $C$ is an internal node, i.e., it has children, mark it as ``unknown''.
+This initial step can be done immediately after the Karras BRT construction without the need to invoke an additional kernel.
+
+We then propagate the BRT labels up the tree in parallel, marking ``unknown'' nodes as ``required'' when the labels of the current node's two child nodes don't match.
 Finally, we generate quadtree nodes from only the required internal binary radix tree nodes.
@@
 We define a bounding cell (BCell) to be the smallest internal quadtree node which entirely contains a given facet.
-Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the BCell is contained in the longest common prefix (LCP) of the Morton codes of the points in $P$.
-If a given LCP is more specific than any quadtree node, ie the LCP lies within a quadtree leaf, we simply take the quadtree address of the leaf that the LCP lies within.
+Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the BCell is the longest common prefix of the Morton codes of the points in $P$.
+If a given LCP is more specific than any quadtree node, i.e., if the LCP lies within a quadtree leaf, we simply take the quadtree address of the leaf that the LCP lies within.
 This is often the case with pruned quadtrees where entire facets may lie within a quadtree leaf.
-\red{This paragraph had some incorrect conclusions, so I tried to clarify a bit more.} Figure \ref{fig:data-structures-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.
+\red{Please check to make sure I got the changes right.} Figure \ref{fig:data-structures-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.
 
@@
 
-Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.
+Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras quadtree construction.
 
@@
 
-    \node[anchor=south west,inner sep=0,outer sep=0] at (0,-4) {
-%      \subfloat[][]{
-%        \label{fig:conflict-find-2}
-%        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
-%      \subfloat[][]{
-%        \label{fig:conflict-find-4}
-%        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
-      \subfloat[][]{
-        \label{fig:conflict-find-5}
-        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
-      \subfloat[][]{
-        \label{fig:conflict-find-6}
-        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }
-    };
+%    \node[anchor=south west,inner sep=0,outer sep=0] at (0,-4) {
+%%      \subfloat[][]{
+%%        \label{fig:conflict-find-2}
+%%        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
+%%      \subfloat[][]{
+%%        \label{fig:conflict-find-4}
+%%        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
+%      \subfloat[][]{
+%        \label{fig:conflict-find-5}
+%        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
+%      \subfloat[][]{
+%        \label{fig:conflict-find-6}
+%        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }
+%    };
 
@@
   \end{tikzpicture}
+
+%      \subfloat[][]{
+%        \label{fig:conflict-find-2}
+%        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
+%      \subfloat[][]{
+%        \label{fig:conflict-find-4}
+%        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
+      \subfloat[][]{
+        \label{fig:conflict-find-5}
+        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
+      \subfloat[][]{
+        \label{fig:conflict-find-6}
+        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }
+
 %      \subfloat[][]{
@@
 %    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and green with facets labeled A-I.
+    \red{Is the pruning correct?}
     We have three objects, blue, red, and green with facets labeled A-I.
-    \protect\subref{fig:conflict-find-2} Initial vertex quadtree. \red{This vertex quadtree is unpruned. Can we prune it for consistency?}
-    \protect\subref{fig:conflict-find-4} Zoomed-in to the region outlined by red in \protect\subref{fig:conflict-find-2} showing the boundary cell (BCell) computation for each facet.
+    \protect\subref{fig:conflict-find-2} Initial pruned vertex quadtree.
+    \protect\subref{fig:conflict-find-4} Zoomed-in to the region outlined by red in \protect\subref{fig:conflict-find-2} and showing the boundary cell (BCell) computation for each facet.
 %These pairs are given in figure \protect\subref{fig:data-structures-1}.
@@
 %    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access.
-We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given octree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
-%\protect\subref{fig:data-structures-3} For a given octree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
+We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given quadtree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
+%\protect\subref{fig:data-structures-3} For a given quadtree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
   }
@@
     \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access.
-We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given octree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
-\protect\subref{fig:data-structures-3} For a given octree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
+We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given quadtree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
+\protect\subref{fig:data-structures-3} For a given quadtree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
   }
@@
 
-Then we use the \texttt{BCells} array and octree data structure to find the conflict cells using algorithm \ref{alg:find-conflict-cells}.
+Then we use the \texttt{BCells} array and quadtree data structure to find the conflict cells using algorithm \ref{alg:find-conflict-cells}.
 We process each leaf cell $L$ in parallel (line \ref{alg:L}).
 First, we set $L$'s color to $-1$ (uninitialized).
-We then traverse each ancestor $A$ of $L$ (line \ref{alg:A}) by using the \texttt{Parent} field in the octree data structure.
+We then investigate each ancestor $A$ of $L$ (line \ref{alg:A}) by using the \texttt{Parent} field in the quadtree data structure.
 Using the \texttt{FFacet} and \texttt{LFacet} fields, we find, respectively, the first and (inclusive) last of possibly multiple facets bounded by $A$ (line \ref{alg:i}).
@@
   \subfloat[][]{
-    \label{fig:conflict-resolution-octree}
+    \label{fig:conflict-resolution-quadtree}
     \includegraphics[width=0.35\columnwidth]{conflict-resolution-octree.pdf} }\\
@@
     \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least two samples.
-This ensures that a quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-octree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
+This ensures that a quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-quadtree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
   }
@@
 that bisects $q$ and $r$.
-Our strategy will be to sample points $P$ on $p(s)$ (figure \ref{fig:conflict-resolution-octree}) such that a quadtree built on $S \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent cell of $c$ will intersect both $q$ and $r$.
+Our strategy will be to sample points $P$ on $p(s)$ (figure \ref{fig:conflict-resolution-quadtree}) such that a quadtree built on $S \cup P$ will completely ``separate'' $q$ and $r$, i.e., no descendent leaf of $C$ will intersect both $q$ and $r$.
 We do this by ensuring that $P$ is sampled such that every box that intersects both $q$ and $r$ also intersects at least two points in $P$.
@@
 We break the problem into two cases:
-\begin{tightenumerate}
+%\begin{tightenumerate}
+\begin{enumerate}
  \item The \textit{opposite} case (see figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
  \item In the \textit{adjacent} case (see figure \ref{fig:conflict-resolution-adjacent-even}), $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
-\end{tightenumerate}
+%\end{tightenumerate}
+\end{enumerate}
 
@@
 Because conflict cell resolution only considers two facets at a time, we may have to iterate multiple times if more than two facets intersect a given cell.
-If new sample points were found then we add them to the current set $S$ of sample points and return to building the octree from points (section \ref{sec:build-initial-octree}).
+If new sample points were found then we add them to the current set $S$ of sample points and return to building the quadtree from points (section \ref{sec:build-initial-quadtree}).
 We finish when the only conflicts identified are at the maximum depth.
 
+\subsection{Optimality}
+\red{Check this}
+Define an optimal quadtree to be one in which only conflict nodes have children, and let an optimal quadtree's size be $n$ total nodes.
+Our iterative sampling algorithm results in a quadtree that has a size within a constant factor of $n$ in the worst case (see figure \ref{fig:optimality}).
+We omit the proof as well as an average case analysis because optimality can be achieved by simply performing one final pruning step (see section \ref{sec:pruning}).
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:optimality-good}
+    \includegraphics[width=0.25\columnwidth]{optimality-good.pdf} }
+  \subfloat[][]{
+    \label{fig:optimality-bad}
+    \includegraphics[width=0.25\columnwidth]{optimality-bad.pdf} }
+  \subfloat[][]{
+    \label{fig:optimality-bad-all}
+    \includegraphics[width=0.25\columnwidth]{optimality-bad-all.pdf} }
+  \caption{
+    Best and worst cases given two lines. The same number of conflict resolution samples are generated regardless of where the lines are located.
+    \protect\subref{fig:optimality-good} Base case: two lines can be resolved by a single quadtree subdivision.
+    \protect\subref{fig:optimality-bad} Worst case: the same two lines translated slightly in $y$ now require five subdivisions to be resolved.
+    \protect\subref{fig:optimality-bad-all} The number of cells generated from the shown resolution samples is within a constant factor of the worst case.
+  }
+  \label{fig:optimality}
+\end{figure}
+
 \subsection{Complexity analysis}
@@
 Let $M = |F|$ and $N = |V|$, where $F$ are the object facets and $V$ are the object vertices.
-Let $D$ be the depth of the octree.
+Let $D$ be the depth of the quadtree.
 In this analysis we assume sufficient parallel units to maximize parallelization.
 
-\vspace{2mm}
-\paragraph{Time complexity}
-
-\begin{tightenumerate}
-\item Build octree using Karras' algorithm \cite{karras2012maximizing} - $O(D)$.
-\item Prune the Binary Radix Tree from Karras \cite{karras2012maximizing} - $O(D)$ \red{Does this sound right?}
+%\vspace{2mm}
+%\paragraph{Time complexity}
+\subsubsection{Time complexity}
+
+%\begin{tightenumerate}
+\begin{enumerate}
+\item Build quadtree using Karras' algorithm \cite{karras2012maximizing}, including pruning - $O(D)$. \red{How about this?}
 \item Detect conflict cells
-  \begin{tightenumerate}
+%  \begin{tightenumerate}
+  \begin{enumerate}
   \item Build BCells array - $O(D)$.
@@
 This requires looking at each bit, of which there are $O(D)$.
-  \item Sort BCells array - $O(\log{M})$.
-\red{Shouldn't the big o here be O(n), where n is the number of levels in the quadtree?
-} The array has $M$ elements, and we use a parallel radix sort with $\log$ complexity.
-  \item Index BCells with octree data structure - $O(D)$. This runs in parallel on leaf cell IDs and each kernel requires a search of the octree for a given cell ID, taking at most $D$ steps.
+  \item Sort BCells array - \red{Shouldn't the Big O here be $O(D)$?} $O(\log{M})$. The array has $M$ elements, and we use a parallel radix sort with $\log$ complexity.
+  \item Index BCells with quadtree data structure - $O(D)$. This runs in parallel on leaf cell IDs and each kernel requires a search of the quadtree for a given cell ID, taking at most $D$ steps.
   \item Find facets that intersect each leaf cell - Worst case $O(M+D)$, average case $O(D)$.
 In unusual datasets, a single leaf cell will be intersected by $O(M)$ facets.
-On average, however, leaf cells intersect a small number of facets, and thus this step is dominated by the depth $D$ of the octree due to visiting each ancestor of the leaf cell.
-  \end{tightenumerate}
+On average, however, leaf cells intersect a small number of facets, and thus this step is dominated by the depth $D$ of the quadtree due to visiting each ancestor of the leaf cell.
+%  \end{tightenumerate}
+  \end{enumerate}
 \item Resolve conflict cells
-  \begin{tightenumerate}
+%  \begin{tightenumerate}
+  \begin{enumerate}
   \item Compute new sample points - $O(1)$.
@@
   \item $S \leftarrow S \cup S'$ - $O(1)$.
-  \end{tightenumerate}
+%  \end{tightenumerate}
+  \end{enumerate}
 \item Iterate - $O(Q)$ iterations. In the worst case, all facets intersect a single cell, requiring potentially $Q=O(M^2)$ iterations. In our testing, $Q$ has not exceeded 4.
-\end{tightenumerate}
+%\end{tightenumerate}
+\end{enumerate}
 
 The final complexity of each iteration is $O(M+D)$ worst case and $O(\log{M}+D)$ average case.
-In practice we must fix the depth of the octree to a constant value in order to use a predetermined integer size for the Morton codes, which brings the average case complexity to $O(\log{M})$.
+In practice we must fix the depth of the quadtree to a constant value in order to use a predetermined integer size for the Morton codes, which brings the average case complexity to $O(\log{M})$.
 Taking iteration into account, the final complexity is $(Q\log{M})$ average case.
 
-\vspace{2mm}
-\paragraph{Space complexity}
+%\vspace{2mm}
+%\paragraph{Space complexity}
+\subsubsection{Space complexity}
 
@@
 In 2D, in the worst case, the facets can form an arrangement of maximum number of intersections, which is $M(M-1)/2 = O(M^2)$.
-If this is the case then we subdivide to the maximum octree depth at each intersection, causing an octree of size $O(DM^2)$.
-%In our testing, our octrees have never exceeded a size of \red{fill this in with experimental results}.
+If this is the case then we subdivide to the maximum quadtree depth at each intersection, causing a quadtree of size $O(DM^2)$.
+%In our testing, our quadtrees have never exceeded a size of \red{fill this in with experimental results}.
 
@@
 of the algorithm supports polygons and polylines which needn't be manifold or connected.
-All tests were run on a Razer Blade Stealth with an Intel i7 6500u 3.10 ghz dual core processor, 8GB of memory, and an Nvidia GTX 1070 graphics card.
-Figure \ref{fig:results} shows results on four datasets: a simple toy dataset showing conflict cell detection and resolution (\ref{fig:simple-conflicts}-\ref{fig:simple-samples}); a more complex maze dataset (\ref{fig:glines}), and a complex dataset with many objects at very different scales (\ref{fig:ut1-samples}-\ref{fig:ut5}).
+All tests were run on a \red{Razer Blade Stealth with an Intel i7 6500u 3.10 GHz dual core processor, 8 GB of memory and an Nvidia GTX 1070 graphics card.} Figure \ref{fig:results-toy} shows results a simple toy dataset showing conflict cell detection and resolution.
+A very complex dataset with many objects at very different scales is shown in figure \ref{fig:results-maze}.
+It demonstrates that our method can handle datasets far beyond the memory limits of uniform grid approaches while still fully resolving between objects.
+The gears dataset (figure \ref{fig:gears}) again shows a large domain-to-object-spacing ratio, as well as non-convexities.
+The vascular dataset shown in figure \ref{fig:vascular} demonstrates our method on polylines derived from medical image data, which is often noisy with non-manifoldness and intersections.
 Table \ref{tab:timings} shows timings for our implementation compared to the previous state-of-the-art.
@@
   \caption{
-    \protect\subref{fig:simple-conflicts} \red{On the bottom right, the box in the 3rd quadrant is not a conflict, and shouldn't be colored red.} A toy dataset showing conflict cells after building the quadtree from object vertices.
+    \protect\subref{fig:simple-conflicts} \red{Incorrect conflict cells.} A toy dataset showing conflict cells after building the quadtree from object vertices.
     \protect\subref{fig:simple-samples} The toy dataset showing how samples are collected.
@@
   \label{fig:results-toy}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfloat[][]{
-    \label{fig:vascular-1}
-    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/04-Pruned.PNG} }
-  \subfloat[][]{
-    \label{fig:vascular-2}
-    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/06-Initial-Conflicts.PNG} }
-  \subfloat[][]{
-    \label{fig:vascular-3}
-%    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/08-Resolved-Quadtree.PNG} }
-    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/13-Resolved-Quadtree.PNG} }
-  \caption{\red{Can you give some details on this dataset -- what it is etc}
-    \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.
-    \protect\subref{fig:vascular-2} All conflict cells of the initial quadtree.
-    \protect\subref{fig:vascular-3} After conflict cell resolution.
-No quadtree cell intersects more than one object.
-\red{Can you get a similar figure to this that has closed polygons and is maybe zoomed in a little more two a few nice-looking polygons?}
-  }
-  \label{fig:vascular}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfloat[][]{
-    \label{fig:gears-1}
-    \includegraphics[width=0.31\columnwidth]{Nathan/gears/2.PNG} }
-  \subfloat[][]{
-    \label{fig:gears-2}
-    \includegraphics[width=0.31\columnwidth]{Nathan/gears/3.PNG} }
-  \subfloat[][]{
-    \label{fig:gears-3}
-    \includegraphics[width=0.31\columnwidth]{Nathan/gears/5.PNG} }
-  \caption{
-    \protect\subref{fig:gears-1} A dataset of gears with close tolerance. The resolved quadtree with sampled points is shown.
-    \protect\subref{fig:gears-2} Showing just the quadtree and sample points.
-    \protect\subref{fig:gears-3} A zoomed-in image showing the close object spacing compared to the large domain.
-  }
-  \label{fig:gears}
 \end{figure}
@@
     \protect\subref{fig:ut1-samples} A complex dataset with 470 objects at vastly different scales in object size and spacing.
-    \protect\subref{fig:ut1}-\protect\subref{fig:ut5} Complex dataset at different zoom levels up to 60K magnification. This shows the importance of an adaptive method such as a quadtree.
+    \protect\subref{fig:ut1}-\protect\subref{fig:ut5} Complex dataset at different zoom levels up to 60K magnification.
+This shows the importance of an adaptive method such as a quadtree.
+A uniform grid would require $2^{48}$ cells to resolve between objects.
+The quadtree shown here has $22429$ cells.
   }
   \label{fig:results-maze}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:gears-1}
+    \includegraphics[width=0.31\columnwidth]{Nathan/gears/2.PNG} }
+  \subfloat[][]{
+    \label{fig:gears-2}
+    \includegraphics[width=0.31\columnwidth]{Nathan/gears/3.PNG} }
+  \subfloat[][]{
+    \label{fig:gears-3}
+    \includegraphics[width=0.31\columnwidth]{Nathan/gears/5.PNG} }
+  \caption{
+    \protect\subref{fig:gears-1} A dataset of gears with close tolerance. The resolved quadtree with sampled points is shown.
+    \protect\subref{fig:gears-2} Showing just the quadtree and sample points.
+    \protect\subref{fig:gears-3} A zoomed-in image showing the close object spacing compared to the large domain.
+  }
+  \label{fig:gears}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:vascular-1}
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/04-Pruned.PNG} }
+  \subfloat[][]{
+    \label{fig:vascular-2}
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/06-Initial-Conflicts.PNG} }
+  \subfloat[][]{
+    \label{fig:vascular-3}
+%    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/08-Resolved-Quadtree.PNG} }
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/13-Resolved-Quadtree.PNG} }
+  \caption{\red{Can you give some details on this dataset -- what it is etc}
+    \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.
+    \protect\subref{fig:vascular-2} All conflict cells of the initial quadtree.
+    \protect\subref{fig:vascular-3} After conflict cell resolution.
+No quadtree cell intersects more than one object.
+Our method works even though objects in this dataset are often non-manifold and have self-intersections.
+  }
+  \label{fig:vascular}
 \end{figure}

### 2017-04-03 paper.tex

@@
 
+%\author{Nathan Morrical}
+%\author{John Edwards}
+%\address{Department of Informatics and Computer Science, Idaho State University, USA}
+
 %% use optional labels to link authors explicitly to addresses:
@@
 %% \address[label2]{<address>}
-
-\author{}
-
-\address{}
 
@@
 
-This work extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution.
+To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory, but it does so in serial, with expected performance liabilities.
+A naive approach to parallelizing quadtree computation would be to assign all available compute units according to a course grid, then run the serial algorithm on each compute unit.
+While simple, there is potential for serious load imbalancing if the close object spacings are not uniformly distributed.
+
+This paper extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution.
 Our algorithm, which is targeted for the GPU, performs an order of magnitude faster than the previous work and will be an important base for later distance transform and generalized Voronoi diagram computation.
@@
   \caption{
-    \protect\subref{fig:simple-conflicts} \red{Incorrect conflict cells.} A toy dataset showing conflict cells after building the quadtree from object vertices.
+    \protect\subref{fig:simple-conflicts} A toy dataset showing conflict cells after building the quadtree from object vertices.
     \protect\subref{fig:simple-samples} The toy dataset showing how samples are collected.
@@
 %    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/08-Resolved-Quadtree.PNG} }
-    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/13-Resolved-Quadtree.PNG} }
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/14-Resolved-Quadtree.PNG} }
   \caption{\red{Can you give some details on this dataset -- what it is etc}

### 2017-04-03 paper.tex
removed red text

@@
 of the algorithm supports polygons and polylines which needn't be manifold or connected.
-All tests were run on a \red{Razer Blade Stealth with an Intel i7 6500u 3.10 GHz dual core processor, 8 GB of memory and an Nvidia GTX 1070 graphics card.} Figure \ref{fig:results-toy} shows results a simple toy dataset showing conflict cell detection and resolution.
+All tests were run on a Razer Blade Stealth with an Intel i7 6500u 3.10 GHz dual core processor, 8 GB of memory and an Nvidia GTX 1070 graphics card.
+Figure \ref{fig:results-toy} shows results a simple toy dataset showing conflict cell detection and resolution.
 A very complex dataset with many objects at very different scales is shown in figure \ref{fig:results-maze}.

### 2017-04-03 paper.tex

@@
 The gears dataset (figure \ref{fig:gears}) again shows a large domain-to-object-spacing ratio, as well as non-convexities.
-The vascular dataset shown in figure \ref{fig:vascular} demonstrates our method on polylines derived from medical image data, which is often noisy with non-manifoldness and intersections.
+The vascular dataset shown in figure \ref{fig:vascular} demonstrates our method on polylines derived from biological image data, which is often noisy with non-manifoldness and intersections.
 Table \ref{tab:timings} shows timings for our implementation compared to the previous state-of-the-art.
@@
     \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/14-Resolved-Quadtree.PNG} }
-  \caption{\red{Can you give some details on this dataset -- what it is etc}
+  \caption{Dataset derived from plant vasculature image.
     \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.

### 2017-04-03 paper.tex
Updated 3.1 to explain Karras in more depth. Adjusted 3.2 to take into account expanded BRT explanation in 3.1. Updated stats table. Updated 3.6.1 big o.

@@
 A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori or because the number of cells required to resolve closely spaced objects exceeds available memory.
-Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, \red{Hierarchical kernel envokation isn't necessarily bad.
-We might move to that to remove some of the Q linear complexity overhead.} either separate points only, or make no guarantees of object separation.
+Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, either separate points only, or make no guarantees of object separation.
 Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
@@
 Our first step is to build a quadtree on the given set of vertices.
-We use the Karras algorithm \cite{karras2012maximizing} which starts by sorting the Morton codes of the given vertices.
-Our implementation uses an efficient parallel radix sorter described by Ha \etal \shortcite{ha2009fast}.
-Once the vertices are sorted, a binary radix tree, and then an initial quadtree can be constructed in parallel.
+We use the Karras algorithm \cite{karras2012maximizing} which begins by placing the given vertices on a Z-Order curve by computing each vertex's cooresponding Morton code in parallel.
+Next, Karras sorts the converted points by using a parallel radix sorter, which has a linear execution time.
+Our implementation uses the efficient four-way parallel radix sorter described by Ha \etal \shortcite{ha2009fast}.
+Once the morton codes are sorted, the Z-Order curve can be exploited to construct a binary radix tree in a parallel bottom up manner by identifying longest common Morton code prefixes between neighboring points.
+This resultant binary radix tree can be analyzed in parallel to identify the size and structure of the required vertex quadtree.
 The strength of this approach lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points.
@@
 \label{sec:pruning}
-During initial quadtree construction, we can prune the quadtree to simplify conflict detection and reduce our memory footprint.
+During Karras' initial binary radix tree (BRT) construction, we can prune the BRT to simplify the resultant quadtree.
+This in turn simplifies the work complexity of conflict cell detection and reduces our overall memory footprint.
 Assume we have a numeric vertex labeling such that each vertex is labeled to match the object it belongs to.
-The binary radix tree (BRT) provided by Karras serves as a bounding volume hierarchy and is used to generate the initial quadtree by seperating vertices regardless of their label.
-Since our objective is to resolve between objects of different labels, we can proactively prune the initial BRT, and subsequently the initial quadtree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label.
-
-%The initial quadtree resolves between vertices, regardless of whether they are labeled differently from each other or not, so each leaf node is guaranteed to contain no more than one vertex.
-Since our objective is to resolve between objects of different labels, we can proactively prune the initial vertex quadtree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label.
-By pruning the initial quadtree we simplify conflict detection and reduce our memory footprint.
-
-%\red{Clarify this passage: We remove internal binary radix tree nodes seperating identically labeled vertices.
-For each binary radix tree leaf in parallel, we first look at the label of the leaf's corresponding vertices.
-If the colors match, we apply that color to the leaf.
-Otherwise, we mark that leaf as "required".} This initial coloring can be done immediately after the Karras binary radix tree construction without the need to invoke an additional kernel.
-Before generating the binary radix tree's corresponding quadtree, we propagate the binary radix leaf colors up the tree in parallel, marking internal nodes as "required" when the colors of two children nodes mismatch.
-Finally, we generate quadtree nodes from only the required internal binary radix tree nodes.
-By pruning the initial quadtree, we simplify conflict detection and reduce our memory footprint.
-
-\red{Talk to Nate}
+The original BRT provided by Karras' algorithm is used to seperate vertices regardless of their label.
+Since our objective is to resolve between objects of different labels, we can proactively prune Karras' initial BRT, and subsequently the initial quadtree (see figure \ref{fig:pruning}) by allowing the generated quadtree leaves to contain multiple vertices as long as those vertices have the same label.
+
 To prune the initial BRT efficiently, we label each BRT node $C$ using the following criterion: if $C$ is a leaf node that separates two vertices with identical labels, label $C$ to match the label of the vertices being separated.
@@
 %    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and green with facets labeled A-I.
-    \red{Is the pruning correct?}
     We have three objects, blue, red, and green with facets labeled A-I.
@@
 \subsection{Optimality}
-\red{Check this}
-Define an optimal quadtree to be one in which only conflict nodes have children, and let an optimal quadtree's size be $n$ total nodes.
+Define an optimal final quadtree to be one in which only conflict nodes have children, and let an optimal final quadtree's size be $n$ total nodes.
 Our iterative sampling algorithm results in a quadtree that has a size within a constant factor of $n$ in the worst case (see figure \ref{fig:optimality}).
-We omit the proof as well as an average case analysis because optimality can be achieved by simply performing one final pruning step (see section \ref{sec:pruning}).
+We omit the proof as well as an average case analysis because optimality can be achieved by simply removing the uneccessary nodes in one final parallel pruning step.
 
@@
 \begin{enumerate}
-\item Build quadtree using Karras' algorithm \cite{karras2012maximizing}, including pruning - $O(D)$. \red{How about this?}
+\item Build quadtree using Karras' algorithm \cite{karras2012maximizing}, including pruning - $O(D)$.
 \item Detect conflict cells
@@
 This requires looking at each bit, of which there are $O(D)$.
-  \item Sort BCells array - \red{Shouldn't the Big O here be $O(D)$?} $O(\log{M})$. The array has $M$ elements, and we use a parallel radix sort with $\log$ complexity.
+  \item Sort BCells array -  \red{$O(D)$}. We use a parallel radix sort with linear complexity dependent on the max quadtree depth.
   \item Index BCells with quadtree data structure - $O(D)$. This runs in parallel on leaf cell IDs and each kernel requires a search of the quadtree for a given cell ID, taking at most $D$ steps.
@@
     \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/14-Resolved-Quadtree.PNG} }
-  \caption{Dataset derived from plant vasculature image.
+  \caption{A large set of uniquely labeled polygons constructed from connected component analysis on a photograph of vascular cambium, a type of plant tissue.
     \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.
@@
   \footnotesize{
-%  \begin{tabular}{lcc|cc|cc|cc|cc}
-  \begin{tabular}{lcc|cc|cc|cc}
+  \begin{tabular}{lcc|c|cc|cc}
     \toprule
-    dataset & objects & object          & \multicolumn{2}{c|}{quadtree}
+    dataset & objects & object          & quadtree
     & \multicolumn{2}{c|}{time} &  \multicolumn{2}{c}{quad cells} 
     \\
-%    & \multicolumn{2}{c}{quad memory}\\
-            &         & facets          & \multicolumn{2}{c|}{depth}
-    & \multicolumn{2}{c|}{(millisec)} & \multicolumn{2}{c}{($\times 10^3$)}
+            &         & facets          & depth
+    & \multicolumn{2}{c|}{(millisec)} & 
     \\
-%    & \multicolumn{2}{c}{(Mb)}    \\
-%            &         & ($\times 10^3$) & Ours & Prev
-            &         &  & Ours & Prev
-%    & Ours & Prev & Ours & Prev & Ours & Prev \\
+           &         &  & 
     & Ours & Prev & Ours & Prev \\
-    %\midrule
     \hline
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/test1*.dat
-    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 10 & 9 & 54 & 3 & 177 & 1168  \\
+    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 9 & 9 & 3 & 138 & 1213  \\
+
+    % ./gvd-viewer2 -l 9 ../data2/glines/200x200/*.dat
+    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/glines/200x200/*
+    Fig. \ref{fig:gears} & 12 & 288 & 10 & 23 & 24 & 408 & 3370  \\
 
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 3 -l 24 data/maze/poly*.dat
-    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 24 & 128 & 465 & 38 & 157  \\
+    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 87 & 277 & 22355 & 156649  \\
 
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 9 data/glines/100x100/*
-    Fig. \ref{fig:glines} & 2 & 27,998 & 9 & 8 & 148 & 429 & 43 & 66  \\
-
-    % ./gvd-viewer2 -l 9 ../data2/glines/200x200/*.dat
-    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/glines/200x200/*
-    Fig. \ref{fig:glines} x2 & 2 & 113,084 & 10 & 9 & 414 & 1778 & 125 & 262  \\
+    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 84 & 376 & 14338 & 157006  \\
 
@@
   \end{tabular}}
-  \caption{\red{Nate, please send me an updated table in whatever format (I can throw it into Latex).
-We should have timings on simple, maze, vascular and gears comparing GVD to PGVD with and without pruning.
-Not sure if we'll end up including the non-pruning results.
-We should also include the \# objects, \# facets and \# quadtree cells, as are included in this table.} Table of quadtree computation statistics and timings on datasets that are unmanageable using other methods.
+  \caption{Table of quadtree computation statistics and timings on datasets that are unmanageable using other methods.
 Columns are: \emph{objects} - the number of objects in the dataset; \emph{object facets} - the number of line segments (2D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{time (ms)} - milliseconds to build the quadtree; \emph{quad cells} - number of quadtree cells.
-Dataset ``\ref{fig:glines} x2'' is a maze dataset increased in size by a factor of two in each dimension from \ref{fig:glines}.
 }

### 2017-04-03 paper.tex
Updated conclusion and other minor edits

@@
 \usepackage{lineno}
-
 %% Avoids linenumbers to collide with text for 5p format:
 \setlength\linenumbersep{3pt}
+
+\usepackage[export]{adjustbox}
 
@@
 
-%\title{Parallel Quadtree Construction on Collections of Objects}
-\title{Fast Construction of Inter-Object Space Representations}
+\title{Parallel Quadtree Construction on Collections of Objects}
+%\title{Fast Construction of Inter-Object Space Representations}
 
@@
 A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori or because the number of cells required to resolve closely spaced objects exceeds available memory.
-Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, either separate points only, or make no guarantees of object separation.
+Previous parallel algorithms that are spatially adaptive, i.e., discretizing finely only where needed, either separate points only or make no guarantees of object separation.
 Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
@@
 This is often the case with pruned quadtrees where entire facets may lie within a quadtree leaf.
-\red{Please check to make sure I got the changes right.} Figure \ref{fig:data-structures-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.
+Figure \ref{fig:data-structures-1} gives the addresses of the BCells of the facets in figure \ref{fig:conflict-find-4}.
 
@@
 Our iterative sampling algorithm results in a quadtree that has a size within a constant factor of $n$ in the worst case (see figure \ref{fig:optimality}).
-We omit the proof as well as an average case analysis because optimality can be achieved by simply removing the uneccessary nodes in one final parallel pruning step.
+We omit the proof as well as an average case analysis because optimality can be achieved by simply removing unnecessary nodes in one final parallel pruning step.
 
@@
 Let $M = |F|$ and $N = |V|$, where $F$ are the object facets and $V$ are the object vertices.
-Let $D$ be the depth of the quadtree.
+Let $D$ be the depth of the quadtree, and $D_{max}$ be the maximum depth of the quadtree.
 In this analysis we assume sufficient parallel units to maximize parallelization.
@@
 \begin{enumerate}
-\item Build quadtree using Karras' algorithm \cite{karras2012maximizing}, including pruning - $O(D)$.
+\item Build quadtree using Karras' algorithm \cite{karras2012maximizing}, including pruning - $O(D_{max})$.
 \item Detect conflict cells
@@
 This requires looking at each bit, of which there are $O(D)$.
-  \item Sort BCells array -  \red{$O(D)$}. We use a parallel radix sort with linear complexity dependent on the max quadtree depth.
+  \item Sort BCells array - $O(D_{max})$. We use a parallel radix sort with linear complexity dependent on the max quadtree depth.
   \item Index BCells with quadtree data structure - $O(D)$. This runs in parallel on leaf cell IDs and each kernel requires a search of the quadtree for a given cell ID, taking at most $D$ steps.
@@
 
-The final complexity of each iteration is $O(M+D)$ worst case and $O(\log{M}+D)$ average case.
+The final complexity of each iteration is $O(M+D_{max})$ worst case and $O(\log{M}+D_{max})$ average case.
 In practice we must fix the depth of the quadtree to a constant value in order to use a predetermined integer size for the Morton codes, which brings the average case complexity to $O(\log{M})$.
@@
 }
-of the algorithm supports polygons and polylines which needn't be manifold or connected.
-All tests were run on a Razer Blade Stealth with an Intel i7 6500u 3.10 GHz dual core processor, 8 GB of memory and an Nvidia GTX 1070 graphics card.
+of the algorithm supports polygons and polylines which needn't be manifold or connected. All tests were run on
+% a Razer Blade Stealth with
+ an Intel i7 6500u 3.10 GHz dual core processor, 8 GB of memory and an Nvidia GTX 1070 graphics card.
 Figure \ref{fig:results-toy} shows results a simple toy dataset showing conflict cell detection and resolution.
@@
 A uniform grid would require $2^{48}$ cells to resolve between objects.
-The quadtree shown here has $22429$ cells.
+The quadtree shown here has 22,429 cells.
   }
@@
   \footnotesize{
+%  \scriptsize{
+%  \subfloat[][]{
+%    \label{fig:table-sub}
   \begin{tabular}{lcc|c|cc|cc}
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 3 -l 24 data/maze/poly*.dat
-    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 87 & 277 & 22355 & 156649  \\
+    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 87 & 277 & 22,355 & 156,649  \\
 
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 9 data/glines/100x100/*
-    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 84 & 376 & 14338 & 157006  \\
+    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 84 & 376 & 14,338 & 157,006  \\
+%    Fig. \ref{fig:vascular} & 2162 & 39K & 12 & 84 & 376 & 14,338 & 157,006  \\
 
@@
   \end{tabular}}
-  \caption{Table of quadtree computation statistics and timings on datasets that are unmanageable using other methods.
+%  \subfloat[][]{
+%    \label{fig:timing-chart}
+%    \includegraphics[width=0.8\columnwidth]{timing-chart.png} }
+  \caption{Table of quadtree computation statistics and timings.
+\emph{Ours} is the approach described in this paper and \emph{Prev} is the approach by Edwards \etal \cite{edwards2015approximating}.
 Columns are: \emph{objects} - the number of objects in the dataset; \emph{object facets} - the number of line segments (2D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{time (ms)} - milliseconds to build the quadtree; \emph{quad cells} - number of quadtree cells.
@@
 In fact, our approach actually performs worse on the toy dataset.
-The power of our algorithm becomes more obvious on large, complex datasets, where our performance time gains are significant.
-
-We are in the process of integrating our algorithm with animated systems, generating quadtrees in real-time for collision detection, distance transforms, and generalized Voronoi diagram computation.
+The power of our algorithm becomes obvious on large, complex datasets, where our performance time gains are significant.
+
+As noted in the introduction, our continuing work is in fast construction of octrees modeling inter-object space in 3D.
+%We are in the process of integrating our algorithm with animated systems, generating quadtrees in real-time for collision detection, distance transforms, and generalized Voronoi diagram computation.
 Our implementation continues to be refined and optimized, and we expect to shortly have a version with an order of magnitude improvement over the state of the art.
@@
 %\bibliographystyle{abbrv}
-%\balance
 \bibliography{paper}
+\balance
 

### 2017-04-25 paper.tex
Updated paper to take into account the clarity concerns addressed by Reviewer 3. Formatting's a little messed up now.

@@
 
-To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory, but it does so in serial, with expected performance liabilities.
-A naive approach to parallelizing quadtree computation would be to assign all available compute units according to a course grid, then run the serial algorithm on each compute unit.
-While simple, there is potential for serious load imbalancing if the close object spacings are not uniformly distributed.
-
 This paper extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution.
@@
 The most similar work to what we do here is Kim and Liu's method~\shortcite{kim2014exact}, which computes the quadtree on the barycenters of triangles, giving an approximation of our quadtree, but without fully resolving between objects.
-We have found no GPU quadtree construction method that is fully adaptive and can resolve between objects.
+We are unaware of any GPU quadtree construction methods that are fully adaptive and resolve the seperation between objects.
 
@@
 Our implementation uses the efficient four-way parallel radix sorter described by Ha \etal \shortcite{ha2009fast}.
-Once the morton codes are sorted, the Z-Order curve can be exploited to construct a binary radix tree in a parallel bottom up manner by identifying longest common Morton code prefixes between neighboring points.
+Once the Morton codes are sorted, the Z-Order curve can be exploited to construct a binary radix tree in a parallel bottom up manner by identifying longest common Morton code prefixes between neighboring points.
 This resultant binary radix tree can be analyzed in parallel to identify the size and structure of the required vertex quadtree.
@@
 Assume we have a numeric vertex labeling such that each vertex is labeled to match the object it belongs to.
-The original BRT provided by Karras' algorithm is used to seperate vertices regardless of their label.
+The original BRT provided by Karras' algorithm is used to generate a quadtree which separates vertices regardless of their label.
 Since our objective is to resolve between objects of different labels, we can proactively prune Karras' initial BRT, and subsequently the initial quadtree (see figure \ref{fig:pruning}) by allowing the generated quadtree leaves to contain multiple vertices as long as those vertices have the same label.
@@
 
-Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from Root to $C$, where the local address is a 2-bit (3-bit in 3D) Morton code.
+Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from the root quadtree cell to cell $C$, where the local address is a 2-bit (3-bit in 3D) Morton code.
 The address of the root cell is defined as the empty string.

### 2017-04-26 paper.tex
Changed course grid to coarse grid (reviewer 4), updated fig 8 b to match a. (reviewer 4)

@@
 To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory, but it does so in serial, with expected performance liabilities.
-A naive approach to parallelizing quadtree computation would be to assign all available compute units according to a course grid, then run the serial algorithm on each compute unit.
+A naive approach to parallelizing quadtree computation would be to assign all available compute units according to a coarse grid, then run the serial algorithm on each compute unit.
 While simple, there is potential for serious load imbalancing if the close object spacings are not uniformly distributed.

### 2017-05-03 paper.tex
Moved complexity analysis to appendix
Updated section 3.2 per reviewer one's request

@@
 Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time).
-We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D.
+We describe our algorithm, demonstrate experimental results, and discuss extension to 3D.
 Our results show significant improvement over the current state of the art.
@@
 
-\subsection{Detect conflict cells}
-
-Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from the root quadtree cell to cell $C$, where the local address is a 2-bit (3-bit in 3D) Morton code.
+\subsection{Identifying conflict cells}
+After the pruning in \ref{sec:pruning}, the Karras quadtree separates all differently labeled vertices in the dataset.
+Our goal is to separate differently labeled facets.
+We first need to identify what quadtree cells require further subdivision.
+We call these cells ``conflict cells'' (see figure \ref{fig:conflict-find-5}).
+To efficiently identify conflict cells, we take advantage of the space filling Morton curve and the existing quadtree hierarchy to reduce the combinatoric complexity of intersection detection between facets and quadtree cells.
+We use the following approach:
+
+\subsubsection{Initializing conflict cell detection}
+\label{sec:initconflict}
+Before we detect conflict cells we create a mapping from each quadtree cell $c$ to all facets bounded by $c$.
+We first find the “bounding cell” $c_f$ for a facet $f$, where the bounding cell is the smallest quadtree cell that completely contains $f$.
+For each facet $f$ in parallel we determine $lcp_f$, which is the longest common prefix of the Morton codes of the vertices of $f$.
+Then, to find the bounding cell $c_f$, we iterate in parallel over each facet $f$ and use $lcp_f$ to direct a search through the quadtree.
+Let $F$ be the number of facets.
+We allocate two parallel arrays of size $F$, BCells and FacetMap.
+As each $c_f$ is found, the index of $c_f$ is stored in the BCells array at index equal to the thread id.
+At the same time, we store the index of each facet in the FacetMap.
+Initially, FacetMap[i] = i.
+(see Figure \ref{fig:data-structures-1}).
+
+Next, we perform a parallel radix sort on the parallel arrays (BCells and FacetMap) using the bounding cell addresses as the sort key (figure \ref{fig:data-structures-2}).
+Finally, since each quadtree cell may bound multiple facets, we compute a range for each quadtree cell by comparing neighboring quadtree indices in the mapping in parallel (the (F/L)Facet array in figure 3b).
+We now have a mapping from a quadtree cell $c$ to facets bounded by $c$ (figure \ref{fig:data-structures-3}).
+
+\subsubsection{Conflict Cell Detection}
+\label{sec:detectconflict}
+To identify conflicts, we begin by processing each leaf cell $L$ in parallel using Algorithm 1.
+First, we set $L$'s color to -1, meaning it is unknown whether $L$ is a conflict cell or not.
+Then, we traverse each direct ancestor $A$ of $L$ using a $Parent$ field stored in the quadtree data structure (line \ref{alg:A}).
+For each ancestor traversed, we iterate over the facets bounded by $A$ by using the quadtree cell to facet mapping found in \ref{sec:initconflict}.
+(line \ref{alg:i})
+
+For each facet $f$ discovered this way, we test for intersection between $f$ and $L$.
+If $f$ intersects $L$ and $L$'s color is -1, we copy $f$'s color to $L$.
+Otherwise if $f$ intersects $L$ and $L$'s color does not match $f$'s color, we set $L$'s color to -2, indicating that $L$ is a conflict cell that must be resolved.
+
+%Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from the root quadtree cell to cell $C$, where the local address is a 2-bit (3-bit in 3D) Morton code.
 The address of the root cell is defined as the empty string.
@@
 
-We define a bounding cell (BCell) to be the smallest internal quadtree node which entirely contains a given facet.
+%We define a bounding cell (BCell) to be the smallest internal quadtree node which entirely contains a given facet.
 Given a facet defined by $n$ endpoints $P=\{p_1, p_2, \dots, p_n\}$, the quadtree address of the BCell is the longest common prefix of the Morton codes of the points in $P$.
@@
 
-We begin by constructing an array \texttt{BCells} and sibling array \texttt{FacetMap} (see figure \ref{fig:data-structures-1}), which is done in parallel over all facets.
+%We begin by constructing an array \texttt{BCells} and sibling array \texttt{FacetMap} (see figure \ref{fig:data-structures-1}), which is done in parallel over all facets.
 Each facet $f$ computes the longest common prefix of its vertices and stores the result in \texttt{BCells[$f$]}.
 
-Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras quadtree construction.
+%Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras quadtree construction.
 
@@
 
-Then we use the \texttt{BCells} array and quadtree data structure to find the conflict cells using algorithm \ref{alg:find-conflict-cells}.
+%Then we use the \texttt{BCells} array and quadtree data structure to find the conflict cells using algorithm \ref{alg:find-conflict-cells}.
 We process each leaf cell $L$ in parallel (line \ref{alg:L}).
@@
 
-\subsection{Complexity analysis}
+%-------------------------------------------------------------------------------
+% Results
+%-------------------------------------------------------------------------------
+\section{Results and conclusions}
+Our implementation\footnote{
+Source code will be made available at our website. %is available at \url{http://cedmav.org/research/project/33-gvds.html}.
+}
+of the algorithm supports polygons and polylines which needn't be manifold or connected. All tests were run on
+% a Razer Blade Stealth with
+ an Intel i7 6500u 3.10 GHz dual core processor, 8 GB of memory and an Nvidia GTX 1070 graphics card.
+Figure \ref{fig:results-toy} shows results a simple toy dataset showing conflict cell detection and resolution.
+A very complex dataset with many objects at very different scales is shown in figure \ref{fig:results-maze}.
+It demonstrates that our method can handle datasets far beyond the memory limits of uniform grid approaches while still fully resolving between objects.
+The gears dataset (figure \ref{fig:gears}) again shows a large domain-to-object-spacing ratio, as well as non-convexities.
+The vascular dataset shown in figure \ref{fig:vascular} demonstrates our method on polylines derived from biological image data, which is often noisy with non-manifoldness and intersections.
+Table \ref{tab:timings} shows timings for our implementation compared to the previous state-of-the-art.
+Our implementation is significantly faster and also generates fewer quadtree cells.
+\blue{See \ref{sec:complexity} for a runtime complexity analysis.}
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:simple-conflicts}
+    \includegraphics[width=0.31\columnwidth]{simple-conflict.png} }
+  \subfloat[][]{
+    \label{fig:simple-samples}
+    \includegraphics[width=0.31\columnwidth]{simple-samples.png} }
+
+  \caption{
+    \protect\subref{fig:simple-conflicts} A toy dataset showing conflict cells after building the quadtree from object vertices.
+    \protect\subref{fig:simple-samples} The toy dataset showing how samples are collected.
+  }
+  \label{fig:results-toy}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:ut1-samples}
+    \includegraphics[width=0.31\columnwidth]{ut1-samples.png} }
+  \subfloat[][1x zoom]{
+    \label{fig:ut1}
+    \includegraphics[width=0.31\columnwidth]{ut1.png} }
+  \subfloat[][16x zoom]{
+    \label{fig:ut2}
+    \includegraphics[width=0.31\columnwidth]{ut2.png} } \\
+  \subfloat[][256x zoom]{
+    \label{fig:ut3}
+    \includegraphics[width=0.31\columnwidth]{ut3.png} }
+  \subfloat[][4,000x zoom]{
+    \label{fig:ut4}
+    \includegraphics[width=0.31\columnwidth]{ut4.png} }
+  \subfloat[][60,000x zoom]{
+    \label{fig:ut5}
+    \includegraphics[width=0.31\columnwidth]{ut5.png} }
+  \caption{
+    \protect\subref{fig:ut1-samples} A complex dataset with 470 objects at vastly different scales in object size and spacing.
+    \protect\subref{fig:ut1}-\protect\subref{fig:ut5} Complex dataset at different zoom levels up to 60K magnification.
+This shows the importance of an adaptive method such as a quadtree.
+A uniform grid would require $2^{48}$ cells to resolve between objects.
+The quadtree shown here has 22,429 cells.
+  }
+  \label{fig:results-maze}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:gears-1}
+    \includegraphics[width=0.31\columnwidth]{Nathan/gears/2.PNG} }
+  \subfloat[][]{
+    \label{fig:gears-2}
+    \includegraphics[width=0.31\columnwidth]{Nathan/gears/3.PNG} }
+  \subfloat[][]{
+    \label{fig:gears-3}
+    \includegraphics[width=0.31\columnwidth]{Nathan/gears/5.PNG} }
+  \caption{
+    \protect\subref{fig:gears-1} A dataset of gears with close tolerance. The resolved quadtree with sampled points is shown.
+    \protect\subref{fig:gears-2} Showing just the quadtree and sample points.
+    \protect\subref{fig:gears-3} A zoomed-in image showing the close object spacing compared to the large domain.
+  }
+  \label{fig:gears}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:vascular-1}
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/04-Pruned.PNG} }
+  \subfloat[][]{
+    \label{fig:vascular-2}
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/06-Initial-Conflicts.PNG} }
+  \subfloat[][]{
+    \label{fig:vascular-3}
+%    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/08-Resolved-Quadtree.PNG} }
+    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/14-Resolved-Quadtree.PNG} }
+  \caption{A large set of uniquely labeled polygons constructed from connected component analysis on a photograph of vascular cambium, a type of plant tissue.
+    \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.
+    \protect\subref{fig:vascular-2} All conflict cells of the initial quadtree.
+    \protect\subref{fig:vascular-3} After conflict cell resolution.
+No quadtree cell intersects more than one object.
+Our method works even though objects in this dataset are often non-manifold and have self-intersections.
+  }
+  \label{fig:vascular}
+\end{figure}
+
+% # quadtree vertices
+%  52K ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
+% 170K ./viewer3 -l 12 ~/data/knife/knife-holder.obj ~/data/knife/knife[2-4].obj
+% 146K ./viewer2 -l 24 ../data2/filled_ut/*.dat
+% 2.8M ./viewer3 --uniform-colors -a 1 -l 8 ~/data/neuron/improved/a*.obj
+% 1.3M ./viewer3 -l 8 ~/data/rice-dwarf/decimated2/n1UF2a-*.obj
+
+% memory (54 bytes per quadtree cell)
+%  2.8Mb ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
+%  9.2Mb ./viewer3 -l 12 ~/data/knife/knife-holder.obj ~/data/knife/knife[2-4].obj
+% 7.9Mb ./viewer2 -l 24 ../data2/filled_ut/*.dat
+% 151.2Mb ./viewer3 --uniform-colors -a 1 -l 8 ~/data/neuron/improved/a*.obj
+% 70.2Mb ./viewer3 -l 8 ~/data/rice-dwarf/decimated2/n1UF2a-*.obj
+
+\begin{table*}
+  \centering
+  \footnotesize{
+%  \scriptsize{
+%  \subfloat[][]{
+%    \label{fig:table-sub}
+  \begin{tabular}{lcc|c|cc|cc}
+    \toprule
+    dataset & objects & object          & quadtree
+    & \multicolumn{2}{c|}{time} &  \multicolumn{2}{c}{quad cells} 
+    \\
+            &         & facets          & depth
+    & \multicolumn{2}{c|}{(millisec)} & 
+    \\
+           &         &  & 
+    & Ours & Prev & Ours & Prev \\
+    \hline
+
+    % ./gvd-viewer2 -l 9 ../data2/test1*.dat
+    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/test1*.dat
+    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 9 & 9 & 3 & 138 & 1213  \\
+
+    % ./gvd-viewer2 -l 9 ../data2/glines/200x200/*.dat
+    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/glines/200x200/*
+    Fig. \ref{fig:gears} & 12 & 288 & 10 & 23 & 24 & 408 & 3370  \\
+
+    % ./gvd-viewer2 -l 24 ../data2/maze/*.dat
+    % ./QUADTREE --disable-conflict-color -d 1 -m 3 -l 24 data/maze/poly*.dat
+    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 87 & 277 & 22,355 & 156,649  \\
+
+    % ./gvd-viewer2 -l 8 ../data2/glines/100x100/*.dat
+    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 9 data/glines/100x100/*
+    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 84 & 376 & 14,338 & 157,006  \\
+%    Fig. \ref{fig:vascular} & 2162 & 39K & 12 & 84 & 376 & 14,338 & 157,006  \\
+
+    \bottomrule
+  \end{tabular}}
+%  \subfloat[][]{
+%    \label{fig:timing-chart}
+%    \includegraphics[width=0.8\columnwidth]{timing-chart.png} }
+  \caption{Table of quadtree computation statistics and timings.
+\emph{Ours} is the approach described in this paper and \emph{Prev} is the approach by Edwards \etal \cite{edwards2015approximating}.
+Columns are: \emph{objects} - the number of objects in the dataset; \emph{object facets} - the number of line segments (2D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{time (ms)} - milliseconds to build the quadtree; \emph{quad cells} - number of quadtree cells.
+}
+  \label{tab:timings}
+\end{table*}
+
+As can be seen in table \ref{tab:timings}, there is overhead with our approach: running our algorithm on small datasets yields smaller gains.
+In fact, our approach actually performs worse on the toy dataset.
+The power of our algorithm becomes obvious on large, complex datasets, where our performance time gains are significant.
+
+As noted in the introduction, our continuing work is in fast construction of octrees modeling inter-object space in 3D.
+%We are in the process of integrating our algorithm with animated systems, generating quadtrees in real-time for collision detection, distance transforms, and generalized Voronoi diagram computation.
+Our implementation continues to be refined and optimized, and we expect to shortly have a version with an order of magnitude improvement over the state of the art.
+Importantly, we are also working on an extension to 3D.
+Every step in our method has a straightforward extension to 3D with the exception of point sampling for conflict resolution (see section \ref{sec:resolve}), which is where we are focusing our efforts.
+
+%-------------------------------------------------------------------------------
+% Conclusions
+%-------------------------------------------------------------------------------
+%\section{Conclusions}
+
+%-------------------------------------------------------------------------------
+% Acknowledgements
+%-------------------------------------------------------------------------------
+%\section*{Acknowledgements}
+%Thanks to Kristen Harris for use of the neuronal data and Jonathan Bronson for the heart data.
+The work of JE and VP was supported in part by NSF IIS-1314896, NSF ACI-0904631, DOE/NEUP 120341, DOE/UV-CDAT DESC0006872, DOE/Codesign P01180734, DOE/SciDAC DESC0007446, DOE/PIPER DESC0010498, and DOE/CCMSC DENA0002375.
+This work initiated at the University of Texas when JE, ED  and CB were supported in part by NIH contract R01-EB00487, NSF Grant OCI-1216701 and SNL contract 1439100.
+
+%-------------------------------------------------------------------------
+% Bibliography
+%-------------------------------------------------------------------------
+
+\bibliographystyle{model3-num-names}
+%\bibliographystyle{abbrv}
+\bibliography{paper}
+\balance
+
+\appendix
+
+\blue{
+\section{Complexity analysis}
+\label{sec:complexity}
+}
 
@@
 %\paragraph{Time complexity}
-\subsubsection{Time complexity}
+\subsection*{Time complexity}
 
@@
 %\paragraph{Space complexity}
-\subsubsection{Space complexity}
+\subsection*{Space complexity}
 
@@
 
-%-------------------------------------------------------------------------------
-% Results
-%-------------------------------------------------------------------------------
-\section{Results and conclusions}
-Our implementation\footnote{
-Source code will be made available at our website. %is available at \url{http://cedmav.org/research/project/33-gvds.html}.
-}
-of the algorithm supports polygons and polylines which needn't be manifold or connected. All tests were run on
-% a Razer Blade Stealth with
- an Intel i7 6500u 3.10 GHz dual core processor, 8 GB of memory and an Nvidia GTX 1070 graphics card.
-Figure \ref{fig:results-toy} shows results a simple toy dataset showing conflict cell detection and resolution.
-A very complex dataset with many objects at very different scales is shown in figure \ref{fig:results-maze}.
-It demonstrates that our method can handle datasets far beyond the memory limits of uniform grid approaches while still fully resolving between objects.
-The gears dataset (figure \ref{fig:gears}) again shows a large domain-to-object-spacing ratio, as well as non-convexities.
-The vascular dataset shown in figure \ref{fig:vascular} demonstrates our method on polylines derived from biological image data, which is often noisy with non-manifoldness and intersections.
-Table \ref{tab:timings} shows timings for our implementation compared to the previous state-of-the-art.
-Our implementation is significantly faster and also generates fewer quadtree cells.
-
-\begin{figure}
-  \centering
-  \subfloat[][]{
-    \label{fig:simple-conflicts}
-    \includegraphics[width=0.31\columnwidth]{simple-conflict.png} }
-  \subfloat[][]{
-    \label{fig:simple-samples}
-    \includegraphics[width=0.31\columnwidth]{simple-samples.png} }
-
-  \caption{
-    \protect\subref{fig:simple-conflicts} A toy dataset showing conflict cells after building the quadtree from object vertices.
-    \protect\subref{fig:simple-samples} The toy dataset showing how samples are collected.
-  }
-  \label{fig:results-toy}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfloat[][]{
-    \label{fig:ut1-samples}
-    \includegraphics[width=0.31\columnwidth]{ut1-samples.png} }
-  \subfloat[][1x zoom]{
-    \label{fig:ut1}
-    \includegraphics[width=0.31\columnwidth]{ut1.png} }
-  \subfloat[][16x zoom]{
-    \label{fig:ut2}
-    \includegraphics[width=0.31\columnwidth]{ut2.png} } \\
-  \subfloat[][256x zoom]{
-    \label{fig:ut3}
-    \includegraphics[width=0.31\columnwidth]{ut3.png} }
-  \subfloat[][4,000x zoom]{
-    \label{fig:ut4}
-    \includegraphics[width=0.31\columnwidth]{ut4.png} }
-  \subfloat[][60,000x zoom]{
-    \label{fig:ut5}
-    \includegraphics[width=0.31\columnwidth]{ut5.png} }
-  \caption{
-    \protect\subref{fig:ut1-samples} A complex dataset with 470 objects at vastly different scales in object size and spacing.
-    \protect\subref{fig:ut1}-\protect\subref{fig:ut5} Complex dataset at different zoom levels up to 60K magnification.
-This shows the importance of an adaptive method such as a quadtree.
-A uniform grid would require $2^{48}$ cells to resolve between objects.
-The quadtree shown here has 22,429 cells.
-  }
-  \label{fig:results-maze}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfloat[][]{
-    \label{fig:gears-1}
-    \includegraphics[width=0.31\columnwidth]{Nathan/gears/2.PNG} }
-  \subfloat[][]{
-    \label{fig:gears-2}
-    \includegraphics[width=0.31\columnwidth]{Nathan/gears/3.PNG} }
-  \subfloat[][]{
-    \label{fig:gears-3}
-    \includegraphics[width=0.31\columnwidth]{Nathan/gears/5.PNG} }
-  \caption{
-    \protect\subref{fig:gears-1} A dataset of gears with close tolerance. The resolved quadtree with sampled points is shown.
-    \protect\subref{fig:gears-2} Showing just the quadtree and sample points.
-    \protect\subref{fig:gears-3} A zoomed-in image showing the close object spacing compared to the large domain.
-  }
-  \label{fig:gears}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfloat[][]{
-    \label{fig:vascular-1}
-    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/04-Pruned.PNG} }
-  \subfloat[][]{
-    \label{fig:vascular-2}
-    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/06-Initial-Conflicts.PNG} }
-  \subfloat[][]{
-    \label{fig:vascular-3}
-%    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/08-Resolved-Quadtree.PNG} }
-    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/14-Resolved-Quadtree.PNG} }
-  \caption{A large set of uniquely labeled polygons constructed from connected component analysis on a photograph of vascular cambium, a type of plant tissue.
-    \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.
-    \protect\subref{fig:vascular-2} All conflict cells of the initial quadtree.
-    \protect\subref{fig:vascular-3} After conflict cell resolution.
-No quadtree cell intersects more than one object.
-Our method works even though objects in this dataset are often non-manifold and have self-intersections.
-  }
-  \label{fig:vascular}
-\end{figure}
-
-% # quadtree vertices
-%  52K ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
-% 170K ./viewer3 -l 12 ~/data/knife/knife-holder.obj ~/data/knife/knife[2-4].obj
-% 146K ./viewer2 -l 24 ../data2/filled_ut/*.dat
-% 2.8M ./viewer3 --uniform-colors -a 1 -l 8 ~/data/neuron/improved/a*.obj
-% 1.3M ./viewer3 -l 8 ~/data/rice-dwarf/decimated2/n1UF2a-*.obj
-
-% memory (54 bytes per quadtree cell)
-%  2.8Mb ./viewer3 -l 8 ~/data/gears/gear[1-3].obj
-%  9.2Mb ./viewer3 -l 12 ~/data/knife/knife-holder.obj ~/data/knife/knife[2-4].obj
-% 7.9Mb ./viewer2 -l 24 ../data2/filled_ut/*.dat
-% 151.2Mb ./viewer3 --uniform-colors -a 1 -l 8 ~/data/neuron/improved/a*.obj
-% 70.2Mb ./viewer3 -l 8 ~/data/rice-dwarf/decimated2/n1UF2a-*.obj
-
-\begin{table*}
-  \centering
-  \footnotesize{
-%  \scriptsize{
-%  \subfloat[][]{
-%    \label{fig:table-sub}
-  \begin{tabular}{lcc|c|cc|cc}
-    \toprule
-    dataset & objects & object          & quadtree
-    & \multicolumn{2}{c|}{time} &  \multicolumn{2}{c}{quad cells} 
-    \\
-            &         & facets          & depth
-    & \multicolumn{2}{c|}{(millisec)} & 
-    \\
-           &         &  & 
-    & Ours & Prev & Ours & Prev \\
-    \hline
-
-    % ./gvd-viewer2 -l 9 ../data2/test1*.dat
-    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/test1*.dat
-    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 9 & 9 & 3 & 138 & 1213  \\
-
-    % ./gvd-viewer2 -l 9 ../data2/glines/200x200/*.dat
-    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/glines/200x200/*
-    Fig. \ref{fig:gears} & 12 & 288 & 10 & 23 & 24 & 408 & 3370  \\
-
-    % ./gvd-viewer2 -l 24 ../data2/maze/*.dat
-    % ./QUADTREE --disable-conflict-color -d 1 -m 3 -l 24 data/maze/poly*.dat
-    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 87 & 277 & 22,355 & 156,649  \\
-
-    % ./gvd-viewer2 -l 8 ../data2/glines/100x100/*.dat
-    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 9 data/glines/100x100/*
-    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 84 & 376 & 14,338 & 157,006  \\
-%    Fig. \ref{fig:vascular} & 2162 & 39K & 12 & 84 & 376 & 14,338 & 157,006  \\
-
-    \bottomrule
-  \end{tabular}}
-%  \subfloat[][]{
-%    \label{fig:timing-chart}
-%    \includegraphics[width=0.8\columnwidth]{timing-chart.png} }
-  \caption{Table of quadtree computation statistics and timings.
-\emph{Ours} is the approach described in this paper and \emph{Prev} is the approach by Edwards \etal \cite{edwards2015approximating}.
-Columns are: \emph{objects} - the number of objects in the dataset; \emph{object facets} - the number of line segments (2D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{time (ms)} - milliseconds to build the quadtree; \emph{quad cells} - number of quadtree cells.
-}
-  \label{tab:timings}
-\end{table*}
-
-As can be seen in table \ref{tab:timings}, there is overhead with our approach: running our algorithm on small datasets yields smaller gains.
-In fact, our approach actually performs worse on the toy dataset.
-The power of our algorithm becomes obvious on large, complex datasets, where our performance time gains are significant.
-
-As noted in the introduction, our continuing work is in fast construction of octrees modeling inter-object space in 3D.
-%We are in the process of integrating our algorithm with animated systems, generating quadtrees in real-time for collision detection, distance transforms, and generalized Voronoi diagram computation.
-Our implementation continues to be refined and optimized, and we expect to shortly have a version with an order of magnitude improvement over the state of the art.
-Importantly, we are also working on an extension to 3D.
-Every step in our method has a straightforward extension to 3D with the exception of point sampling for conflict resolution (see section \ref{sec:resolve}), which is where we are focusing our efforts.
-
-%-------------------------------------------------------------------------------
-% Conclusions
-%-------------------------------------------------------------------------------
-%\section{Conclusions}
-
-%-------------------------------------------------------------------------------
-% Acknowledgements
-%-------------------------------------------------------------------------------
-%\section*{Acknowledgements}
-%Thanks to Kristen Harris for use of the neuronal data and Jonathan Bronson for the heart data.
-The work of JE and VP was supported in part by NSF IIS-1314896, NSF ACI-0904631, DOE/NEUP 120341, DOE/UV-CDAT DESC0006872, DOE/Codesign P01180734, DOE/SciDAC DESC0007446, DOE/PIPER DESC0010498, and DOE/CCMSC DENA0002375.
-This work initiated at the University of Texas when JE, ED  and CB were supported in part by NIH contract R01-EB00487, NSF Grant OCI-1216701 and SNL contract 1439100.
-
-%-------------------------------------------------------------------------
-% Bibliography
-%-------------------------------------------------------------------------
-
-\bibliographystyle{model3-num-names}
-%\bibliographystyle{abbrv}
-\bibliography{paper}
-\balance
-
 \end{document}

### 2017-05-05 paper.tex
Added that no atomic operations are required for algorithm 1, as per reviewer one's request
updated timings to not include OpenGL overhead
Responding to reviewer comments. Changes in blue.

@@
 
-\paragraph{Parallel} Many recent works on fast quadtree construction using the GPU are limited to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data}.
+\paragraph{Parallel} Many recent works on fast quadtree construction using the GPU are limited either to point sites \cite{bedorf2012sparse,karras2012maximizing,zhou2011data} \blue{or to sites that don't overlap octree cells \cite{li2014fully}.}
 %Kim and Liu's work~\shortcite{kim2014exact} is similar, computing the quadtree on the barycenters of triangles, giving an approximation of an object-resolving quadtree.
- Most quadtree approaches that support surfaces \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed} are designed for efficient rendering, and actual construction of the quadtree is implemented on the CPU.
-Two works \cite{bastos2008gpu,park2010cuda} implement Adaptive Distance Fields in parallel on quadtrees but building the quadtree itself is done sequentially.
-Yin \etal~\shortcite{yin2011fast} compute the octree entirely on the GPU using a bottom-up approach by initially subdividing into a complete quadtree, resulting in memory usage that is no better than using a uniform grid.
+\blue{Most quadtree approaches that support surfaces are designed for efficient rendering and not inter-object resolution.
+Most of these approaches construct the quadtree on the CPU \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed}, although Choi \etal \cite{choi2010parallel} succeed in constructing k-D trees in parallel.} Two works \cite{bastos2008gpu,park2010cuda} implement Adaptive Distance Fields in parallel on quadtrees but building the quadtree itself is done sequentially.
+Yin \etal~\shortcite{yin2011fast} \blue{and Crassin and Green \cite{crassin2012octree}} compute the octree entirely on the GPU using a bottom-up approach by initially subdividing into a complete octree, resulting in memory usage that is no better than using a uniform grid.
 The most similar work to what we do here is Kim and Liu's method~\shortcite{kim2014exact}, which computes the quadtree on the barycenters of triangles, giving an approximation of our quadtree, but without fully resolving between objects.
-We are unaware of any GPU quadtree construction methods that are fully adaptive and resolve the seperation between objects.
+We are unaware of any GPU quadtree construction methods that are fully adaptive and resolve between objects.
 
@@
 We refer to quadtree leaf cells that intersect two or more objects as ``conflict cells.'' A necessary and sufficient condition for a quadtree to resolve objects is to have no conflict cells.
-Our approach to computing such a quadtree is to first build an initial quadtree, called the ``vertex quadtree,'' using a set $S$ of point samples.
+Our approach to computing such a quadtree is \blue{in two stages.
+We first build an initial quadtree, called the ``vertex quadtree,'' using a set $S$ of point samples.
 We initialize $S$ to be the object vertices.
-We then detect conflict cells in parallel, followed by augmenting $S$ with sample points such that a subsequent quadtree built on $S$ resolves conflict cells.
-If $S$ changed, then we iterate (see section \ref{sec:iterate}).
+The second stage is to detect conflict cells in parallel, followed by augmenting $S$ with sample points such that a subsequent quadtree built on $S$ resolves conflict cells.} If $S$ changed, then we iterate (see section \ref{sec:iterate}) \blue{which is necessary only if a conflict cell has multiple intersecting objects and is minimized by starting from an initial vertex quadtree}.
+\blue{This two-stage approach enabled by our sample point computation is the first method that can resolve between objects fully in parallel regardless of object spacing, i.e., we do not iterate through levels of the quadtree, subdividing as we go.}
 
@@
 
+\blue{
 \subsection{Identifying conflict cells}
@@
 To efficiently identify conflict cells, we take advantage of the space filling Morton curve and the existing quadtree hierarchy to reduce the combinatoric complexity of intersection detection between facets and quadtree cells.
-We use the following approach:
-
-\subsubsection{Initializing conflict cell detection}
+We use the following approach.}
+
+\blue{\subsubsection{Initializing conflict cell detection}
 \label{sec:initconflict}
-Before we detect conflict cells we create a mapping from each quadtree cell $c$ to all facets bounded by $c$.
-We first find the “bounding cell” $c_f$ for a facet $f$, where the bounding cell is the smallest quadtree cell that completely contains $f$.
+Before we detect conflict cells we create a mapping from each quadtree cell $c$ to all facets bounded by $c$, a technique similar in spirit to the fragment emission and sorting done by Pantaleoni \cite{pantaleoni2011voxelpipe}.
+We first find the ``bounding cell'' $c_f$ for a facet $f$, where the bounding cell is the smallest quadtree cell that completely contains $f$.
 For each facet $f$ in parallel we determine $lcp_f$, which is the longest common prefix of the Morton codes of the vertices of $f$.
@@
 Let $F$ be the number of facets.
-We allocate two parallel arrays of size $F$, BCells and FacetMap.
-As each $c_f$ is found, the index of $c_f$ is stored in the BCells array at index equal to the thread id.
-At the same time, we store the index of each facet in the FacetMap.
-Initially, FacetMap[i] = i.
-(see Figure \ref{fig:data-structures-1}).
-
-Next, we perform a parallel radix sort on the parallel arrays (BCells and FacetMap) using the bounding cell addresses as the sort key (figure \ref{fig:data-structures-2}).
-Finally, since each quadtree cell may bound multiple facets, we compute a range for each quadtree cell by comparing neighboring quadtree indices in the mapping in parallel (the (F/L)Facet array in figure 3b).
-We now have a mapping from a quadtree cell $c$ to facets bounded by $c$ (figure \ref{fig:data-structures-3}).
+We allocate two parallel arrays of size $F$, \texttt{BCells} and \texttt{FacetMap}.
+As each $c_f$ is found, the index of $c_f$ is stored in the \texttt{BCells} array at index equal to the thread id.
+At the same time, we store the index of each facet in the \texttt{FacetMap}.
+Initially, {FacetMap[i] = i} (see figure \ref{fig:data-structures-1}).}
+
+\blue{
+Next, we perform a parallel radix sort on the parallel arrays (\texttt{BCells} and \texttt{FacetMap}) using the bounding cell addresses as the sort key (figure \ref{fig:data-structures-2}).
+Finally, since each quadtree cell may bound multiple facets, we compute a range for each quadtree cell by comparing neighboring quadtree indices in the mapping in parallel (the \texttt{(F/L)Facet} array in figure \ref{fig:data-structures-2}).
+We now have a mapping from a quadtree cell $c$ to facets bounded by $c$ (figure \ref{fig:data-structures-3}).}
 
@@
 \label{sec:detectconflict}
-To identify conflicts, we begin by processing each leaf cell $L$ in parallel using Algorithm 1.
+To identify conflicts, we begin by processing each leaf cell $L$ in parallel using Algorithm \ref{alg:find-conflict-cells}.
 First, we set $L$'s color to -1, meaning it is unknown whether $L$ is a conflict cell or not.
 Then, we traverse each direct ancestor $A$ of $L$ using a $Parent$ field stored in the quadtree data structure (line \ref{alg:A}).
-For each ancestor traversed, we iterate over the facets bounded by $A$ by using the quadtree cell to facet mapping found in \ref{sec:initconflict}.
-(line \ref{alg:i})
+For each ancestor traversed, we iterate over the facets bounded by $A$ by using the quadtree cell to facet mapping computed in \ref{sec:initconflict} (line \ref{alg:i}).
 
@@
 Otherwise if $f$ intersects $L$ and $L$'s color does not match $f$'s color, we set $L$'s color to -2, indicating that $L$ is a conflict cell that must be resolved.
+Note that in Algorithm \ref{alg:find-conflict-cells}, no atomic operations are required.
 
@@
     \protect\subref{fig:conflict-resolution-x} A conflict cell with two lines from different objects.
-    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least one sample (red dots).
+    \protect\subref{fig:conflict-resolution-even}-\protect\subref{fig:conflict-resolution-all} Fitting boxes such that any box intersecting both lines contains at least one sample (red dots).
     \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least two samples.
 This ensures that a quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-quadtree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
+    \blue{\protect\subref{fig:conflict-resolution-adjacent-even}-\protect\subref{fig:conflict-resolution-adjacent-all} The adjacent case.}
   }
@@
 We begin by fitting the smallest box centered on a point $p$ that intersects both $q$ and $r$.
-We break the problem into two cases:
+\blue{The smallest box sampled at point $p(s)$ has edge length $a(s)$ as shown in figure \ref{fig:conflict-resolution-even}.
+We break the problem of finding $a(s)$ into two cases:}
 %\begin{tightenumerate}
 \begin{enumerate}
- \item The \textit{opposite} case (see figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
- \item In the \textit{adjacent} case (see figure \ref{fig:conflict-resolution-adjacent-even}), $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
+ \item The \textit{opposite} case (figure \ref{fig:conflict-resolution-even}) is where $w^y > 0$, so each box intersects $q$ and $r$ at its top-left and bottom-right corners, respectively.
+ \item In the \textit{adjacent} case (figure \ref{fig:conflict-resolution-adjacent-even}), $w^y < 0$, so the line intersections are adjacent at the top-left and bottom-left corners of the box.
 %\end{tightenumerate}
@@
 \section{Results and conclusions}
-Our implementation\footnote{
+Our \blue{OpenCL} implementation\footnote{
 Source code will be made available at our website. %is available at \url{http://cedmav.org/research/project/33-gvds.html}.
@@
 
-\begin{table*}
+%\begin{table*}
+%  \centering
+%  \footnotesize{
+%%  \scriptsize{
+%%  \subfloat[][]{
+%%    \label{fig:table-sub}
+%  \begin{tabular}{lcc|c|cc|cc}
+%    \toprule
+%    dataset & objects & object          & quadtree
+%    & \multicolumn{2}{c|}{time} &  \multicolumn{2}{c}{quad cells} 
+%    \\
+%            &         & facets          & depth
+%    & \multicolumn{2}{c|}{(millisec)} & 
+%    \\
+%           &         &  & 
+%    & Ours & Prev & Ours & Prev \\
+%    \hline
+%
+%    % ./gvd-viewer2 -l 9 ../data2/test1*.dat
+%    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/test1*.dat
+%    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 9 & 9 & 3 & 138 & 1213  \\
+%
+%    % ./gvd-viewer2 -l 9 ../data2/glines/200x200/*.dat
+%    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/glines/200x200/*
+%    Fig. \ref{fig:gears} & 12 & 288 & 10 & 23 & 24 & 408 & 3370  \\
+%
+%    % ./gvd-viewer2 -l 24 ../data2/maze/*.dat
+%    % ./QUADTREE --disable-conflict-color -d 1 -m 3 -l 24 data/maze/poly*.dat
+%    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 87 & 277 & 22,355 & 156,649  \\
+%
+%    % ./gvd-viewer2 -l 8 ../data2/glines/100x100/*.dat
+%    % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 9 data/glines/100x100/*
+%    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 84 & 376 & 14,338 & 157,006  \\
+%%    Fig. \ref{fig:vascular} & 2162 & 39K & 12 & 84 & 376 & 14,338 & 157,006  \\
+%
+%    \bottomrule
+%  \end{tabular}}
+%%  \subfloat[][]{
+%%    \label{fig:timing-chart}
+%%    \includegraphics[width=0.8\columnwidth]{timing-chart.png} }
+%  \caption{Table of quadtree computation statistics and timings.
+\emph{Ours} is the approach described in this paper and \emph{Prev} is the approach by Edwards \etal \cite{edwards2015approximating}.
+Columns are: \emph{objects} - the number of objects in the dataset; \emph{object facets} - the number of line segments (2D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{time (ms)} - milliseconds to build the quadtree; \emph{quad cells} - number of quadtree cells.
+}
+%  \label{tab:timings}
+%\end{table*}
+
+\begin{table}
   \centering
   \footnotesize{
-%  \scriptsize{
-%  \subfloat[][]{
-%    \label{fig:table-sub}
-  \begin{tabular}{lcc|c|cc|cc}
+  \begin{tabular}{lcc|c|cc}
     \toprule
     dataset & objects & object          & quadtree
-    & \multicolumn{2}{c|}{time} &  \multicolumn{2}{c}{quad cells} 
-    \\
+    & \multicolumn{2}{c}{time} \\
             &         & facets          & depth
-    & \multicolumn{2}{c|}{(millisec)} & 
-    \\
-           &         &  & 
-    & Ours & Prev & Ours & Prev \\
+    & \multicolumn{2}{c}{(millisec)} \\
+           &&&
+    & Ours & Prev \\
     \hline
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/test1*.dat
-    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 9 & 9 & 3 & 138 & 1213  \\
+    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 9 & 9 & 3 \\
 
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/glines/200x200/*
-    Fig. \ref{fig:gears} & 12 & 288 & 10 & 23 & 24 & 408 & 3370  \\
+    Fig. \ref{fig:gears} & 12 & 288 & 10 & 23 & 24 \\
 
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 3 -l 24 data/maze/poly*.dat
-    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 87 & 277 & 22,355 & 156,649  \\
+    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 87 & 277 \\
 
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 9 data/glines/100x100/*
-    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 84 & 376 & 14,338 & 157,006  \\
+    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 84 & 376 \\
 %    Fig. \ref{fig:vascular} & 2162 & 39K & 12 & 84 & 376 & 14,338 & 157,006  \\
@@
 \emph{Ours} is the approach described in this paper and \emph{Prev} is the approach by Edwards \etal \cite{edwards2015approximating}.
-Columns are: \emph{objects} - the number of objects in the dataset; \emph{object facets} - the number of line segments (2D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{time (ms)} - milliseconds to build the quadtree; \emph{quad cells} - number of quadtree cells.
-}
+Columns are: \emph{objects} - the number of objects in the dataset; \emph{object facets} - the number of line segments (2D) of all objects in the dataset; \emph{quadtree depth} - required quadtree depth in order to resolve objects; \emph{time (ms)} - milliseconds to build the quadtree }
   \label{tab:timings}
-\end{table*}
+\end{table}
 
@@
 The power of our algorithm becomes obvious on large, complex datasets, where our performance time gains are significant.
+
+\begin{figure}
+  \centering
+%  \subfloat[][]{
+%    \label{fig:timings-data}
+%    % trim={<left> <lower> <right> <upper>}
+%    \includegraphics[width=0.3\columnwidth]{facet-test-image} }\\
+  \subfloat[][]{
+    \label{fig:timings1}
+    % trim={<left> <lower> <right> <upper>}
+    \includegraphics[trim={4mm 0 40mm 0},clip,width=0.48\columnwidth]{facetxtime} }
+  \subfloat[][]{
+    \label{fig:timings2}
+    \includegraphics[trim={4mm 0 40mm 0},clip,width=0.48\columnwidth]{facetxcells.png} }
+  \caption{Scaling tests. The dataset used is a square domain with increasing numbers of non-intersecting lines.
+%    \protect\subref{fig:timings-data} Data used in the scaling study. Lines are shortened and added to the perimeter of the square to increase the number of facets.
+    \protect\subref{fig:timings1} Our algorithm consistently performs an order of magnitude faster than the algorithm of Edwards \etal \cite{edwards2015approximating} as the number of facets increases.
+    \protect\subref{fig:timings2} The quadtree that we build is roughly equal in size to that of the previous work.
+  }
+  \label{fig:timings}
+\end{figure}
+
+Figure \ref{fig:timings} shows the results of a scaling study, where we increased the number of objects and facets by orders of magnitude.
+Our algorithm consistently shows timings an order of magnitude faster than the state of the art.
+The approach of Edwards \etal failed on datasets with $10^6$ facets or more.
 
@@
 \bibliography{paper}
-\balance
+%\balance
 

### 2017-05-05 paper.tex
Added kernel stages flow chart. Added implementation section. Updated data structures figure.

@@
 %-------------------------------------------------------------------------------
+% Implementation
+%-------------------------------------------------------------------------------
+\section{Implementation}
+\blue{We have implemented\footnote{
+Source code will be made available at our website. %is available at \url{http://cedmav.org/research/project/33-gvds.html}.
+}
+our algorithm using OpenCL.
+Figure \ref{fig:flow-chart} shows the stages of the major kernels.
+Every kernel call is parallel on vertices, facets, quadtree nodes or Morton code bits.
+Our implementation uses 64-bit Morton codes, which were sufficient for all of our datasets.
+There is no way of knowing \textit{a priori} what the object spacing is going to be, however, 64-bit codes were sufficient for the most demanding datasets while retaining reasonable radix-sort timings.}
+
+\blue{Our implementation of the algorithm supports polygons and polylines which needn't be manifold or connected.
+Intersecting lines are not handled as a special case, i.e., the quadtree is simply resolved to its maximum depth.}
+
+\begin{figure}
+  \centering
+  % trim={<left> <lower> <right> <upper>}
+  \includegraphics[trim={0 50mm 0 0},clip,width=0.95\columnwidth]{kernel-stages.pdf}
+  \caption{Kernel calls, with some calls omitted for clarity.
+The name of the kernel is in larger font while the the elements on which the parallelism runs are given in smaller font.
+The majority of calls are facet- or vertex-parallel.
+  }
+  \label{fig:flow-chart}
+\end{figure}
+
+%-------------------------------------------------------------------------------
 % Results
@@
 \section{Results and conclusions}
-Our \blue{OpenCL} implementation\footnote{
-Source code will be made available at our website. %is available at \url{http://cedmav.org/research/project/33-gvds.html}.
-}
-of the algorithm supports polygons and polylines which needn't be manifold or connected. All tests were run on
+All tests were run on
 % a Razer Blade Stealth with
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/test1*.dat
-    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 9 & 9 & 3 \\
+    Fig. \ref{fig:simple-conflicts} & 5 & 24 & 9 & 5 & 3 \\
 
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 10 data/glines/200x200/*
-    Fig. \ref{fig:gears} & 12 & 288 & 10 & 23 & 24 \\
+    Fig. \ref{fig:gears} & 12 & 288 & 10 & 8 & 24 \\
 
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 3 -l 24 data/maze/poly*.dat
-    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 87 & 277 \\
+    Fig. \ref{fig:ut1-samples} & 470 & 4943 & 24 & 40 & 277 \\
 
@@
     % ./QUADTREE --disable-conflict-color -d 1 -m 4 -l 9 data/glines/100x100/*
-    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 84 & 376 \\
+    Fig. \ref{fig:vascular} & 2162 & 39,338 & 12 & 36 & 376 \\
 %    Fig. \ref{fig:vascular} & 2162 & 39K & 12 & 84 & 376 & 14,338 & 157,006  \\

### 2017-05-05 paper.tex
Added discussion of Morton code size selection to implementation section

@@
 Our implementation uses 64-bit Morton codes, which were sufficient for all of our datasets.
-There is no way of knowing \textit{a priori} what the object spacing is going to be, however, 64-bit codes were sufficient for the most demanding datasets while retaining reasonable radix-sort timings.}
+There is no way of knowing \textit{a priori} what the object spacing is going to be, however, 64-bit codes were sufficient for the most demanding datasets while retaining reasonable radix-sort timings.
+Choosing a large number of bits for the Morton code results in little or no wasted effort in later refinement.
+Increasing the size of the Morton codes affects the amount of later pruning only if it results in extraneous resolution of vertices in the initial vertex quadtree computation.
+The number of bits also has no effect on the number of conflict cells unless the Morton codes are too small to resolve vertices.}
 

### 2017-05-06 paper.tex
added discussion of the use of an atomic for propagating brt colors for pruning

@@
 We then propagate the BRT labels up the tree in parallel, marking ``unknown'' nodes as ``required'' when the labels of the current node's two child nodes don't match.
-Finally, we generate quadtree nodes from only the required internal binary radix tree nodes.
+\blue{Labels are applied using an atomic compare and swap, and threads terminate if the current ancestor's label was previously ``unknown''.} Finally, we generate quadtree nodes from only the required internal binary radix tree nodes.
 

### 2017-05-06 paper.tex
Layout changes

@@
     \label{fig:conflict-resolution-x}
-    \includegraphics[width=0.35\columnwidth]{conflict-resolution-x.pdf} }
+    \includegraphics[width=0.3\columnwidth]{conflict-resolution-x.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-even}
-    \includegraphics[width=0.35\columnwidth]{conflict-resolution-even.pdf} }\\
+    \includegraphics[width=0.3\columnwidth]{conflict-resolution-even.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-all}
-    \includegraphics[width=0.35\columnwidth]{conflict-resolution-all.pdf} }
+    \includegraphics[width=0.3\columnwidth]{conflict-resolution-all.pdf} }\\
   \subfloat[][]{
     \label{fig:conflict-resolution-quadtree}
-    \includegraphics[width=0.35\columnwidth]{conflict-resolution-octree.pdf} }\\
+    \includegraphics[width=0.3\columnwidth]{conflict-resolution-octree.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-adjacent-even}
-    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-even.pdf} }
+    \includegraphics[width=0.3\columnwidth]{conflict-resolution-adjacent-even.pdf} }
   \subfloat[][]{
     \label{fig:conflict-resolution-adjacent-all}
-    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-all.pdf} }
+    \includegraphics[width=0.3\columnwidth]{conflict-resolution-adjacent-all.pdf} }
   \caption{

### 2017-05-06 paper.tex
Layout changes

@@
 We initialize $S$ to be the object vertices.
-The second stage is to detect conflict cells in parallel, followed by augmenting $S$ with sample points such that a subsequent quadtree built on $S$ resolves conflict cells.} If $S$ changed, then we iterate (see section \ref{sec:iterate}) \blue{which is necessary only if a conflict cell has multiple intersecting objects and is minimized by starting from an initial vertex quadtree}.
-\blue{This two-stage approach enabled by our sample point computation is the first method that can resolve between objects fully in parallel regardless of object spacing, i.e., we do not iterate through levels of the quadtree, subdividing as we go.}
+The second stage is to detect conflict cells in parallel, followed by augmenting $S$ with sample points such that a subsequent quadtree built on $S$ resolves conflict cells.} If $S$ changed, then we iterate (see section \ref{sec:iterate}) \blue{which is necessary only if a conflict cell has multiple intersecting objects.
+The number of iterations is minimized by starting from an initial vertex quadtree}.
+\blue{This two-stage approach enables us to resolve between objects fully in parallel regardless of object spacing, i.e., we do not iterate through levels of the quadtree, subdividing as we go.}
 
@@
 There is no way of knowing \textit{a priori} what the object spacing is going to be, however, 64-bit codes were sufficient for the most demanding datasets while retaining reasonable radix-sort timings.
-Choosing a large number of bits for the Morton code results in little or no wasted effort in later refinement.
-Increasing the size of the Morton codes affects the amount of later pruning only if it results in extraneous resolution of vertices in the initial vertex quadtree computation.
+Choosing a large number of bits for the Morton code results in little or no wasted effort in later refinement, affecting the amount of later pruning only if intra-object vertex spacing is small compared to spacing between objects.
 The number of bits also has no effect on the number of conflict cells unless the Morton codes are too small to resolve vertices.}
@@
 \blue{Our implementation of the algorithm supports polygons and polylines which needn't be manifold or connected.
-Intersecting lines are not handled as a special case, i.e., the quadtree is simply resolved to its maximum depth.}
+Intersecting lines are not handled as a special case, i.e., the quadtree is simply resolved to its maximum depth.
+Special handling can be implemented per application as needed, e.g., for collision detection applications.}
 
@@
 
-\begin{figure}
+\begin{figure}[t]
   \centering
@@
 
+Figure \ref{fig:timings} shows the results of a scaling study, where we increased the number of objects and facets by orders of magnitude.
+Our algorithm consistently shows timings an order of magnitude faster than the state of the art.
+The approach of Edwards \etal failed on datasets with $10^6$ facets or more.
+
+As noted in the introduction, our continuing work is in fast construction of octrees modeling inter-object space in 3D.
+%We are in the process of integrating our algorithm with animated systems, generating quadtrees in real-time for collision detection, distance transforms, and generalized Voronoi diagram computation.
+Our implementation continues to be refined and optimized, and we expect to shortly have a version with an order of magnitude improvement over the state of the art.
+Importantly, we are also working on an extension to 3D.
+Every step in our method has a straightforward extension to 3D with the exception of point sampling for conflict resolution (see section \ref{sec:resolve}), which is where we are focusing our efforts.
+
 \begin{figure}
@@
 \end{figure}
-
-Figure \ref{fig:timings} shows the results of a scaling study, where we increased the number of objects and facets by orders of magnitude.
-Our algorithm consistently shows timings an order of magnitude faster than the state of the art.
-The approach of Edwards \etal failed on datasets with $10^6$ facets or more.
-
-As noted in the introduction, our continuing work is in fast construction of octrees modeling inter-object space in 3D.
-%We are in the process of integrating our algorithm with animated systems, generating quadtrees in real-time for collision detection, distance transforms, and generalized Voronoi diagram computation.
-Our implementation continues to be refined and optimized, and we expect to shortly have a version with an order of magnitude improvement over the state of the art.
-Importantly, we are also working on an extension to 3D.
-Every step in our method has a straightforward extension to 3D with the exception of point sampling for conflict resolution (see section \ref{sec:resolve}), which is where we are focusing our efforts.
-
-%-------------------------------------------------------------------------------
-% Conclusions
-%-------------------------------------------------------------------------------
-%\section{Conclusions}
 

### 2017-05-17 paper.tex
final changes

@@
 
-\title{Parallel Quadtree Construction on Collections of Objects}
+\title{Parallel quadtree construction on collections of objects}
 %\title{Fast Construction of Inter-Object Space Representations}
 
-%\author{Nathan Morrical}
-%\author{John Edwards}
-%\address{Department of Informatics and Computer Science, Idaho State University, USA}
+\author{Nathan Morrical}
+\ead{bitinat2@isu.edu}
+\author{John Edwards\corref{cor1}}%\fnref{fn1}}
+\ead{edwajohn@isu.edu}
+\cortext[cor1]{Corresponding author}
+%\fntext[fn1]{Editor-in-Chief, Computers and Graphics Journal.}
+
+%\address{Idaho State University, 921 S 8th Ave, Pocatello, ID, USA}
+\address{Idaho State University, Pocatello, ID, USA}
+
+%\author[label1]{Nathan Morrical}
+%\author[label1]{John Edwards\corref{Corresponding author}}
+%\address[label1]{Department of Informatics and Computer Science, Idaho State University, USA}
 
@@
 Most of these approaches construct the quadtree on the CPU \cite{baert2013out,crassin2009gigavoxels,laine2011efficient,lefebvre2007compressed}, although Choi \etal \cite{choi2010parallel} succeed in constructing k-D trees in parallel.} Two works \cite{bastos2008gpu,park2010cuda} implement Adaptive Distance Fields in parallel on quadtrees but building the quadtree itself is done sequentially.
-Yin \etal~\shortcite{yin2011fast} \blue{and Crassin and Green \cite{crassin2012octree}} compute the octree entirely on the GPU using a bottom-up approach by initially subdividing into a complete octree, resulting in memory usage that is no better than using a uniform grid.
+Yin \etal~\shortcite{yin2011fast} compute the octree entirely on the GPU using a bottom-up approach by initially subdividing into a complete octree, resulting in memory usage that is no better than using a uniform grid.
+Crassin and Green \cite{crassin2012octree} build the octree top-down by performing subdivisions at each level.
 The most similar work to what we do here is Kim and Liu's method~\shortcite{kim2014exact}, which computes the quadtree on the barycenters of triangles, giving an approximation of our quadtree, but without fully resolving between objects.
